### 2014-03-27 paper.tex
added everything

%\documentclass[a4paper,10pt]{article}
%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
\documentclass[10pt,twocolumn,twoside]{IEEEtran}
%\documentclass[journal]{IEEEtran}

\usepackage{mathbf-abbrevs}
\input{defs}

\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{pgfplots}

%\usepackage{xr}
%\externaldocument{paper2}

\usepackage{amsmath,amsfonts,amssymb, amsthm, bm}

\usepackage[square,comma,numbers,sort&compress]{natbib}

\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\sinc}{\operatorname{sinc}}
\newcommand{\rect}[1]{\operatorname{rect}\left(#1\right)}

%opening
\title{Fast sparse period estimation}
\author{R.~G.~McKilliam, I.~V.~L.~Clarkson and B.~G.~Quinn    
\thanks{
R.~G.~McKilliam is with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.  I.~V.~L.~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.  B.~G.~Quinn is with the Department of Statistics, Macquarie University, Sydney, NSW, 2109, Australia.
}}

\begin{document}

\maketitle

\begin{abstract}

The problem of estimating the period of a point process from observations that are both sparse and noisy is considered.  By sparse it is meant that only a potentially small unknown subset of the process is observed.  By noisy it is meant that the subset that is observed, is observed with some error, or noise.  Existing accurate algorithms for estimating the period require $O(N^2)$ operations where $N$ is the number of observations.  By quantising the observations we produce an estimator that requires only $O(N\log N)$ operations by use of the chirp z-transform or the fast Fourier transform.  The quantisation has the adverse affect of decreasing the accuracy of the estimator.  This is investigated by Monte-Carlo simulation.  The simulations indicate that significant computational savings are possible with negligible loss in statistical accuracy.

\end{abstract}
\begin{IEEEkeywords}
Period estimation, fast Fourier transform
\end{IEEEkeywords}

\section{Introduction}

Consider the periodic sequence of real numbers
\begin{equation}\label{eq:periodicprocess}
T_0 k + \theta_0, \qquad k \in \ints
\end{equation}
with period $T_0$ and phase $\theta_0$.  For example, the sequence might represent times at which the beat occurs in a musical score.  A problem of interest is to estimate $T_0$ and $\theta_0$ from $N$ observations $y_1,\dots,y_N$ of the form
\begin{equation} \label{eq:sigmodel}
y_n = T_0 s_n + \theta_0 + w_n, \qquad n = 1,\dots,N
\end{equation}
where $w_1,\dots,w_N$ represent noise and $s_1,\dots,s_N$ are unknown integers representing the subset of the periodic sequence that is observed.  The $y_1, \dots, y_N$ are \emph{sparse}, owing to the unknown integers $s_1,\dots,s_N$, and \emph{noisy}, owing to $w_1,\dots,w_N$, observations of the sequence~\eqref{eq:periodicprocess}.  For example, $y_1,\dots,y_N$ might represent times at which a bass drum hit is observed.  The drum hits might not occur at every beat and might occur off beat on some occasions.  Our aim in this case would be to estimate the beat of the music from the times of the drum hits.  The model is depicted in Figure~\ref{fig_stat_model}. %It is clear that any estimator of $T_0$ and $\theta_0$ must, either explicitly or implicitly, also produce and estimate of the unknow integers $s_1,\dots,s_N$ describing the subset of the periods that are observed.

The problem of estimating $T_0$ is called the \emph{sparse, noisy period estimation problem}~\cite{Clarkson2007,McKilliam2007}.  Solutions have applications to synchronisation for telecommunications devices~\cite{Fogel1988,Fogel1989_bit_synch_zero_crossings,Sidiropoulos2005,5621928}, estimation of the pulse repetition interval in electronic support~\cite{EltonGray_puilse_train_rader_1994,Gray_more_pri_1994,Clarkson_thesis,clarkson_estimate_period_pulse_train_1996,Hauochan_pri_2012}, and beat and tempo estimation in music~\cite{dixon_beat_extraction_2001}.  The problem is closely related to spike interval estimation in neurology~\cite{Arnett_neuro_pri_1976,Brillinger_spike_trains_1988,Rossoni200630} and a similar model has been proposed for the detection of anomolies, such as denial-of-service attacks, in internet traffic~\cite{He_detecting_periodic_patterns_in_internet_2006,5585849,5947313}.  The problem is also connected with classical problems in number theory such as Diophantine approximation, lattice reduction, and the study of prime numbers~\cite{Cassels_geom_numbers_1997,490557,Clarkson_thesis,Lenstra1982,Wubben_2011,Casey1995,726812,CaseySadler_primes_2013}.

Least squares estimation of period has been studied in~\cite{Clarkson2007,McKilliam2007,Quinn_sparse_noisy_SSP_2012}.  The least squares estimator can be closely approximated using $O(N^2)$ operations by using special purpose algorithms from computational number theory~\cite{McKilliam2009CoxeterLattices,McKilliam2008,McKilliam2008b}.  The least squares estimator is also the maximum likelihood estimator in the case that the noise variables $w_1,\dots,w_N$ are independent and identically normally distributed.  An alternative estimator, called the~\emph{periodogram estimator}, was studied by Fogel and Gavish~\cite{Fogel1988,Fogel1989_bit_synch_zero_crossings} and involves the maximisation of Bartlett's periodogram for point processes~\cite{Bartlest_periodgram_point_process_1963}.  The statistical properties of the periodogram estimator were recently studied by Quinn~et.~al.~\cite{Quinn20013asilomar_period_est} who show the estimator to be strongly consistent and asymptotically normal under mild conditions on the noise $w_1,\dots,w_N$ and integers $s_1,\dots,s_N$.  The fastest existing algorithms can closely approximate the periodogram estimator in $O(N^2)$ operations.  We describe the periodogram estimator in Section~\ref{sec:peri-estim}.

In Section~\ref{sec:quant-peri} we show that, by quantising the observations $y_1,\dots,y_N$, an approximation to the periodogram estimator can be computed using only $O(N\log N)$ operations.  This is facilitated by either the chirp z-transform~\cite{Rabiner1969} or the fast Fourier transform.  The quantisation has the adverse affect of decreasing the accuracy of the estimator.  The estimator is accurate if a sufficiently fine quantiser is chosen, but this requires a larger transform.  The trade off between quantisation level, accuracy, and computational complexity is investigated by computer simulation in Section~\ref{sec:simulations}.  The simulations indicate that significant computational savings are possible with negligible loss in accuracy.  Sadler and Casey~\cite{726812} have previously suggested use of the fast Fourier transform for period estimation.  However, their estimator requires far larger transforms and produces less accurate results than the estimator described here (see Section~VI and Figures 1-4 in~\cite{726812}).  In contrast, our estimator has accuracy similar to the least squares estimator, i.e., similar to the maximum likelihood estimator if the noise variables $w_1,\dots,w_N$ are assumed independent and identically normally distributed.

{
\def\vertgap{2}
\def\ph{0.4}
\def\T{1.1}
\newcommand{\raxis}{\draw[->] (-0.25,0) -- (8,0) node[above] {$\reals$}; \draw (0,-0.06)-- node[below] {$0$} (0,0.06) }
\newcommand{\pulse}[1]{ \draw[->,>=latex] (#1,0) -- (#1,1) }
\newcommand{\pulsewithnode}[2]{ \draw[->,>=latex] (#1,0) -- node[right] {#2} (#1,1) }
\begin{figure}[t]
	\centering
\begin{tikzpicture}[font=\small]
    %\draw[very thin,color=gray] (-0.1,-1.1) grid (3.9,3.9);
\begin{scope}
    \raxis;
    \foreach \k in {0,1,...,6} { \k, \pulse{\T*\k+\ph}; }
    \draw[<->] (\T*2+\ph,1.2) -- node[above] {$T_0$} (\T*3+\ph,1.2) ;
    \draw[<->] (0.0,1.2) -- node[above] {$\theta_0$} (\ph,1.2) ; 
%    \draw[->] (-0.25,-2*\vertgap) -- (8,-2*\vertgap) node[above] {$\reals$};
\end{scope}
\begin{scope}[yshift=-\vertgap cm]
  \raxis;
  \node[below] at (\T*1+\ph-0.3,0) {$y_1$}; \pulsewithnode{\T*1+\ph-0.3}{$s_1=1$}; \draw[->] (\T*1+\ph,1.2) -- node[above] {$w_1$} (\T*1+\ph-0.3,1.2) ; 
  \node[below] at (\T*3+\ph+0.5,0) {$y_2$}; \pulsewithnode{\T*3+\ph+0.5}{$s_2=3$}; \draw[->] (\T*3+\ph,1.2) -- node[above] {$w_2$} (\T*3+\ph+0.5,1.2) ;
  \node[below] at (\T*6+\ph-0.4,0) {$y_3$}; \pulsewithnode{\T*6+\ph-0.4}{$s_3=6$};\draw[->] (\T*6+\ph,1.2) -- node[above] {$w_3$} (\T*6+\ph-0.4,1.2) ;
\end{scope}
\end{tikzpicture}
%		\includegraphics{figs/figures-1.mps}
		\caption{Sparse noisy observations of a periodic sequence.  The original sequence~\eqref{eq:periodicprocess} is shown on the top and sparse noisy observations according to~\eqref{eq:sigmodel} are shown on the bottom.  In this example the observed subset is represented by integers $s_1=1$, $s_2=3$ and $s_3=6$.}
		\label{fig_stat_model}
\end{figure}
}

\section{The periodogram estimator}\label{sec:peri-estim}

Before describing the periodogram estimator we state some required assumptions.  Observe that
\[
y_n = T_0 s_n + \theta_0 + w_n = \frac{T_0}{2} r_n + \theta_0 + w_n
\]
where $r_n = 2s_n \in \ints$.  Thus, to ensure identifiability, it is necessary to assume that $T_0$ lies in some interval $(T_{\text{min}}, T_{\text{max}})$ where $T_{\text{max}} < 2T_{\text{min}}$.  We also assume that $y_1,\dots,y_N$ are in ascending order.  This assumption can be made without loss of generality since it amounts only to sorting the observations if required.

%BLERG. $T_{\text{min}}, T_{\text{max}}$ identifiability and $y_1,\dots,y_N$ in accending order.

Fogel and Gavish~\cite{Fogel1988,Fogel1989_bit_synch_zero_crossings} proposed the estimator $\hat{T} = 1/\hat{f}$ where $\hat{f}$ is the maximiser of the~\emph{periodogram}
\[
I_y(f) = \babs{ \sum_{n=1}^N e^{ 2\pi j f y_n} }^2 = \babs{ \sum_{n=1}^N e^{ -2\pi j f y_n} }^2 
\]
over the interval $(f_{\text{min}}, f_{\text{max}}) = (1/T_{\text{max}}, 1/T_{\text{min}})$ and $j = \sqrt{-1}$.  That is,
\[
\hat{f} = \arg\max_{f \in (f_{\text{min}}, f_{\text{max}})} I_y(f).
\] 
This is called the \emph{periodogram estimator}.  This definition of the periodogram differs from the common definition in the spectral and frequency estimation literature~\cite{Hannan_time_series_1967,Quinn2001}.  The statistical properties of the periodogram estimator were examined in~\cite{Quinn20013asilomar_period_est}.

The periodogram $I_y$ is not convex and has many local maxima.  In order to compute $\hat{f}$ it is usual to first compute $I_y$ on a grid of frequencies
\[
G = \{ f_{\text{min}} + Dk \mid k = 0, \dots K-1 \},
\]
where $D$ is the distance between frequencies in the grid, 
\[
K =  \floor{\frac{f_{\text{max}} - f_{\text{min}}}{D}}
\]
is the number of frequencies in the grid, and $\floor{\cdot}$ denotes the largest integer less than or equal to its argument.  
Let $\widetilde{f}$ be the maximiser of $I_y$ over $G$, that is, let
\[
\widetilde{f} = \arg\max_{f \in G} I_y(f).
\]
If the distance $D$ is sufficiently small, then $\widetilde{f}$ is a good approximation to the maximiser $\hat{f}$.  An iterative procedure, such as the Newton-Raphson method, %, or Brent's method~\cite{Brent_opt_no_derivs_1973} 
can then be used to compute $\hat{f}$ starting from $\widetilde{f}$.

In~\cite{McKilliam2007} it was observed that $D$ must shrink as $N$ increases in order for $\widetilde{f}$ to be a sufficiently good approximation to $\hat{f}$.  This was recently studied by Ye~et.~al.~\cite{Haohuan2013435,6492742} who heuristically suggest choosing
\begin{equation}\label{eq:gridwidthrecommended}
D = \frac{1}{2(y_N - y_1)}
\end{equation}
which is $O(N^{-1})$ under reasonable assumptions about the integers $s_1,\dots,s_N$ and the noise $w_1,\dots,w_N$.  This choice is supported by the simulations in~\cite{Haohuan2013435,6492742} and also those in Section~\ref{sec:simulations}.  The asymptotic standard deviation of $\hat{f}$ is of order $O(N^{-3/2})$ as shown in~\cite{Quinn20013asilomar_period_est} and this initially suggests that $D = O(N^{-1})$ is too large.  However, the choice~\eqref{eq:gridwidthrecommended} does appear effective in practice.  A statistically rigorous justification for this choice of $D$ is beyond the scope of the present paper, but could potentially be developed using the techniques in~\cite{Quinn2008maximizing_the_periodogram}.

With this choice for $D$ the number of elements in the grid grows linearly with $N$, that is, $K = O(N)$.  Computing $I_y(f)$ for each $f$ requires $O(N)$ operations and so computing $I_y(f)$ for all $f \in G$ requires $O(N^2)$ operations.  It follows that computation of $\widetilde{f}$ requires $O(N^2)$ operations and that the periodogram estimator $\hat{T} = 1/\hat{f}$ requires at least $O(N^2)$ operations to compute by this method.  %The choice for $D$ from~\eqref{eq:gridwidthrecommended} might not guarantee that application of a numerical optimisation proceedure, starting from $\widetilde{f}$, will result in $\hat{f}$.  For this reason, this approach can only be considered approximate.  However, the simulations presented in~\cite{Haohuan2013435,6492742} and those in Section~\ref{sec:simulations} suggest that~\eqref{eq:gridwidthrecommended} the approximation is very accurate.  
In the next section we describe a modification of the periodogram estimator.  The modification is such that the fast Fourier transform or the chirp z-transform~\cite{Rabiner1969} can be used to compute an approximation $\widetilde{f}^\prime$ to $\hat{f}$ with only $O(N\log N)$ operations.

\section{Quantisation and the periodogram}\label{sec:quant-peri}

Denote by $\round{x}$ the nearest integer to $x$.  The direction of rounding for half integers is not important.  Half integers are rounded up in our implementation.  Let $q$ be a real number and define the quantisation function
\[
Q(x) = \round{qx}/q
\]
that rounds its argument to a nearest multiple of $1/q$.  Large $q$ corresponds with fine quantisation and small $q$ corresponds with coarse quantisation.  Compute quantised observations
\[
z_n = Q(y_n) =  \ell_n/q, \qquad n = 1,\dots,N
\]
where $\ell_n = \round{qy_n} \in \ints$.  The periodogram of the quantised observations $z_1,\dots,z_N$ is
\begin{align*}
I_z(f) = \babs{ \sum_{n=1}^N e^{ -2\pi j f z_n} }^2 = \babs{ \sum_{n=1}^N e^{ -2\pi j f \ell_n/q} }^2.
\end{align*}
Let $\widetilde{f}^\prime$ be the maximiser of $I_z$ over the grid $G$, that is,
\[
\widetilde{f}^\prime = \arg\max_{f \in G} I_z(f).
\]
If the quantisation parameter $q$ is sufficiently large we expect that $I_z$ is close to $I_y$ and so $\widetilde{f}^\prime$ will be close to $\widetilde{f}$.  Thus, if $q$ is sufficiently large and the distance $D$ is sufficiently small we expect $\widetilde{f}^\prime$ to be a good approximation to $\hat{f}$.  An iterative procedure %or Brent's method~\cite{Brent_opt_no_derivs_1973} 
can then be used to compute $\hat{f}$ starting from $\widetilde{f}^\prime$.  
Observe that $I_z$ is periodic with period $q$.  This immediately indicates that we should choose the quantisation fine enough that 
\begin{equation}\label{eq:boundaliasingP}
q > f_{\text{max}} - f_{\text{min}}.
\end{equation}
Otherwise there would exist frequencies in the interval $(f_{\text{min}}, f_{\text{max}})$ that are aliases with respect to $I_z$.  %That is, there would exists two frequences $f_1,f_2 \in (f_{\text{min}}, f_{\text{max}})$ such that $f_1 = f_2 + 1/P$ and $I_z(f_1) = I_z(f_2)$.

The maximiser $\widetilde{f}^\prime$ can be computed efficiently using the chirp z-transform.  Let $b_i$ be the number of integers from $\ell_1,\dots,\ell_N$ that are equal to $i$.  That is,
\[
b_i= \#\{ n \mid \ell_n = i, n = 1,\dots, N\}
\]
where $\#$ indicates the number of elements in a set.  The integers $\ell_1,\dots,\ell_N$ are in ascending order because $y_1,\dots,y_N$ are in ascending order.  So $b_i = 0$ if $i < \ell_1$ or $i > \ell_N$.  Now
\[
I_z(f) = \babs{ \sum_{i = \ell_1}^{\ell_N} e^{ -2\pi j f i/q} b_i }^2 = \babs{ \sum_{m = 0}^{\ell_N-\ell_1} e^{ -2\pi j f m/q} b_{m+\ell_1} }^2.
\]
We want to compute $I_z(f)$ for all $f$ in the grid $G$.  Putting
\[
v_m = e^{ -2\pi j f_{\text{min}}m/q} b_{m+\ell_1}
\]
we have
\[
I_z(f_{\text{min}} + Dk) = \babs{ \sum_{m = 0}^{\ell_N-\ell_1} e^{ -2\pi j Dk m/q} v_m }^2 = \abs{V_k}^2
\]   
where, using the notation of Rabiner~et.~al.~\cite[eq.~(10)]{Rabiner1969},
\[
V_k = \sum_{m = 0}^{\ell_N-\ell_1} e^{ -2\pi j Dk m/q} v_m = \sum_{m=0}^{\ell_N-\ell_1} A^{-m} W^{km} v_m
\]
where $A = 1$ and $W = e^{ -2\pi j D/q}$.  All of $V_0,\dots,V_{K-1}$ can be efficiently computed by the chirp z-transform~\cite{Rabiner1969}.  Putting
\[
\hat{k} = \arg\max_{k = 0,1,\dots,K-1} \abs{ V_k }^2
\]
we have
\[ 
\widetilde{f}^\prime = f_{\text{min}} + D\hat{k} = \arg\max_{f \in G} I_z(f).
\]
The frequency $\widetilde{f}^\prime$ can be used to initialize an iterative procedure to maximise $I_y$.  The procedure will converge to $\hat{f}$ if $\widetilde{f}^\prime$ is sufficiently close to $\hat{f}$.

An alternative to the chirp z-transform is to use a single fast Fourier transform.  We have found this to be faster than the chirp z-transform for all but exceptionally large $N$.  Suppose that we choose the distance $D$ and quantisation parameter $q$ so that $M = q/D$ is an integer and
\[
M = \frac{q}{D} > \ell_N-\ell_1.
\]
With this choice
\[
I_z(f_{\text{min}} + Dk) = \sabs{\sum_{m = 0}^{M-1} e^{-2\pi j k m/M} v_m}^2 = \sabs{V_k}^2
\]
where $V_0,\dots,V_{M-1}$ are the Fourier coefficients of $v_0,\dots,v_{M-1}$ and can be computed using a fast Fourier transform of length $M$.  From inequality~\eqref{eq:boundaliasingP} we have
\[
M = \frac{q}{D} > \frac{f_{\text{max}} - f_{\text{min}}}{D} \geq \floor{\frac{f_{\text{max}} - f_{\text{min}}}{D}} = K
\]
and so, the required values  $V_0,\dots,V_{K-1}$ are the first $K$ Fourier coefficients.

\section{Computational complexity}\label{sec:comp-compl}

The previous section showed how $\widetilde{f}^\prime$ could be computed using either the chirp z-transform or a fast Fourier transform of length $M$.  We now describe recommended values for the parameters $q, M$, and $D$.  The simulations presented in Section~\ref{sec:simulations} suggest that a reasonable value for the quantisation parameter is $q = 4 f_{\text{max}}$.  If the chirp z-transform is used then $D$ is chosen according to~\eqref{eq:gridwidthrecommended}.  Computation of the chirp z-transform requires the computation of three fast Fourier transforms, each of length $L + K - 1$, where 
\[
L = \ell_N - \ell_1 + 1 = \round{qy_N} - \round{qy_1} + 1.
\]
If instead a single fast Fourier transform is used then our recommended value for the length of the transform is $M= 2L$.  In this case the grid distance
\[ 
D = \frac{q}{M} = \frac{q}{2(\round{qy_N} - \round{qy_1} + 1)} \approx  \frac{1}{2(y_N - y_1)}
\]
is approximately equal to~\eqref{eq:gridwidthrecommended}.

Observe that $L$ grows proportionally with $q$ and, under reasonable assumptions about the integers $s_1,\dots,s_N$ and the noise variables $w_1,\dots,w_N$, we have $L = O(qN)$.  The most computationally expensive part of the chirp z-transform is the computation of three fast Fourier transforms of length $L+K-1$.  Alternatively, a single fast Fourier transform of length $M = 2L$ can be used.  Both approaches require $O(qN\log(qN)) = O(N\log N)$ operations because $q$ does not depend on $N$.  However, it is apparent that the length of the transforms, and hence the complexity, increases as the quantisation becomes finer, i.e., as $q$ increases.  This presents a trade off between computationally complexity and quantisation accuracy.  This trade off is investigated in the next section.

\begin{figure}[t]
  \centering 
  \begin{tikzpicture}
    \selectcolormodel{gray} 
    \begin{axis}[font=\footnotesize,xmode=log,ymode=log,height=7cm,width=9cm,xlabel={$N$},ylabel={time (ms)},ylabel style={at={(0.215,0.41)}},xlabel style={at={(0.5,0.205)}}, legend style={draw=none,fill=none,legend pos=north west,cells={anchor=west},font=\footnotesize}]
      \addplot[mark=o,mark options={scale=1}] table {code/data/PeriodogramBenchmark};
      \addplot[mark=x,mark options={solid,fill=black,scale=1}] table {code/data/LeastSquaresBenchmark};
      \addplot[mark=*,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramq4.0Benchmark};
      \addplot[mark=triangle,mark options={solid,fill=black,scale=0.9}] table {code/data/QuantisedPeriodogramChirpZq4.0Benchmark};
      %\addplot[mark=square,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpBenchmark};
      \addplot[mark=square,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2newBenchmark};
      \legend{Periodogram, Least squares, Fast Fourier transform, Chirp z-transform, SLS2-new}
   \end{axis}
  \end{tikzpicture}  
  \caption{Computation time in milliseconds versus $N$ for the periodogram, least squares, and quantised periodogram estimators computed using the chirp z-transform and a single fast Fourier transform.}\label{plot:benchmark}
\end{figure}

\begin{figure*}[t] 
  \centering 
  \begin{tikzpicture} 
    \selectcolormodel{gray} 
    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.19,0.28)}},xlabel style={at={(0.5,0.14)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=6, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30geom3}; 
            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q1.5geom3};
      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN30q4.0geom3};
%      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN30};
      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN30geom3};
       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/NormalisedSamplingLLSN30geom3}; 
 %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN30geom3};
      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN30geom3};
%      \addplot[mark=none,dashed,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q2};
 %     \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200};
 %     \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200}; 
 %     \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/PeriodogramN200};
 %     \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q3};
 %     \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q10};
%      \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/NormalisedSamplingLLSN200};
      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN1200geom3};
 %     \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1200};
       \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN1200geom3};
       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/NormalisedSamplingLLSN1200geom3};
 %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1200geom3};
       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN1200geom3};
       \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q1.5geom3};
      %\addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q3};
      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q4.0geom3};
 %     \legend{Periodogram Theory, Least squares theory, Quantised periodogram, Periodogram, Least squares}
     \legend{Periodogram theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=4 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
     \node[rotate=15,font=\footnotesize] at (axis cs:6e-4,1e-8) {$N=30$};
     \node[rotate=15,font=\footnotesize] at (axis cs:7e-4,2e-13) {$N=1200$}; 
   \end{axis} 
 \begin{axis}[name=plot2,at={(11,0)},font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.19,0.28)}},xlabel style={at={(0.5,0.14)}}]
      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30geom15}; 
            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q1.5geom15};
      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN30q4.0geom15};
%      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN30};
      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN30geom15};
       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/NormalisedSamplingLLSN30geom15}; 
  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN30geom15};
       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN30geom15};
%      \addplot[mark=none,dashed,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q2};
 %     \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200};
 %     \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200}; 
 %     \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/PeriodogramN200};
 %     \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q3};
 %     \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q10};
%      \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/NormalisedSamplingLLSN200}; 
      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN1200geom15};
 %     \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1100};
       \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN1200geom15};
       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/NormalisedSamplingLLSN1200geom15};
 %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1100geom15};
       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN1200geom15};
       \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q1.5geom15};
      %\addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1100q3};
      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q4.0geom15};
 %     \legend{Periodogram Theory, Least squares theory, Quantised periodogram, Periodogram, Least squares}
     %\legend{Periodogram theory, $q=\tfrac{3}{2}T_{\text{max}}$, $q=4T_{\text{max}}$, Periodogram, Least squares, SLS2-adj}
     \node[rotate=15,font=\footnotesize] at (axis cs:6e-4,4e-10) {$N=30$};
     \node[rotate=15,font=\footnotesize] at (axis cs:7e-4,0.6e-14) {$N=1200$}; 
   \end{axis}
  \end{tikzpicture}  
  \caption{Mean square period error versus noise variance $\sigma^2$.  The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=2,\dots,N$ are independent and identically geometrically distributed with mean $\mu = 3$ in the plot on the left and mean $\mu=15$ in the plot on the right.}\label{plot:multipleN}
\end{figure*}

\section{Simulations} \label{sec:simulations}

This section presents the results of Monte-Carlo simulations with the periodogram estimator~\cite{Haohuan2013435,Fogel1989_bit_synch_zero_crossings}, the modified least squares estimator~\cite{Clarkson2007,McKilliam2007}, Clarkson's modification of the separable least squares line search (SLS2-new) estimator~\cite{Clarkson2007,Sidiropoulos2005}, and the quantised periodogram estimator from Section~\ref{sec:quant-peri}.  The grid distance $D$ used for the modified least squares, periodogram, and SLS2-new estimators is given by~\eqref{eq:gridwidthrecommended}.  The quantised periodogram estimator is computed using a single fast Fourier transform of length $M = 2L = 2(\ell_N - \ell_N + 1)$.  Both the periodogram and quanitised periodogram estimators use the Newton-Raphson method.  In all simulations the noise variables $w_1,\dots,w_N$ are independent and identically normally distributed with variance $\sigma^2$.  The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=2,\dots,N$ are independent and identically geometrically distributed starting from $1$ with means $\mu=3$ and $15$.  The true period is $T_0 = \pi/3$, the true phase is $\theta_0 = 0.2$, and $T_{\text{min}} = 0.8$ and $T_{\text{max}} = 1.5$.

Figure~\ref{plot:multipleN} shows the mean square error of the estimators versus $\sigma^2$ for $N = 30$ and $N = 1200$.  In the left hand  plot $\mu=3$ and in the right hand plot $\mu=15$.  Thus, on average $\tfrac{1}{3}$ of the elements of the point process~\eqref{eq:periodicprocess} are observed in the left hand  plot and only $\tfrac{1}{15}$ in the right hand plot.  Four thousand Monte-Carlo trials are run for each value of $\sigma^2$.  The plots show the mean square error of the quantised periodogram estimator for two quantisation parameters $q = \tfrac{3}{2}f_{\text{max}} = \tfrac{15}{8}$ and $q = 4 f_{\text{max}} = 5$.  The estimator performs poorly when $q = \tfrac{3}{2}f_{\text{max}}$ (triangles) but performs almost identically to the original periodogram estimator (circles) when $q = 4 f_{\text{max}}$ (dots).  For comparison purposes, the plot also shows the mean square error of the modified least squares estimator~\cite{Clarkson2007,McKilliam2007} and the SLS2-new estimator~\cite{Clarkson2007,Sidiropoulos2005}.  %The statistical performance of the SLS2-new estimator is poor when compared with the other estimators.  %For small $\sigma^2$ the mean square error of the SLS2 estimator is almost three orders of magnitude less accurate when $N = 50$ and almost five orders of magnitude less accurate when $N=1000$.   
The solid line shows the mean square error of the periodogram estimator predicted by Theorem~2 from~\cite{Quinn20013asilomar_period_est}.

Figure~\ref{plot:benchmark} shows the average running time for the estimators for $N$ in the range $10$ to $25000$ when $\mu=3$.  The running times when $\mu=15$ lead to similar conclusions and are omitted.  The quantisation parameter is $q = 4 f_{\text{max}}$ so that the performance of the quantised periodogram estimator is similar to that of the original periodogram estimator as observed in Figure~\ref{plot:multipleN}.  The average running times for the quantised periodogram estimator computed using the chirp z-transform and using a single fast Fourier transform are shown.  The chirp z-transform is slower in these simulations, but is expected to be faster when $N$ is exceptionally large.  The quantised periodogram estimator computed using a single fast Fourier transform is faster than the periodogram estimator when $N \geq 20$.  It is faster than the least squares estimator when $N \geq 80$ and it is faster than the SLS2-new estimator when $N \geq 400$.  The quantised periodogram estimator is significantly faster than the other estimator for large $N$.  For example, when $N = 25000$ the periodogram estimator requires approximately 530 seconds, whereas the quantised periodogram estimator requires approximately 960 milliseconds.  The software is written in Java and the computer used is an Intel Core2 Q6600 running at \unit[2.4]{GHz}.

\section{Conclusion}

We have considered the problem of estimating the period of a point process from $N$ observations that are both sparse and noisy.  By quantising the observations we describe an estimator that can be computed in $O(N\log N)$ operations by use of the fast Fourier transform or chirp z-transform.  This is in contrast to existing accurate estimators that require at least $O(N^2)$ operations.  The trade off between computational complexity and quantisation accuracy was examined by simulation.  The simulations indicate that significant computational savings are possible with negligible loss in statistical accuracy.

The new estimator has potential application to the synchronisation of communications devices~\cite{Fogel1988,Fogel1989_bit_synch_zero_crossings}, pulse repetition interval estimation in electronic support~\cite{EltonGray_puilse_train_rader_1994,Gray_more_pri_1994,Clarkson_thesis,clarkson_estimate_period_pulse_train_1996,Hauochan_pri_2012}, beat and tempo estimation in music~\cite{dixon_beat_extraction_2001}, spike interval estimation in neurology~\cite{Arnett_neuro_pri_1976,Brillinger_spike_trains_1988,Rossoni200630}, and in the analysis of internet traffic~\cite{He_detecting_periodic_patterns_in_internet_2006,5585849,5947313}.

%Currently, as is common An open theoretical question is a rigorously

\small
%\bibliography{zotero}
\bibliography{bib}

\end{document}

%%% Local Variables:
%%% TeX-source-correlate-mode: t
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2014-03-27 build.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+
+#recurse() {
+#    for D in `find -type d -not -name .`
+#   do
+#	pushd $D
+#	recurse
+#	popd
+#    done
+#    for f in *.mp
+#    do
+#	mpost -interaction=nonstopmode ${f}
+#    done
+#}
+
+#compile all metapost files in this directory
+#recurse .
+
+pdflatex --shell-escape paper.tex
+bibtex paper
+pdflatex --shell-escape paper.tex
+pdflatex --shell-escape paper.tex

### 2014-03-27 code/benchmark.scala

@@ -1,0 +1,90 @@
+/**
+* Run benchmarks.
+*/
+
+import pubsim.distributions.discrete.GeometricRandomVariable
+import pubsim.distributions.GaussianNoise
+import pubsim.distributions.RealRandomVariable
+import snpe.SparseNoisyPeriodicSignal
+import snpe.PRIEstimator
+import snpe.YeSampling
+import snpe.PeriodogramEstimator
+import snpe.NormalisedSamplingLLS
+import snpe.QuantisedPeriodogramFFT
+import snpe.QuantisedPeriodogramChirpZ
+import snpe.SLS2novlp
+import snpe.SLS2new
+
+val T_0 = scala.math.Pi/3 //true period
+val T_min = 0.8 //minimum period
+val T_max = 1.5 //maximum period
+val f_min = 1.0/T_max //minimum frequency
+val f_max = 1.0/T_min //maximum frequency
+val theta_0 = 0.2 //true phase
+val p = 1.0/3.0 //parameter of geometric random variable for generating sparse noise
+def discretedist = new GeometricRandomVariable(p) //the discrete distribution for generating different sN - sN-1
+def noisedist = new GaussianNoise(0,0.01) //the noise distribution we use with variance v
+
+//time used for benchmark warmup in nanoseconds.  Benchmark runtime is approximately twice this.
+val MIN_BENCH_DURATION : Long = 5000000000L; // (10 secs)  
+
+val Ns = List(10,14,20,30,40,50,60,80,100,120,150,200,250,300,400,600,900,1200,1500,2000,2600,3500,4700,6000,7800,10000,13000,18000,25000)
+//val Ns = List(10,14,20,30,40,50,60,80,100,120,150,200,250,300,400,600,900,1200,1500,2000,2600,3500,4700,6000,7800,10000,13000,18000,25000,35000,50000,70000)
+//val Ls = List(1000)
+//val Ns = (1 to 40).map(i => (5*scala.math.pow(1.1,i)).toInt).distinct //number of symbols
+
+//runbench(Ns, (N : Int) => new YeSampling(new PeriodogramEstimator(N,T_min,T_max)), "Periodogram")
+//runbench(Ns, (N : Int) => new YeSampling(new NormalisedSamplingLLS(N,T_min,T_max)), "LeastSquares")
+//runbench(Ns, (N : Int) => new YeSampling(new SLS2novlp(N,T_min,T_max)), "SLS2novlp")
[…]

### 2014-03-27 code/java/src/snpe/AbstractPRIEstimator.java

@@ -1,0 +1,50 @@
+package snpe;
+
+/**
+ * Abstract class for period estimators.  Includes a least squares phase estimator 
+ * and various recommended final definitions
+ * @author Robby McKilliam
+ */
+public abstract class AbstractPRIEstimator implements PRIEstimator {
+    
+    public final int N;
+    public final double Tmin, Tmax;
+    public final double fmin, fmax;
+    public final PhaseEstimator phasestor;
+    
+    public AbstractPRIEstimator(int N, double Tmin, double Tmax){
+        this.N = N;
+        this.Tmin = Tmin;
+        this.Tmax = Tmax;
+        fmin = 1/Tmax;
+        fmax = 1/Tmin;
+        phasestor = new PhaseEstimator(N);
+    }
+    
+    @Override
+    public int getLength() {
+        return N;
+    }
+    
+    /** Abstract class for period estimators that sample their objective function */
+    public static abstract class Sampled extends AbstractPRIEstimator implements PRIEstimator.Sampled {
+        
+        public final int NUM_SAMPLES; //number of samples used by the periodogram estimator
+        
+        public Sampled(int N, double Tmin, double Tmax, int samples){
+            super(N,Tmin,Tmax);
+            NUM_SAMPLES = samples;
+        }
+        
+        /// Default estimate uses NUM_SAMPLES set during construction
+        @Override
+        public void estimate(Double[] y) {
+            double fmin = 1 / Tmax;
+            double fmax = 1 / Tmin;
+            double W = (fmax - fmin) / NUM_SAMPLES; //search grid width
+            estimate(y, W);
+        }
+
+    }
+    
+}

### 2014-03-27 code/java/src/snpe/NormalisedPeriodogram.java

@@ -1,0 +1,88 @@
+package snpe;
+
+// Written by Vaughan Clarkson, 11-Jan-07 (fork from PeriodogramEstimator).
+public class NormalisedPeriodogram extends PeriodogramEstimator  {
+
+    public NormalisedPeriodogram(int N, double Tmin, double Tmax, int samples) {
+        super(N, Tmin, Tmax, samples);
+    }
+
+    /**
+     * Default number of samples is 4*N
+     */
+    public NormalisedPeriodogram(int N, double Tmin, double Tmax) {
+        this(N, Tmin, Tmax, 4 * N);
+    }
+
+    private static double calculateObjective(Double[] y, double f) {
+        double sumur = 0, sumui = 0;
+        int N = y.length;
+        for (int i = 0; i < y.length; i++) {
+            sumur += Math.cos(2 * Math.PI * f * y[i]);
+            sumui += Math.sin(2 * Math.PI * f * y[i]);
+        }
+        return (N * N - sumur * sumur - sumui * sumui) / (f * f);
+    }
+
+    @Override
+    public void estimate(Double[] y, double fstep) {
+        // Coarse search
+        double minp = Double.POSITIVE_INFINITY;
+        double fhat = fmin;
+        for (double f = fmin; f <= fmax; f += fstep) {
+            double p = calculateObjective(y, f);
+            if (p < minp) {
+                minp = p;
+                fhat = f;
+            }
+        }
+        fhat = refine(fhat, y); //refine coarse frequency estimate to local maximiser of normalised periodogram
+        That = 1.0 / fhat;
+        phat = phasestor.getPhase(y, That); //now compute the phase estimate
+    }
+
+    ///Modified Newton step minimses the normalised periodogram
+    @Override
+    protected double refine(double fhat, Double[] y) {
+        double minp = calculateObjective(y, fhat);
+        int numIter = 0;
+        double f = fhat;
+        double lastf = f - 2 * EPSILON;
+        double lastp = Double.POSITIVE_INFINITY;
+        while (Math.abs(f - lastf) > EPSILON && numIter++ < MAX_ITER && f >= fmin && f <= fmax) {
+            double sumur = 0, sumui = 0, sumvr = 0, sumvi = 0, sumwr = 0, sumwi = 0;
+            for (int i = 0; i < N; i++) {
+                double ur = Math.cos(2 * Math.PI * f * y[i]);
+                double ui = Math.sin(2 * Math.PI * f * y[i]);
+                double vr = -2 * Math.PI * y[i] * ui;
[…]

### 2014-03-27 code/java/src/snpe/NormalisedSamplingLLS.java

@@ -1,0 +1,58 @@
+/*
+ * NormalisedSamplingLLS.java
+ *
+ * Created on 16 July 2007, 21:57
+ */
+package snpe;
+
+import pubsim.VectorFunctions;
+import pubsim.lattices.Anstar.AnstarVaughan;
+
+/**
+ * Version of the SamplingLLS that uses the modified liklihood function.
+ *
+ * @author Robby McKilliam
+ */
+public class NormalisedSamplingLLS extends SamplingLLS {
+
+    public NormalisedSamplingLLS(int N, double Tmin, double Tmax, int samples) {
+        super(N,Tmin,Tmax,samples);
+    }
+
+    /** Default number of samples is 4N */
+    public NormalisedSamplingLLS(int N, double Tmin, double Tmax) {
+        this(N,Tmin,Tmax,4*N);
+    }
+
+    @Override
+    public void estimate(Double[] y, double fstep) {
+
+        //first compute the period estimate
+        AnstarVaughan.project(y, zeta);
+        double ztz = VectorFunctions.sum2(zeta);
+        double bestL = Double.POSITIVE_INFINITY;
+        double fhat = fmin;
+        for (double f = fmin; f <= fmax; f += fstep) {
+            for (int i = 0; i < N; i++) fzeta[i] = f * zeta[i];
+            lattice.nearestPoint(fzeta);
+            double[] v = lattice.getLatticePoint();
+            double vtz = 0, vtv = 0;
+            for (int i = 0; i < N; i++) {
+                vtv += v[i] * v[i];
+                vtz += v[i] * zeta[i];
+            }
+            //double f0 = vtv / vtz;
+            double f0 = vtz / ztz;
+            double L = f0 * f0 * ztz - 2 * f0 * vtz + vtv;
+            if (L < bestL) {
+                bestL = L;
+                fhat = f0;
+            }
+        }
+        That = 1 / fhat;
+
+        //now compute the phase estimate
+        phat = phasestor.getPhase(y, That);
+    }
+    
+}

### 2014-03-27 code/java/src/snpe/NormalisedZnLLS.java

@@ -1,0 +1,79 @@
+/*
+ * NormalisedZnLLS.java
+ *
+ * Created on 17 July 2007, 07:44
+ */
+
+package snpe;
+
+import pubsim.lattices.Anstar.AnstarVaughan;
+
+/**
+ * ZnLLS with the unormalised log-likelihood function, ie. it finds the
+ * closest point to the line rather than the point of minimal angle.
+ * @author Robby McKilliam
+ */
+public class NormalisedZnLLS extends ZnLLS implements PRIEstimator {
+
+    public NormalisedZnLLS(int N, double Tmin, double Tmax) {
+        super(N, Tmin, Tmax);
+    }
+
+    @Override
+    public void estimate(Double[] y){
+
+        AnstarVaughan.project(y, z);
+
+        double bestf = 0.0;
+        double mindist = Double.POSITIVE_INFINITY;
+
+        for(int i = 0; i <=n ; i++){
+            map.clear();
+            glueVector(i);
+
+            //setup map and variables for this glue vector
+            double ztz = 0.0, ztv = 0.0, vtv = 0.0;
+            for(int j=0; j<=n; j++){
+                v[j] = Math.round(fmin*z[j] - g[j]) + g[j];
+                map.put(new Double((Math.signum(z[j])*0.5 + v[j])/z[j]), new Integer(j));
+                ztz += z[j]*z[j];
+                ztv += z[j]*v[j];
+                vtv += v[j]*v[j];
+            }
+
+            //double f = vtv/ztv;
+            double f = ztv / ztz;
+            //line search loop
+            while(f < fmax){
+
+                double dist = f*f*ztz - 2*f*ztv + vtv;
+
+                if(dist < mindist /*&& f > fmin && f < fmax*/){
+                    mindist = dist;
+                    bestf = f;
+                }
+
+                Double key = ((Double) map.firstKey());
+                int k = ((Integer)map.get(key)).intValue();
+                double d = Math.signum(z[k]);
+                v[k] += d;
+                map.remove(key);
+                map.put(new Double((d*0.5 + v[k])/z[k]), new Integer(k));
+
+                ztv += d*z[k];
+                vtv += 2*d*(v[k]-d) + 1;
+
+                //update f
+                //f = vtv/ztv;
+                f = ztv / ztz;
+            }
+        }
+        That = 1.0/bestf;
+
+        //now compute the phase estimate
+        phat = phasestor.getPhase(y, That);
+
+
+    }
+    
+}

### 2014-03-27 code/java/src/snpe/PRIEstimator.java

@@ -1,0 +1,75 @@
+package snpe;
+
+import java.io.Serializable;
+import static pubsim.Util.fracpart;
+import pubsim.lattices.Anstar.Anstar;
+import pubsim.lattices.Anstar.AnstarLinear;
+import pubsim.lattices.LatticeAndNearestPointAlgorithmInterface;
+
+
+/**
+ * Interface that describes the functions of a PRI estimator.
+ * @author Vaughan Clarkson, 15-Jun-05.
+ */
+public interface PRIEstimator extends Serializable{
+
+    /**
+     * Runs the estimator and return the estimated frequency.
+     */
+    public void estimate(Double[] y);
+
+    /**
+     * @return the period estimate
+     */
+    public double getPeriod();
+
+    /**
+     *
+     * @return the phase estimate
+     */
+    public double getPhase();
+
+    /// @return the number of observations this estimator uses.
+    public int getLength();
+    
+    /// Interface for period estimators that operate by sampling their objective functions
+    public static interface Sampled extends PRIEstimator {
+        
+        /// Run a search with grid with W
+        public void estimate(Double[] y, double W);
+
+    }
+
+    /**
+     * Given the period, you only need to solved the nearest point problem
+     * in Anstar to get the phase estimate. This class does that.
+     */
+    public static class PhaseEstimator {
+
+        protected final LatticeAndNearestPointAlgorithmInterface lattice;
+        protected final double[] z, fz;
+        protected final int N;
+
+        public PhaseEstimator(int N){
+            this.N = N;
+            lattice = new AnstarLinear(N-1);
+            z = new double[N];
+            fz = new double[N];
+        }
+
+        public double getPhase(Double[] y, double That){
+            Anstar.project(y, z);
+            for(int i = 0; i < N; i++) fz[i] =  z[i] / That;
+            lattice.nearestPoint(fz);
+            double[] s = lattice.getIndex();
+            double sumyTs = 0;
+            for(int i = 0; i < N; i++) sumyTs += y[i] - That*s[i];
+            double phase = sumyTs/N;
+            //return phase;
+            return fracpart(phase/That)*That;
+        }
+
+    }
+
+
+}

### 2014-03-27 code/java/src/snpe/PeriodogramEstimator.java

@@ -1,0 +1,109 @@
+package snpe;
+
+/**
+ * Written by Vaughan Clarkson, 05-Jan-07. New method calculatePeriodogram and
+ * fixed the Newton iteration steps, 08-Jan-07.
+ *
+ */
+public class PeriodogramEstimator extends AbstractPRIEstimator.Sampled {
+
+    public static final int MAX_ITER = 10;
+    public static final double EPSILON = 1e-10;
+    /**
+     * Period and phase estimates
+     */
+    protected double That, phat;
+    
+    protected final double[] kappa;
+
+    public PeriodogramEstimator(int N, double Tmin, double Tmax, int samples) {
+        super(N, Tmin, Tmax,samples);
+        kappa = new double[N];
+    }
+
+    /**
+     * Default number of samples is 4N
+     */
+    public PeriodogramEstimator(int N, double Tmin, double Tmax) {
+        this(N, Tmin, Tmax, 4 * N);
+    }
+
+    public static double calculatePeriodogram(Double[] y, double f) {
+        double sumur = 0, sumui = 0;
+        for (int i = 0; i < y.length; i++) {
+            sumur += Math.cos(2 * Math.PI * f * y[i]);
+            sumui += Math.sin(2 * Math.PI * f * y[i]);
+        }
+        return sumur * sumur + sumui * sumui;
+    }
+
+    @Override
+    public void estimate(Double[] y, double fstep) {
+        double maxp = 0;
+        double fhat = fmin;
+        for (double f = fmin; f <= fmax; f += fstep) {
+            double p = calculatePeriodogram(y, f);
+            if (p > maxp) {
+                maxp = p;
+                fhat = f;
+            }
+        }
+        fhat = refine(fhat, y); //refine coarse frequencey estimate
+        That = 1.0 / fhat;
+        phat = phasestor.getPhase(y, That); //now compute the phase estimate
+    }
+
+    /// Modified Newton step
+    protected double refine(double fhat, Double[] y) {
+        double maxp = calculatePeriodogram(y, fhat);
+        int numIter = 0;
+        double f = fhat;
+        double lastf = f - 2 * EPSILON;
+        double lastp = 0;
+        while (Math.abs(f - lastf) > EPSILON && numIter <= MAX_ITER && f >= fmin && f <= fmax) {
+            double sumur = 0, sumui = 0, sumvr = 0, sumvi = 0, sumwr = 0, sumwi = 0;
[…]

### 2014-03-27 code/java/src/snpe/QuantisedPeriodogramChirpZ.java

@@ -1,0 +1,72 @@
+package snpe;
+
+import java.util.Arrays;
+import pubsim.ChirpZ;
+import pubsim.Complex;
+
+/**
+ * A version of the quantised periodogram estimator that uses the Chirp z-transform
+ * rather than the FFT.  I don't expect it to be faster.
+ * @author Robby McKilliam
+ */
+public class QuantisedPeriodogramChirpZ extends PeriodogramEstimator {
+    
+    ///the quantisation level. Large is fine quantisation and small is coarse quantisation
+    public final double q;
+    
+    /** Stores sorted copy of received data. 
+     * Makes internal method signatures mimic the paper better. Mutable and not particularly safe.
+     */
+    protected final Double[] y; 
+    
+    public QuantisedPeriodogramChirpZ(int N, double Tmin, double Tmax, double q) {
+        super(N,Tmin,Tmax);
+        this.q = q;
+        if( q <= fmax - fmin ) throw new RuntimeException("q must be larger than fmax - fmin");
+        y = new Double[N];
+    }
+    
+    @Override
+    public void estimate(Double[] y) {
+        if(y.length != N) throw new ArrayIndexOutOfBoundsException("input vector y must have length " + N);
+        System.arraycopy(y, 0, this.y, 0, N); //copy to internal memory
+        Arrays.sort(this.y); //sort the recieved signal (following code assumes accending order).  This sort is not strictly neccessary.  You could instead just take the min and max.
+        
+        final double D = 1.0/2.0/(y[N-1]-y[0]); //grid width is that suggested by Ye
+        final int K = (int) Math.floor( (fmax - fmin) / D ); //number of elements in the grid
+        Complex[] V = computeV(D, K);
+        
+        //find the index that maximises D
+        int khat = 0;
+        double Vmax = V[0].abs2();
+        for(int k = 1; k < K; k++){
+            if( V[k].abs2() > Vmax ) {
+                khat = k;
+                Vmax = V[k].abs2();
+            }
+        }        
+        double ftilde = fmin + D*khat;
+        double fhat = refine(ftilde, y); //refine coarse frequency estimate
+        That = 1.0 / fhat;
+        phat = phasestor.getPhase(y, That); //now compute the phase estimate        
+    }
+
+    private Complex[] computeV(final double D, final int K) {
+        Complex A = Complex.one;
[…]

### 2014-03-27 code/java/src/snpe/QuantisedPeriodogramFFT.java

@@ -1,0 +1,102 @@
+package snpe;
+
+import java.util.Arrays;
+import pubsim.Complex;
+import static pubsim.FFT.fft;
+
+/**
+ * Class computes the periodogram estimator after quantising the observed time of arrivals.
+ * This enables use of the FFT.  Function names are notation are described in the paper
+ * 
+ * R. G. McKilliam, I. V. L. Clarkson and B. G. Quinn "Fast sparse period estimation", Feb. 2014.
+ * 
+ * @author Robby McKilliam
+ */
+public class QuantisedPeriodogramFFT extends PeriodogramEstimator {
+
+    ///the quantisation level. Large is fine quantisation and small is coarse quantisation
+    public final double q;
+    
+    /** Stores sorted copy of received data. 
+     * Makes internal method signatures mimic the paper better. Mutable and not particularly safe.
+     */
+    protected final Double[] y; 
+
+    public QuantisedPeriodogramFFT(int N, double Tmin, double Tmax, double q) {
+        super(N, Tmin, Tmax);
+        this.q = q;
+        if( q <= fmax - fmin ) throw new RuntimeException("q must be larger than fmax - fmin");
+        y = new Double[N];
+    }
+    
+    @Override
+    public void estimate(Double[] y) {
+        if(y.length != N) throw new ArrayIndexOutOfBoundsException("input vector y must have length " + N);
+        System.arraycopy(y, 0, this.y, 0, N); //copy to internal memory
+        Arrays.sort(this.y); //sort the recieved signal (following code assumes accending order).  This sort is not strictly neccessary.  You could instead just take the min and max.
+        double D = q/M(); //width of the search grid
+        Complex[] V = computeV(); //compute the vector V by FFT
+        
+        //find the index that maximises V
+        int K = (int) Math.floor( (fmax - fmin) / D );
+        int khat = 0;
+        double Vmax = V[0].abs2();
+        for(int k = 1; k < K; k++){
+            if( V[k].abs2() > Vmax ) {
+                khat = k;
+                Vmax = V[k].abs2();
+            }
+        }        
+        double ftilde = fmin + D*khat;
+        double fhat = refine(ftilde, y); //refine coarse frequency estimate
+        That = 1.0 / fhat;
+        phat = phasestor.getPhase(y, That); //now compute the phase estimate        
+    }
+    
+    /// Calculates the vector v and returns its Fourier transform computed by FFT
[…]

### 2014-03-27 code/java/src/snpe/SLS2all.java

@@ -1,0 +1,81 @@
+package snpe;
+
+/**
+ * Implementation of Sidiropoulos et al.'s SLS2-ALL algorithm for PRI
+ * estimation.
+ * @author Vaughan Clarkson, 16-Jun-05.
+ * Stupid bug fix, 17-Jun-05.
+ * Modifed Robby McKilliam 24-Feb-2014
+ */
+public class SLS2all extends AbstractPRIEstimator.Sampled {
+
+    public final int m;
+    
+    final double[] d, kappa;
+    final int[] u;
+
+    /** Period and phase estimates */
+    protected double That, phat;
+    
+    public SLS2all(int N, double Tmin, double Tmax, int samples){
+        super(N,Tmin,Tmax,samples);
+        m = N * (N-1) / 2;
+        d = new double[m];
+        u = new int[m];
+        kappa = new double[N];
+    }
+    
+    /** Default number of samples is 100 */
+    public SLS2all(int N, double Tmin, double Tmax) {
+        this(N,Tmin,Tmax,100);
+    }
+
+    @Override
+    public void estimate(Double[] y, double fstep) {
+        if (N != y.length) 
+            throw new RuntimeException("Data length " + y.length + " does not match estimator length " + N);
+
+        //first compute the period estimate
+	int k = 0;
+	for (int i = 0; i < N-1; i++)
+	    for (int j = i+1; j < N; j++)
+		d[k++] = y[j] - y[i];
+	double bestL = Double.POSITIVE_INFINITY;
+	double fhat = fmin;
+	for (double f = fmin; f <= fmax; f += fstep) {
+	    double sumu2 = 0;
+	    double sumud = 0;
+	    for (int i = 0; i < m; i++) {
+		u[i] = (int) Math.round(f * d[i]);
+		sumu2 += u[i] * u[i];
+		sumud += u[i] * d[i];
+	    }
+	    double f0 = sumu2 / sumud;
+	    double L = 0;
+	    for (int i = 0; i < m; i++) {
+		double diff = d[i] - (u[i] / f0);
+		L += diff * diff;
+	    }
+	    if (L < bestL) {
+		bestL = L;
+		fhat = f0;
+	    }
+	}
+	That = 1/fhat;
+
+        //now compute the phase estimate
+        phat = phasestor.getPhase(y, That);
+    }
+
+    @Override
+    public double getPeriod() {
+        return That;
+    }
+
+    @Override
+    public double getPhase() {
+        return phat;
+    }
+
+
+}

### 2014-03-27 code/java/src/snpe/SLS2new.java

@@ -1,0 +1,77 @@
+package snpe;
+
+
+/**
+ * A variant on the SLS2 algorithm that should do significantly better than
+ * SLS2-NOVLP.
+ * @author Vaughan Clarkson, 15-Jun-05.
+ * Added setSize method, 16-Jun-05.
+ */
+public class SLS2new extends AbstractPRIEstimator.Sampled {
+
+    public final int m, s;
+    
+    final double[] d;
+    final int[] u;
+
+    /** Period and phase estimates */
+    protected double That, phat;
+    
+    public SLS2new(int N, double Tmin, double Tmax, int samples) {
+        super(N, Tmin, Tmax, samples);
+        m = N / 2;
+        s = m + (N % 2);
+        d = new double[m];
+        u = new int[m];
+    }
+    
+    /** Default number of samples is 100 */
+    public SLS2new(int N, double Tmin, double Tmax) {
+        this(N,Tmin,Tmax,100);
+    }
+
+    @Override
+    public void estimate(Double[] y, double fstep) {
+        if (N != y.length) 
+            throw new RuntimeException("Data length " + y.length + " does not match estimator length " + N);
+
+        for (int i = 0; i < m; i++) {
+            d[i] = y[i + s] - y[i];
+        }
+        double bestL = Double.POSITIVE_INFINITY;
+        double fhat = fmin;
+        for (double f = fmin; f <= fmax; f += fstep) {
+            double sumu2 = 0;
+            double sumud = 0;
+            for (int i = 0; i < m; i++) {
+                u[i] = (int) Math.round(f * d[i]);
+                sumu2 += u[i] * u[i];
+                sumud += u[i] * d[i];
+            }
+            double f0 = sumu2 / sumud;
+            double L = 0;
+            for (int i = 0; i < m; i++) {
+                double diff = d[i] - (u[i] / f0);
+                L += diff * diff;
+            }
+            if (L < bestL) {
+                bestL = L;
+                fhat = f0;
+            }
+        }
+        That = 1 / fhat;
+
+        //now compute the phase estimate
+        phat = phasestor.getPhase(y, That);
+    }
+
+    @Override
+    public double getPeriod() {
+        return That;
+    }
+
+    @Override
+    public double getPhase() {
+        return phat;
+    }
+}

### 2014-03-27 code/java/src/snpe/SLS2novlp.java

@@ -1,0 +1,77 @@
+package snpe;
+
+/**
+ * Implementation of Sidiropoulos et al.'s SLS2-NOVLP algorithm for PRI
+ * estimation.
+ *
+ * @author Vaughan Clarkson, 15-Jun-05. Added setSize method, 16-Jun-05.
+ */
+public class SLS2novlp extends AbstractPRIEstimator.Sampled {
+
+    public final int m;
+    final double[] d;
+    final int[] u;
+    /**
+     * Period and phase estimates
+     */
+    protected double That, phat;
+
+    public SLS2novlp(int N, double Tmin, double Tmax, int samples) {
+        super(N, Tmin, Tmax, samples);
+        m = N / 2;
+        d = new double[m];
+        u = new int[m];
+    }
+
+    /**
+     * Default number of samples is 100
+     */
+    public SLS2novlp(int N, double Tmin, double Tmax) {
+        this(N, Tmin, Tmax, 100);
+    }
+
+    @Override
+    public void estimate(Double[] y, double fstep) {
+        if (N != y.length) 
+            throw new RuntimeException("Data length " + y.length + " does not match estimator length " + N);
+
+        for (int i = 0; i < m; i++) {
+            d[i] = y[(2 * i) + 1] - y[2 * i];
+        }
+        double bestL = Double.POSITIVE_INFINITY;
+        double fhat = fmin;
+        for (double f = fmin; f <= fmax; f += fstep) {
+            double sumu2 = 0;
+            double sumud = 0;
+            for (int i = 0; i < m; i++) {
+                u[i] = (int) Math.round(f * d[i]);
+                sumu2 += u[i] * u[i];
+                sumud += u[i] * d[i];
+            }
+            double f0 = sumu2 / sumud;
+            double L = 0;
+            for (int i = 0; i < m; i++) {
+                double diff = d[i] - (u[i] / f0);
+                L += diff * diff;
+            }
+            if (L < bestL) {
+                bestL = L;
+                fhat = f0;
+            }
+        }
+        That = 1 / fhat;
+
+        //now compute the phase estimate
+        phat = phasestor.getPhase(y, That);
+    }
+
+    @Override
+    public double getPeriod() {
+        return That;
+    }
+
+    @Override
+    public double getPhase() {
+        return phat;
+    }
+}

### 2014-03-27 code/java/src/snpe/SamplingLLS.java

@@ -1,0 +1,81 @@
+package snpe;
+
+import pubsim.VectorFunctions;
+import pubsim.lattices.Anstar.AnstarLinear;
+import pubsim.lattices.Anstar.AnstarVaughan;
+import pubsim.lattices.LatticeAndNearestPointAlgorithmInterface;
+
+/**
+ * A modification of the Bresenham estimator which samples the line segment
+ * rather than tracing through all the Voronoi cell boundaries.
+ *
+ * @author Vaughan Clarkson, 16-Jun-05. Add calculateObjective method,
+ * 13-Jan-07. Modified by Robby McKilliam 24-Feb-2014.
+ */
+public class SamplingLLS extends AbstractPRIEstimator.Sampled {
+
+    final double[] zeta, fzeta;
+    /**
+     * Period and phase estimates
+     */
+    protected double That, phat;
+    protected final LatticeAndNearestPointAlgorithmInterface lattice;
+    
+    public SamplingLLS(int N, double Tmin, double Tmax, int samples) {
+        super(N,Tmin,Tmax,samples);
+        lattice = new AnstarLinear(N-1);
+        zeta = new double[N];
+        fzeta = new double[N];
+    }
+
+    /** Default number of samples is 4N */
+    public SamplingLLS(int N, double Tmin, double Tmax) {
+        this(N,Tmin,Tmax,4*N);
+    }
+    
+    ///Run estimator with given search grid width
+    @Override
+    public void estimate(Double[] y, double fstep) {
+        if (N != y.length) 
+            throw new RuntimeException("Data length " + y.length + " does not match estimator length " + N);
+        
+        //first compute the period estimate
+        AnstarVaughan.project(y, zeta);
+        double ztz = VectorFunctions.sum2(zeta);
+        double bestL = Double.POSITIVE_INFINITY;
+        double fhat = fmin;
+        for (double f = fmin; f <= fmax; f += fstep) {
+            for (int i = 0; i < N; i++) {
+                fzeta[i] = f * zeta[i];
+            }
+            lattice.nearestPoint(fzeta);
+            double[] v = lattice.getLatticePoint();
+            double vtv = 0, vtz = 0;
+            for (int i = 0; i < N; i++) {
+                vtv += v[i] * v[i];
+                vtz += v[i] * zeta[i];
+            }
+            double f0 = vtv / vtz;
+            double L = ztz - 2 * vtz / f0 + vtv / (f0 * f0);
+            if (L < bestL) {
+                bestL = L;
[…]

### 2014-03-27 code/java/src/snpe/SparseNoisyPeriodicSignal.java

@@ -1,0 +1,101 @@
+/*
+ * SparseNoisyPeriodicSignal.java
+ *
+ * Created on 13 April 2007, 14:55
+ */
+
+package snpe;
+
+import pubsim.SignalGenerator;
+import pubsim.distributions.NoiseGenerator;
+
+/**
+ * Generates a set of recieved times that are sparse and
+ * noisy versions of the received signal.
+ * <p>
+ * The specific sparse pulses that will be transmitted can
+ * be specified by @func setSparseSignal.
+ * 
+ * @author Robby McKilliam
+ */
+public class SparseNoisyPeriodicSignal implements SignalGenerator<Double> {
+    
+    public final double T;
+    public final int N;
+    public final double phase;
+    
+    protected final Integer[] sparseSignal;
+    protected final Double[] recievedSignal;
+    protected final NoiseGenerator<Double> noise;
+    protected final NoiseGenerator<Integer> sparsenoise;
+    
+    public SparseNoisyPeriodicSignal(int N, double T, double phase, NoiseGenerator<Integer> sparsenoise, NoiseGenerator<Double> noise){
+        this.N = N;
+        this.T = T;
+        this.phase = phase;
+        sparseSignal = new Integer[N];
+        recievedSignal = new Double[N];
+        this.noise = noise;
+        this.sparsenoise = sparsenoise;
+        generateSparseSignal();   
+    }
+    
+    public void setSparseSignal(Integer[] S){
+        if(S.length != N) throw new ArrayIndexOutOfBoundsException("Tried to set sparse signal to the wrong length");
+        System.arraycopy(S, 0, sparseSignal, 0, N);
+    }
+    
+    public double getPeriod(){ return T; }
+    public double getPhase(){ return phase; }
+    
+    /** {@inheritDoc} */
+    @Override
+    public int getLength() {return N; }
+    
+    public Integer[] generateSparseSignal(){
+        int sum = 0;
+        for(int i = 0; i < N; i++){
+            sum += sparsenoise.getNoise();
+            sparseSignal[i] = sum;
+        }
+        //arguably this should be sorted here?
+        return sparseSignal;
+    }
+
+    /**
+     * Generate sparse noisy signal
+     */
+    @Override
+    public Double[] generateReceivedSignal() {
+          
+          for(int i = 0; i< sparseSignal.length; i++){
+              recievedSignal[i] = T * sparseSignal[i]
+                                    + noise.getNoise() + phase;
+          }
[…]

### 2014-03-27 code/java/src/snpe/Util.java

@@ -1,0 +1,34 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package snpe;
+
+import static pubsim.Util.fracpart;
+
+/**
+ * Utilities for the sparse noisy period estimators.
+ * @author Robby McKilliam
+ */
+public class Util {
+
+        /**
+         * Returns square error between estimated period That and
+         * true period T.
+         */
+        public static double periodError(double That, double T){
+            return (That - T)*(That - T);
+        }
+
+        /**
+         * Returns square error of the fraction part difference between
+         * estimated phase phat and true phase p. Requires the
+         * true period T.
+         */
+        public static double phaseError(double phat, double p, double T){
+            double fdiff = fracpart( (phat - p) / T )*T;
+            return fdiff*fdiff;
+        }
+
+}

### 2014-03-27 code/java/src/snpe/YeSampling.java

@@ -1,0 +1,53 @@
+package snpe;
+
+/**
+ * Wraps a PRIEstimator.Sampled and choose the sampling interval according to
+ * the given data. Specifically the width of the search grid is 1/(scale*(y_n-y_1)) where scale
+ * is by default 2.0, but can be set at the constructor.
+ * 
+ * @author Robby McKilliam
+ */
+public class YeSampling implements PRIEstimator {
+    
+    public final double scale;
+    
+    protected final PRIEstimator.Sampled periodest;
+    
+    public YeSampling(PRIEstimator.Sampled est){
+        periodest = est;
+        scale = 2.0;
+    }
+    
+    public YeSampling(PRIEstimator.Sampled est, double scale){
+        periodest = est;
+        this.scale = scale;
+    }
+
+    @Override
+    public void estimate(Double[] y) {
+        double Ymax = y[y.length - 1]; //supposes data is in accending order
+        double Ymin = y[0];
+        for(double v : y){ //get guaranteed minimum and maximum y
+            if(Ymax < v) Ymax = v;
+            if(Ymin > v) Ymin = v;
+        }
+        double W = 1.0 / scale / (Ymax - Ymin); //Haohuan et. al. suggested grid width
+        periodest.estimate(y, W);
+    }
+
+    @Override
+    public double getPeriod() {
+        return periodest.getPeriod();
+    }
+
+    @Override
+    public double getPhase() {
+        return periodest.getPhase();
+    }
+
+    @Override
+    public int getLength() {
+        return periodest.getLength();
+    }
+    
+}

### 2014-03-27 code/java/src/snpe/ZnLLS.java

@@ -1,0 +1,120 @@
+/*
+ * ZnLLS.java
+ *
+ * Created on 8 July 2007, 08:55
+ */
+
+package snpe;
+
+import java.util.TreeMap;
+import pubsim.lattices.Anstar.AnstarVaughan;
+
+/**
+ * Zn->An->An* line search.  Uses the sorted
+ * Zn search proposed by Dan and Vaughan.  This
+ * runs O(n^3log(n)) which is the same as the An*
+ * Bresenham line search.
+ * @author Robby McKilliam
+ */
+public class ZnLLS extends AbstractPRIEstimator {
+    
+    public final double n;
+    
+    final double[] g, z, v, fz;
+    final TreeMap map;
+
+
+    /** Period and phase estimates */
+    protected double That, phat;
+
+    public ZnLLS(int N, double Tmin, double Tmax){
+        super(N,Tmin,Tmax);
+        this.n = N-1;
+        g = new double[N];
+        z = new double[N];
+        fz = new double[N];
+        v = new double[N];
+        map = new TreeMap();
+    }
+     
+    /** 
+     * Sets protected variable g to the glue
+     * vector [i].  See SPLAG pp109.
+     */
+    protected void glueVector(double i){
+        double j = n + 1 - i;
+        for(int k = 0; k < j; k++)
+            g[k] = i/(n+1);
+        for(int k = (int)j; k < n + 1; k++)
+            g[k] = -j/(n+1);
+    }
+    
+    @Override
+    public void estimate(Double[] y){
+        if (N != y.length) 
+            throw new RuntimeException("Data length " + y.length + " does not match estimator length " + N);
+
+        AnstarVaughan.project(y, z);
+        
+        double bestf = 0.0;
+        double mindist = Double.POSITIVE_INFINITY;
+        
+        for(int i = 0; i <=n ; i++){
+            map.clear();
+            glueVector(i);
+            
+            //setup map and variables for this glue vector
+            double ztz = 0.0, ztv = 0.0, vtv = 0.0;
+            for(int j=0; j<=n; j++){
+                v[j] = Math.round(fmin*z[j] - g[j]) + g[j];
+                map.put(new Double((Math.signum(z[j])*0.5 + v[j])/z[j]), new Integer(j));
+                ztz += z[j]*z[j];
+                ztv += z[j]*v[j];
+                vtv += v[j]*v[j];
+            }
+            
+            double f = vtv/ztv;
+            //line search loop
[…]

### 2014-03-27 code/java/src/snpe/bounds/BoundCalculator.java

@@ -1,0 +1,38 @@
+package snpe.bounds;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public interface BoundCalculator {
+
+    /**
+     * Return the clairvoyant CRB for the set parameters
+     */
+    double getBound();
+
+    /**
+     *
+     * Set the parameter for the geometrically distributed
+     * index differences.
+     * @param p the geometric parameter in range [0,1]
+     */
+    void setGeometicParameter(double p);
+
+    /**
+     * Set the data length
+     */
+    void setLength(int N);
+
+    /**
+     * Set the vector of indicies
+     */
+    void setS(double[] s);
+    
+    /** Set the true period of the signal */
+    void setPeriod(double T);
+    
+    /** Set the variance of the noise */
+    void setVariance(double var);
+
+}

### 2014-03-27 code/java/src/snpe/bounds/CLT.java

@@ -1,0 +1,16 @@
+package snpe.bounds;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public interface CLT {
+    
+    public double phaseVar(int N);
+
+    public double periodVar(int N);
+
+    public double periodPhaseCoVar(int N);
+
+    
+}

### 2014-03-27 code/java/src/snpe/bounds/ClairvoyantCRB.java

@@ -1,0 +1,48 @@
+package snpe.bounds;
+
+import pubsim.lattices.Anstar.Anstar;
+import pubsim.VectorFunctions;
+
+/**
+ * This is Sirdopulous's clairvoyant Cramer-Rao bound.  It is
+ * just the bound given by linear regression under the assumption
+ * that the vector of indicies, s, is known.
+ * @author Robby McKilliam
+ */
+public class ClairvoyantCRB implements BoundCalculator {
+    
+    protected double[] s, x;
+    
+    protected double T, var, mag2x;
+    
+    /** Set the vector of indicies */
+    public void setS(double[] s){
+        this.s = s;
+        x = new double[s.length];
+        Anstar.project(s, x);
+        mag2x = VectorFunctions.sum2(x);
+    }
+    
+    /** Set the true period of the signal */
+    public void setPeriod(double T){
+        this.T = T;
+    }
+    
+    /** Set the variance of the noise */
+    public void setVariance(double var){
+        this.var = var;
+    }
+    
+    /** Return the clairvoyant CRB for the set parameters */
+    public double getBound(){
+        return var/mag2x;
+    }
+
+    public void setGeometicParameter(double p) {
+    }
+
+    public void setLength(int N) {
+    }
+    
+        
+}

### 2014-03-27 code/java/src/snpe/bounds/ClairvoyantCRBRandomS.java

@@ -1,0 +1,50 @@
+package snpe.bounds;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public class ClairvoyantCRBRandomS extends ClairvoyantCRB{
+    
+    double p;
+    int N;
+    
+    /** 
+     * Set the parameter for the geometrically distributed
+     * index differences.
+     * @param p the geometric parameter in range [0,1]
+     */
+    @Override
+    public void setGeometicParameter(double p){
+        this.p = p;
+    }
+    
+    /** Set the vector of indicies */
+    @Override
+    public void setS(double[] s){
+        setLength(s.length);
+    }
+    
+    /** Set the data length */
+    @Override
+    public void setLength(int N){
+        this.N = N-1;
+    }
+    
+    
+    /** Return the clairvoyant CRB for the set parameters */
+    @Override
+    public double getBound(){
+        double ntn = N*N*N/3.0 + N*N/2.0 + N/6.0;
+        double nt1 = N*N/2.0 + N/2.0;
+        double ep = 1/p;
+        double vp = (1-p)/(p*p);
+        double ests = ntn*ep*ep + nt1*vp;
+        double est1 = nt1*ep;
+        
+        double extx = ests - est1*est1/N;
+        
+        return var/extx;
+    }
+
+}

### 2014-03-27 code/java/src/snpe/bounds/Hamersley.java

@@ -1,0 +1,65 @@
+package snpe.bounds;
+
+import Jama.Matrix;
+import pubsim.VectorFunctions;
+
+/**
+ * Hamersley-Chapman-Robbins bound that applies
+ * when the index differences are geometrically
+ * distributed.
+ * @author Robby McKilliam
+ */
+public class Hamersley extends ClairvoyantCRB implements BoundCalculator{
+    
+    int N;
+    double p;
+    
+    /** Set the vector of indicies */
+    @Override
+    public void setS(double[] s){
+        setLength(s.length);
+    }
+    
+    /** Set the data length */
+    @Override
+    public void setLength(int N){
+        this.N = N-1;
+    }
+    
+    /** 
+     * Set the parameter for the geometrically distributed
+     * index differences.
+     * @param p the geometric parameter in range [0,1]
+     */
+    @Override
+    public void setGeometicParameter(double p){
+        this.p = p;
+    }
+    
+   /** Return the clairvoyant CRB for the set parameters */
+    @Override
+    public double getBound(){
+        double ntn = N*N*N/3.0 + N*N/2.0 + N/6.0;
+        double nt1 = N*N/2.0 + N/2.0;
+        double ep = 1/p;
+        double vp = (1-p)/(p*p);
+        double ests = ntn*ep*ep + nt1*vp;
+        double est1 = nt1*ep;
+        double estd = N*ep;
+        
+                //set the numerator matrix
+        Matrix num = new Matrix(2,2);
+        num.set(0,0, N/var); num.set(0,1, T/var);
+        num.set(1,0, T/var); num.set(1,1, Math.expm1(T*T/var));
+        
+        //set the numerator matrix
+        Matrix den = new Matrix(3,3);
+        den.set(0,0, ests/var); den.set(0,1, est1/var); den.set(0,2, T*estd/var);
+        den.set(1,0, est1/var); den.set(1,1, N/var); den.set(1,2, T/var);
+        den.set(2,1, T*estd/var); den.set(2,1, T/var); den.set(2,2, Math.expm1(T*T/var));
+        
+        return VectorFunctions.stableDet(num)/VectorFunctions.stableDet(den);
+        
+    }
+
+}

### 2014-03-27 code/java/src/snpe/bounds/HamersleyFixedS.java

@@ -1,0 +1,75 @@
+package snpe.bounds;
+
+import Jama.Matrix;
+import java.util.Iterator;
+import pubsim.lattices.util.ZnWithinSphere;
+import pubsim.VectorFunctions;
+
+/**
+ * This is the Hamerley-Chapmin-Robbins bound that
+ * includes the discrete parameter.  It is only tight when
+ * sigma^2/T^2 is less than log(3).  This is generally in
+ * the range that the estimators work so it's not a
+ * terrible restriction.
+ * @author Robby McKilliam
+ */
+public class HamersleyFixedS extends ClairvoyantCRB{
+    
+    int N;
+    double sts, st1;
+    
+    /** Set the vector of indicies */
+    @Override
+    public void setS(double[] s){
+        this.s = s;
+        sts = VectorFunctions.sum2(s);
+        st1 = VectorFunctions.sum(s);
+        N = s.length;
+        
+    }
+    
+    /** Return the Hamerley-Chapmin-Robbins bound for the set parameters */
+    @Override
+    public double getBound(){
+        
+        Matrix num = new Matrix(2,2);
+        Matrix den = new Matrix(3,3);
+        
+        System.out.println(N);
+        System.out.println(2*Math.sqrt(sts));
+        
+        ZnWithinSphere deltas = new ZnWithinSphere(N, 2*Math.sqrt(sts));
+        
+        double B = Double.NEGATIVE_INFINITY;
+        Iterator itr = deltas.iterator();
+        while(itr.hasNext()){
+            
+            double[] delta = (double[]) itr.next();
+            
+            //System.out.println(VectorFunctions.print(delta));
+            
+            
+            double dt1 = VectorFunctions.sum(delta);
+            double dts = VectorFunctions.dot(delta, s);
+            double magd = VectorFunctions.magnitude(delta);
+            double dtd = VectorFunctions.sum2(delta);
+            
+            //set the numerator matrix
+            num.set(0,0, N/var); num.set(0,1, T*dt1/(var*magd));
+            num.set(1,0, T*dt1/(var*magd)); num.set(1,1, Math.expm1(T*T*dtd/var)/dtd);
+
+            //set the numerator matrix
+            den.set(0,0, sts/var); den.set(0,1, st1/var); den.set(0,2, T*dts/(var*magd));
+            den.set(1,0, st1/var); den.set(1,1, N/var); den.set(1,2, T*dt1/(var*magd));
[…]

### 2014-03-27 code/java/src/snpe/bounds/NormalisedLLSCLT.java

@@ -1,0 +1,55 @@
+package snpe.bounds;
+
+import pubsim.distributions.RealRandomVariable;
+import pubsim.distributions.circular.CircularRandomVariable;
+
+/**
+ * Barry's remarkable central limit theorem for the LLS estimator
+ * @author Robby McKilliam
+ */
+public class NormalisedLLSCLT implements CLT{
+
+    protected final double scalefac;
+    protected final double dmean;
+
+    /**
+     * Computes all the bits needed (like the unwrapped variance etc) for
+     * the NormalisedLLSCLT.
+     * @param noise: A version of the noise distribution that is scaled by 1/T0.
+     *              This is u_n in Barry's writeup.
+     * @param discretemean: The mean of the discrete r.v. that
+     *                      described the sparseness of the signal
+     * @param T0: The `true' period of the signal
+     */
+    public NormalisedLLSCLT(RealRandomVariable noise, double discretemean, double T0){
+
+        dmean = discretemean;
+        CircularRandomVariable wrped = noise.getWrapped();
+        double h = wrped.pdf(-0.5);
+
+        //Is is the wrapped variance or not?  Need to check Barry's
+        //CLT a little more closely.
+        double wrappedvar = wrped.intrinsicVariance(0);
+        //double wrappedvar = noise.getVariance();
+
+        //this is the the scale factor contructed from these numbers;
+        scalefac = 12.0*T0*T0*wrappedvar/( (1-h)*(1-h)*dmean*dmean );
+        //System.out.println(scalefac);
+    }
+
+    @Override
+    public double phaseVar(int N){
+        return scalefac * dmean * dmean / 3.0 / N;
+    }
+
+    @Override
+    public double periodVar(int N){
+        return scalefac / N / N / N;
+    }
+
+    @Override
+    public double periodPhaseCoVar(int N){
+        return - scalefac * dmean / 2.0 / N / N;
+    }
+
+}

### 2014-03-27 code/java/src/snpe/bounds/PeriodogramCLT.java

@@ -1,0 +1,43 @@
+package snpe.bounds;
+
+import pubsim.distributions.RealRandomVariable;
+
+/**
+ * Computes the asymptotic variance of the periodogram estimator of a sparse
+ * noisy periodic signal.
+ * @author Robby McKilliam
+ */
+public class PeriodogramCLT implements CLT {
+
+    protected final double scalefac;
+
+    /* 
+     * @param noise: A version of the noise distribution that is scaled by 1/T0. This is U_n in Barry's writeup.
+     */
+    public PeriodogramCLT(RealRandomVariable noise, double discretemean, double T0) {
+
+        double numer = 3 * T0 * T0 * (1 - noise.characteristicFunction(4 * Math.PI).re());
+        double chrf2pi = noise.characteristicFunction(2*Math.PI).re();
+        double denom = 2 * Math.PI * Math.PI * discretemean * discretemean * chrf2pi * chrf2pi;
+        scalefac = numer / denom;
+    }
+
+    /** 
+     * This is wrong, but is set this way for convenience.  Should really add this later
+     */
+    @Override
+    public double phaseVar(int N) {
+        return 0.0;
+        //throw new UnsupportedOperationException("Not supported yet.");
+    }
+
+    @Override
+    public double periodVar(int N) {
+        return scalefac/N/N/N;
+    }
+
+    @Override
+    public double periodPhaseCoVar(int N) {
+        throw new UnsupportedOperationException("Not supported yet.");
+    }
+}

### 2014-03-27 code/java/test/snpe/NormalisedSamplingLLSTest.java

@@ -1,0 +1,60 @@
+package snpe;
+
+import static org.junit.Assert.assertEquals;
+import org.junit.*;
+import pubsim.distributions.GaussianNoise;
+import pubsim.distributions.discrete.PoissonRandomVariable;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public class NormalisedSamplingLLSTest {
+
+    public NormalisedSamplingLLSTest() {
+    }
+
+    @BeforeClass
+    public static void setUpClass() throws Exception {
+    }
+
+    @AfterClass
+    public static void tearDownClass() throws Exception {
+    }
+
+    @Before
+    public void setUp() {
+    }
+
+    @After
+    public void tearDown() {
+    }
+
+    /**
+     * Test of estimate method, of class NormalisedSamplingLLS.
+     */
+    @Test
+    public void testEstimate() {
+        System.out.println("NormalisedSamplingLLS estimate");
+
+        double Tmin = 0.7;
+        double Tmax = 1.3;
+        int n = 30;
+        double T = 1.1;
+        double phase = 0.4;
+        NormalisedSamplingLLS instance = new NormalisedSamplingLLS(n, Tmin, Tmax, 2*n);
+
+        double noisestd = 0.001;
+        GaussianNoise noise = new pubsim.distributions.GaussianNoise(0.0,noisestd*noisestd);
+
+        SparseNoisyPeriodicSignal sig = new SparseNoisyPeriodicSignal(n,T, phase, new PoissonRandomVariable(2),noise);
+        sig.generateSparseSignal();
+        Double[] y = sig.generateReceivedSignal();
+
+        instance.estimate(y);
+        System.out.println(instance.getPeriod() + "\t" + instance.getPhase());
+        assertEquals(T, instance.getPeriod(), 0.001);
+        assertEquals(phase, instance.getPhase(), 0.001);
+    }
+
+}

### 2014-03-27 code/java/test/snpe/PeriodogramEstimatorTest.java

@@ -1,0 +1,64 @@
+package snpe;
+
+import org.junit.After;
+import org.junit.AfterClass;
+import org.junit.Before;
+import org.junit.BeforeClass;
+import org.junit.Test;
+import static org.junit.Assert.*;
+import pubsim.distributions.GaussianNoise;
+import pubsim.distributions.discrete.PoissonRandomVariable;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public class PeriodogramEstimatorTest {
+    
+    public PeriodogramEstimatorTest() {
+    }
+    
+    @BeforeClass
+    public static void setUpClass() {
+    }
+    
+    @AfterClass
+    public static void tearDownClass() {
+    }
+    
+    @Before
+    public void setUp() {
+    }
+    
+    @After
+    public void tearDown() {
+    }
+
+    /**
+     * Test of estimate method, of class NormalisedSamplingLLS.
+     */
+    @Test
+    public void testEstimate() {
+        System.out.println("Periodogram estimator");
+
+        double Tmin = 0.7;
+        double Tmax = 1.3;
+        int n = 30;
+        double T = 1.1;
+        double phase = 0.4;
+        PeriodogramEstimator instance = new PeriodogramEstimator(n, Tmin, Tmax, 2*n);
+
+        double noisestd = 0.001;
+        GaussianNoise noise = new pubsim.distributions.GaussianNoise(0.0,noisestd*noisestd);
+
+        SparseNoisyPeriodicSignal sig = new SparseNoisyPeriodicSignal(n, T, phase, new PoissonRandomVariable(2),noise);
+        sig.generateSparseSignal();
+        Double[] y = sig.generateReceivedSignal();
+
+        instance.estimate(y);
+        System.out.println(instance.getPeriod() + "\t" + instance.getPhase());
+        assertEquals(T, instance.getPeriod(), 0.001);
+        assertEquals(phase, instance.getPhase(), 0.001);
+    }
+    
+}

### 2014-03-27 code/java/test/snpe/QuantisedPeriodogramChirpZTest.java

@@ -1,0 +1,65 @@
+package snpe;
+
+import org.junit.After;
+import org.junit.AfterClass;
+import org.junit.Before;
+import org.junit.BeforeClass;
+import org.junit.Test;
+import static org.junit.Assert.*;
+import pubsim.distributions.GaussianNoise;
+import pubsim.distributions.discrete.PoissonRandomVariable;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public class QuantisedPeriodogramChirpZTest {
+    
+    public QuantisedPeriodogramChirpZTest() {
+    }
+    
+    @BeforeClass
+    public static void setUpClass() {
+    }
+    
+    @AfterClass
+    public static void tearDownClass() {
+    }
+    
+    @Before
+    public void setUp() {
+    }
+    
+    @After
+    public void tearDown() {
+    }
+
+        /**
+     * Test of estimate method, of class Quantised Periodogram FFT.
+     */
+    @Test
+    public void testEstimate() {
+        System.out.println("Quantised Periodogram ChirpZ estimator");
+
+        double Tmin = 0.7;
+        double Tmax = 1.3;
+        int n = 100;
+        double T = 1.0;
+        double phase = 0.4;
+        QuantisedPeriodogramChirpZ instance = new QuantisedPeriodogramChirpZ(n, Tmin, Tmax, 4/Tmin);
+
+        double noisestd = 0.0001;
+        GaussianNoise noise = new pubsim.distributions.GaussianNoise(0.0,noisestd*noisestd);
+
+        SparseNoisyPeriodicSignal sig = new SparseNoisyPeriodicSignal(n, T, phase, new PoissonRandomVariable(2),noise);
+        sig.generateSparseSignal();
+        Double[] y = sig.generateReceivedSignal();
+
+        instance.estimate(y);
+        System.out.println(instance.getPeriod() + "\t" + instance.getPhase());
+        assertEquals(T, instance.getPeriod(), 0.001);
+        assertEquals(phase, instance.getPhase(), 0.001);
+    }
+    
+}
+    

### 2014-03-27 code/java/test/snpe/QuantisedPeriodogramFFTTest.java

@@ -1,0 +1,106 @@
+package snpe;
+
+import org.junit.After;
+import org.junit.AfterClass;
+import org.junit.Before;
+import org.junit.BeforeClass;
+import org.junit.Test;
+import static org.junit.Assert.*;
+import pubsim.Complex;
+import pubsim.distributions.GaussianNoise;
+import pubsim.distributions.discrete.PoissonRandomVariable;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public class QuantisedPeriodogramFFTTest {
+    
+    public QuantisedPeriodogramFFTTest() {
+    }
+    
+    @BeforeClass
+    public static void setUpClass() {
+    }
+    
+    @AfterClass
+    public static void tearDownClass() {
+    }
+    
+    @Before
+    public void setUp() {
+    }
+    
+    @After
+    public void tearDown() {
+    }
+
+    /**
+     * Test of the Iz periodgram.
+     */
+    @Test
+    public void testPeriodgramIzWithZeroData() {
+        System.out.println("test Iz periodogram with data all zeros");
+        final int N = 10;
+        final double Tmin = 0.7;
+        final double Tmax = 1.2;
+        final double q = 10;
+        final Double[] y = new Double[N];
+        for(int i = 0; i < y.length; i++) y[i] = 0.0;
+        QuantisedPeriodogramFFT instance = new QuantisedPeriodogramFFT(N,Tmin,Tmax,q);
+        instance.estimate(y);
+        for(double f = 0; f <= q; f+=0.01) assertEquals(instance.Iz(f),  N*N, 0.000001);
+    }
+    
+    /**
+     * Test of the D against B
+     */
+    @Test
+    public void testPeriodgramIzAgainstD() {
+        System.out.println("test slow Iz versus fast FFT computed D");
+        final int N = 10;
+        final double Tmin = 0.7;
+        final double Tmax = 1.2;
+        final double q = 10;
+        final Double[] y = new Double[N];
+        for(int i = 0; i < y.length; i++) y[i] = 1.0*i;
+        QuantisedPeriodogramFFT instance = new QuantisedPeriodogramFFT(N,Tmin,Tmax,q);
+        instance.estimate(y);
+        int M = instance.M(); //set M to minimum possible
+        double W = q/M; //width of the search grid
+        Complex[] D = instance.computeV(); //compute the vector D by FFT
[…]

### 2014-03-27 code/java/test/snpe/SparseNoisyPeriodicSignalTest.java

@@ -1,0 +1,60 @@
+package snpe;
+
+import junit.framework.TestCase;
+import pubsim.VectorFunctions;
+import pubsim.distributions.GaussianNoise;
+import pubsim.distributions.discrete.PoissonRandomVariable;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public class SparseNoisyPeriodicSignalTest extends TestCase {
+    
+    public SparseNoisyPeriodicSignalTest(String testName) {
+        super(testName);
+    }
+
+    /**
+     * Test of generateReceivedSignal method, of class simulator.SparseNoisyPeriodicSignal.
+     */
+    public void testGenerateTransmittedSignal() {
+        System.out.println("generateReceivedSignal");
+        
+        SparseNoisyPeriodicSignal instance = new SparseNoisyPeriodicSignal(20,1,0.0,
+                new PoissonRandomVariable(2),new GaussianNoise(0,1));
+        
+        Integer[] sig = instance.generateSparseSignal();
+        boolean result = false;
+        result = VectorFunctions.increasing(sig);
+        for(int i = 0; i < sig.length; i++){
+            result &= (((long)sig[i]) == Math.round(sig[i]));
+        }
+        assertEquals(true, result);
+        
+    }
+    
+    /**
+     * Test of generateSparseSignal method, of class simulator.SparseNoisyPeriodicSignal.
+     */
+    public void testGenerateRecievedSignal() {
+        System.out.println("generateTransmittedSignal");
+        
+        int length = 20;
+        double T = 2.0;
+        SparseNoisyPeriodicSignal instance = new SparseNoisyPeriodicSignal(length,T,0.0,
+                new PoissonRandomVariable(2),new pubsim.distributions.UniformNoise(0.0, 1.0/3.0));
+
+        Integer[] rec_sig = instance.generateSparseSignal();
+        instance.setSparseSignal(rec_sig);
+       
+        Double[] sig = instance.generateReceivedSignal();
+        boolean result;
+        for (int i = 0; i < length; i++){
+            result = (sig[i] <= (T*rec_sig[i] + 1.0)) && (sig[i] >= (T*rec_sig[i] - 1.0));
+            assertEquals(true, result);
+        }
+        
+    }
+    
+}

### 2014-03-27 code/java/test/snpe/YeSamplingTest.java

@@ -1,0 +1,92 @@
+package snpe;
+
+import org.junit.After;
+import org.junit.AfterClass;
+import org.junit.Before;
+import org.junit.BeforeClass;
+import org.junit.Test;
+import static org.junit.Assert.*;
+import pubsim.distributions.GaussianNoise;
+import pubsim.distributions.discrete.PoissonRandomVariable;
+
+/**
+ *
+ * @author harprobey
+ */
+public class YeSamplingTest {
+    
+    public YeSamplingTest() {
+    }
+    
+    @BeforeClass
+    public static void setUpClass() {
+    }
+    
+    @AfterClass
+    public static void tearDownClass() {
+    }
+    
+    @Before
+    public void setUp() {
+    }
+    
+    @After
+    public void tearDown() {
+    }
+
+        /**
+     * Test of estimate method, of class YeSampling periodogram estimator.
+     */
+    @Test
+    public void testHaohuanSamplingPeriodogramEstimate() {
+        System.out.println("HaohuanSampling Periodogram estimator");
+
+        double Tmin = 0.7;
+        double Tmax = 1.3;
+        int n = 30;
+        double T = 1.1;
+        double phase = 0.4;
+        YeSampling instance = new YeSampling(new PeriodogramEstimator(n, Tmin, Tmax));
+
+        double noisestd = 0.001;
+        GaussianNoise noise = new pubsim.distributions.GaussianNoise(0.0,noisestd*noisestd);
+
+        SparseNoisyPeriodicSignal sig = new SparseNoisyPeriodicSignal(n,T,phase,new PoissonRandomVariable(2),noise);
+        sig.generateSparseSignal();
+        Double[] y = sig.generateReceivedSignal();
+
+        instance.estimate(y);
+        System.out.println(instance.getPeriod() + "\t" + instance.getPhase());
+        assertEquals(T, instance.getPeriod(), 0.001);
+        assertEquals(phase, instance.getPhase(), 0.001);
+    }
+    
+    
+     /**
+     * Test of estimate method, of class YeSampling with normalised LLS estimator.
+     */
+    @Test
+    public void testHaohuanSamplingNormaliedLLSEstimate() {
+        System.out.println("HaohuanSampling Normalised LLS estimator");
+
+        double Tmin = 0.7;
+        double Tmax = 1.3;
+        int n = 30;
+        double T = 1.1;
+        double phase = 0.4;
[…]

### 2014-03-27 code/java/test/snpe/ZnLLSTest.java

@@ -1,0 +1,92 @@
+/*
+ * ZnLLSTest.java
+ * JUnit based test
+ *
+ * Created on 9 July 2007, 12:24
+ */
+
+package snpe;
+
+import junit.framework.TestCase;
+import pubsim.distributions.GaussianNoise;
+import pubsim.distributions.discrete.PoissonRandomVariable;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public class ZnLLSTest extends TestCase {
+    
+    public ZnLLSTest(String testName) {
+        super(testName);
+    }
+
+    /**
+     * Test of estimateFreq method, of class simulator.ZnLLS.
+     */
+    public void testEstimateFreq() {
+        System.out.println("ZnLLS estimate frequency");
+        
+        double Tmin = 0.7;
+        double Tmax = 1.3;
+        int N = 30;
+        double f = 0.9;
+        double T = 1/f;
+        ZnLLS instance = new ZnLLS(N,Tmin,Tmax);
+        
+        double noisestd = 0.0001;
+        GaussianNoise noise = new pubsim.distributions.GaussianNoise(0.0,noisestd*noisestd);
+        
+        SparseNoisyPeriodicSignal sig = new SparseNoisyPeriodicSignal(N,T,0.0,
+                new PoissonRandomVariable(2),noise);
+
+          
+        long seed = 1331;
+        noise.setSeed(seed);
+        Integer[] trans = sig.generateSparseSignal();
+        Double[] y = sig.generateReceivedSignal();
+        
+        double expResult = T;
+        instance.estimate(y);
+        assertEquals(expResult, instance.getPeriod(), 0.001);
+
+    }
+
+   /**
+     * Test of estimateFreq method, of class simulator.ZnLLS.
+     */
+    public void testEstimateWithPhase() {
+        System.out.println("ZnLLS estimate with phase");
+
+        double fmin = 0.7;
+        double fmax = 1.3;
+        int N = 30;
+        double T = 1.0;
+        double f = 1.0/T;
+        double phase = 0.15;
+        
+        ZnLLS instance = new ZnLLS(N,1/fmax,1/fmin);
+
+        double noisestd = 0.00001;
+        GaussianNoise noise = new pubsim.distributions.GaussianNoise(0.0,noisestd*noisestd);
+
+        SparseNoisyPeriodicSignal sig = new SparseNoisyPeriodicSignal(N,T,phase,
+                new PoissonRandomVariable(2),noise);
+
[…]

### 2014-03-27 code/montecarlo.scala

@@ -1,0 +1,125 @@
+/**
+ * Run simulations of various sparse noise period estimators.
+ */
+import pubsim.distributions.discrete.DiscreteRandomVariable
+import pubsim.distributions.discrete.GeometricRandomVariable
+import pubsim.distributions.GaussianNoise
+import pubsim.distributions.RealRandomVariable
+import snpe.SparseNoisyPeriodicSignal
+import snpe.PRIEstimator
+import snpe.YeSampling
+import snpe.PeriodogramEstimator
+import snpe.NormalisedSamplingLLS
+import snpe.QuantisedPeriodogramFFT
+import snpe.SLS2novlp
+import snpe.SLS2new
+import snpe.bounds.PeriodogramCLT
+import snpe.bounds.NormalisedLLSCLT
+
+val iters = 4000
+val T_0 = scala.math.Pi/3 //true period
+val T_min = 0.8 //minimum period
+val T_max = 1.5 //maximum period
+val f_min = 1.0/T_max //minimum frequency
+val f_max = 1.0/T_min //maximum frequency
+val theta_0 = 0.2 //true phase
+def sparsedist(m : Double) = new GeometricRandomVariable(1.0/m)
+def noisedist(v : Double) = new GaussianNoise(0,v) //the noise distribution we use with variance v
+
+val starttime = (new java.util.Date).getTime
+
+for( N <- List(30,1200) ) {
+  
+  //for discrete distributions with mean 3 and 15
+  for( m <- List( 3, 15 ) ) {
+    def discretedist() = sparsedist(m)
+    val varbs = -38.0 to -5.0 by 1.5 //variances we use in dB
+    runsim(varbs, N, () => discretedist, () => new YeSampling(new PeriodogramEstimator(N,T_min,T_max)), "PeriodogramN" + N + "geom" + m)
+    runsim(varbs, N, () => discretedist, () => new YeSampling(new NormalisedSamplingLLS(N,T_min,T_max)), "NormalisedSamplingLLSN" + N + "geom" + m)
+    //runsim(varbs, N, () => discretedist, () => new YeSampling(new SLS2novlp(N,T_min,T_max)), "SLS2novlpN" + N + "geom" + m)
+    runsim(varbs, N, () => discretedist, () => new YeSampling(new SLS2new(N,T_min,T_max)), "SLS2newN" + N + "geom" + m)
[…]

### 2014-03-28 paper.tex
minor
added newext build

@@
 By quantising the observations we produce an estimator that requires only $O(N\log N)$ operations by use of the chirp z-transform or the fast Fourier transform.
-The quantisation has the adverse affect of decreasing the accuracy of the estimator.
+The quantisation has the adverse effect of decreasing the accuracy of the estimator.
 This is investigated by Monte-Carlo simulation.
@@
 \]
-where $A = 1$ and $W = e^{ -2\pi j D/q}$.  All of $V_0,\dots,V_{K-1}$ can be efficiently computed by the chirp z-transform~\cite{Rabiner1969}.  Putting
+where $A = 1$ and $W = e^{ -2\pi j D/q}$.  The $V_0,\dots,V_{K-1}$ can be efficiently computed by the chirp z-transform~\cite{Rabiner1969}.  Putting
 \[
@@
       \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30geom3}; 
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN30geom3};
             \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q1.5geom3};
       \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN30q4.0geom3};
-%      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN30};
       \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN30geom3};
@@
       \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN1200geom3};
- %     \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1200};
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1200geom3};
        \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN1200geom3};
@@
  %     \legend{Periodogram Theory, Least squares theory, Quantised periodogram, Periodogram, Least squares}
-     \legend{Periodogram theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=4 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
+     \legend{Periodogram theory, Least squares theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=4 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
      \node[rotate=15,font=\footnotesize] at (axis cs:6e-4,1e-8) {$N=30$};
@@
       \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30geom15}; 
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN30geom15};
             \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q1.5geom15};
       \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN30q4.0geom15};
-%      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN30};
       \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN30geom15};
@@
       \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN1200geom15};
- %     \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1100};
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1200geom15};
        \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN1200geom15};
@@
 
-%Currently, as is common An open theoretical question is a rigorously
-
 \small
-%\bibliography{zotero}
 \bibliography{bib}

### 2014-03-28 code/java/src/snpe/bounds/NormalisedLLSCLT.java

@@ -13,7 +13,7 @@ public class NormalisedLLSCLT implements CLT{
     protected final double dmean;
 
     /**
-     * Computes all the bits needed (like the unwrapped variance etc) for
+     * Computes all the bits needed (like the intrinsic variance etc) for
      * the NormalisedLLSCLT.
      * @param noise: A version of the noise distribution that is scaled by 1/T0.
      *              This is u_n in Barry's writeup.

### 2014-03-28 code/montecarlo.scala
made geom variable start at zero

@@ -23,7 +23,7 @@ val T_max = 1.5 //maximum period
 val f_min = 1.0/T_max //minimum frequency
 val f_max = 1.0/T_min //maximum frequency
 val theta_0 = 0.2 //true phase
-def sparsedist(m : Double) = new GeometricRandomVariable(1.0/m)
+def sparsedist(m : Double) = new GeometricRandomVariable.StartingAtZero(1.0/(m+1.0))
 def noisedist(v : Double) = new GaussianNoise(0,v) //the noise distribution we use with variance v
 
 val starttime = (new java.util.Date).getTime

### 2014-03-28 code/montecarlo.scala
minor

@@ -16,7 +16,7 @@ import snpe.SLS2new
 import snpe.bounds.PeriodogramCLT
 import snpe.bounds.NormalisedLLSCLT
 
-val iters = 4000
+val iters = 3000
 val T_0 = scala.math.Pi/3 //true period
 val T_min = 0.8 //minimum period
 val T_max = 1.5 //maximum period
@@ -31,7 +31,7 @@ val starttime = (new java.util.Date).getTime
 for( N <- List(30,1200) ) {
   
   //for discrete distributions with mean 3 and 15
-  for( m <- List( 3, 15 ) ) {
+  for( m <- List( 1, 10 ) ) {
     def discretedist() = sparsedist(m)
     val varbs = -38.0 to -5.0 by 1.5 //variances we use in dB
     runsim(varbs, N, () => discretedist, () => new YeSampling(new PeriodogramEstimator(N,T_min,T_max)), "PeriodogramN" + N + "geom" + m)

### 2014-03-31 paper.tex
added more data

@@
 We also assume that $y_1,\dots,y_N$ are in ascending order.
-This assumption can be made without loss of generality since it amounts only to sorting the observations if required.
+The periodogram estimator is unchanged by a permutation observations, and so, this assumption can be made without loss of generality because it amounts only to sorting the observations if required.
 
@@
 \]
-where $A = 1$ and $W = e^{ -2\pi j D/q}$.  The $V_0,\dots,V_{K-1}$ can be efficiently computed by the chirp z-transform~\cite{Rabiner1969}.  Putting
+where $A = 1$ and $W = e^{ -2\pi j D/q}$.  Now $V_0,\dots,V_{K-1}$ can be efficiently computed by the chirp z-transform~\cite{Rabiner1969}.  Putting
 \[
@@
     \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.19,0.28)}},xlabel style={at={(0.5,0.14)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=6, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
-      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30geom3}; 
-      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN30geom3};
-            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q1.5geom3};
-      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN30q4.0geom3};
-      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN30geom3};
-       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/NormalisedSamplingLLSN30geom3}; 
- %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN30geom3};
-      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN30geom3};
+      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30geom1}; 
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN30geom1};
+            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q1.5geom1};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN30q5.0geom1};
+      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN30geom1};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/NormalisedSamplingLLSN30geom1}; 
+ %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN30geom1};
+      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN30geom1};
 %      \addplot[mark=none,dashed,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q2};
@@
 %      \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/NormalisedSamplingLLSN200};
-      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN1200geom3};
-      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1200geom3};
-       \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN1200geom3};
-       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/NormalisedSamplingLLSN1200geom3};
- %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1200geom3};
-       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN1200geom3};
-       \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q1.5geom3};
+      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN1200geom1};
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1200geom1};
+       \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN1200geom1};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/NormalisedSamplingLLSN1200geom1};
+ %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1200geom1};
+       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN1200geom1};
+       \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q1.5geom1};
       %\addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q3};
-      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q4.0geom3};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q5.0geom1};
  %     \legend{Periodogram Theory, Least squares theory, Quantised periodogram, Periodogram, Least squares}
@@
  \begin{axis}[name=plot2,at={(11,0)},font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.19,0.28)}},xlabel style={at={(0.5,0.14)}}]
-      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30geom15}; 
-      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN30geom15};
-            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q1.5geom15};
-      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN30q4.0geom15};
-      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN30geom15};
-       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/NormalisedSamplingLLSN30geom15}; 
-  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN30geom15};
-       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN30geom15};
+      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30geom10}; 
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN30geom10};
+            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q1.5geom10};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN30q5.0geom10};
+      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN30geom10};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/NormalisedSamplingLLSN30geom10}; 
+  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN30geom10};
+       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN30geom10};
 %      \addplot[mark=none,dashed,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q2};
@@
 %      \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/NormalisedSamplingLLSN200}; 
-      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN1200geom15};
-      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1200geom15};
-       \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN1200geom15};
-       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/NormalisedSamplingLLSN1200geom15};
- %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1100geom15};
-       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN1200geom15};
-       \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q1.5geom15};
+      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN1200geom10};
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1200geom10};
+       \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN1200geom10};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/NormalisedSamplingLLSN1200geom10};
+ %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1100geom10};
+       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN1200geom10};
+       \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q1.5geom10};
       %\addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1100q3};
-      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q4.0geom15};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q4.0geom10};
  %     \legend{Periodogram Theory, Least squares theory, Quantised periodogram, Periodogram, Least squares}
@@
   \caption{Mean square period error versus noise variance $\sigma^2$.
-The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=2,\dots,N$ are independent and identically geometrically distributed with mean $\mu = 3$ in the plot on the left and mean $\mu=15$ in the plot on the right.}\label{plot:multipleN}
+The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=2,\dots,N$ are independent and identically geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right.}\label{plot:multipleN}
 \end{figure*}
@@
 In all simulations the noise variables $w_1,\dots,w_N$ are independent and identically normally distributed with variance $\sigma^2$.
-The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=2,\dots,N$ are independent and identically geometrically distributed starting from $1$ with means $\mu=3$ and $15$.
+The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=2,\dots,N$ are independent and identically geometrically distributed starting from zero with means $\mu=1$ and $10$.
 The true period is $T_0 = \pi/3$, the true phase is $\theta_0 = 0.2$, and $T_{\text{min}} = 0.8$ and $T_{\text{max}} = 1.5$.
@@
 Figure~\ref{plot:multipleN} shows the mean square error of the estimators versus $\sigma^2$ for $N = 30$ and $N = 1200$.
-In the left hand  plot $\mu=3$ and in the right hand plot $\mu=15$.
-Thus, on average $\tfrac{1}{3}$ of the elements of the point process~\eqref{eq:periodicprocess} are observed in the left hand  plot and only $\tfrac{1}{15}$ in the right hand plot.
+In the left hand  plot $\mu=1$ and in the right hand plot $\mu=10$.
+%Thus, on average $\tfrac{1}{3}$ of the elements of the point process~\eqref{eq:periodicprocess} are observed in the left hand  plot and only $\tfrac{1}{15}$ in the right hand plot.
 Four thousand Monte-Carlo trials are run for each value of $\sigma^2$.
@@
 
-Figure~\ref{plot:benchmark} shows the average running time for the estimators for $N$ in the range $10$ to $25000$ when $\mu=3$.
-The running times when $\mu=15$ lead to similar conclusions and are omitted.
+Figure~\ref{plot:benchmark} shows the average running time for the estimators for $N$ in the range $10$ to $25000$ when $\mu=1$.
+The running times when $\mu=10$ lead to similar conclusions and are omitted.
 The quantisation parameter is $q = 4 f_{\text{max}}$ so that the performance of the quantised periodogram estimator is similar to that of the original periodogram estimator as observed in Figure~\ref{plot:multipleN}.

### 2014-03-31 code/montecarlo.scala

@@ -34,11 +34,11 @@ for( N <- List(30,1200) ) {
   for( m <- List( 1, 10 ) ) {
     def discretedist() = sparsedist(m)
     val varbs = -38.0 to -5.0 by 1.5 //variances we use in dB
-    runsim(varbs, N, () => discretedist, () => new YeSampling(new PeriodogramEstimator(N,T_min,T_max)), "PeriodogramN" + N + "geom" + m)
-    runsim(varbs, N, () => discretedist, () => new YeSampling(new NormalisedSamplingLLS(N,T_min,T_max)), "NormalisedSamplingLLSN" + N + "geom" + m)
+    //runsim(varbs, N, () => discretedist, () => new YeSampling(new PeriodogramEstimator(N,T_min,T_max)), "PeriodogramN" + N + "geom" + m)
+    //runsim(varbs, N, () => discretedist, () => new YeSampling(new NormalisedSamplingLLS(N,T_min,T_max)), "NormalisedSamplingLLSN" + N + "geom" + m)
     //runsim(varbs, N, () => discretedist, () => new YeSampling(new SLS2novlp(N,T_min,T_max)), "SLS2novlpN" + N + "geom" + m)
-    runsim(varbs, N, () => discretedist, () => new YeSampling(new SLS2new(N,T_min,T_max)), "SLS2newN" + N + "geom" + m)
-    for( q <- List(1.5, 4.0) ) runsim(varbs, N, () => discretedist, () => new QuantisedPeriodogramFFT(N,T_min,T_max,q*f_max), "QuantisedPeriodogramFFTN" + N + "q" + q + "geom" + m)
+    //runsim(varbs, N, () => discretedist, () => new YeSampling(new SLS2new(N,T_min,T_max)), "SLS2newN" + N + "geom" + m)
+    for( q <- List(1.5,2.0,4.0,5.0) ) runsim(varbs, N, () => discretedist, () => new QuantisedPeriodogramFFT(N,T_min,T_max,q*f_max), "QuantisedPeriodogramFFTN" + N + "q" + q + "geom" + m)
 
     //now compute CLTs
     val vardbsfiner = -38.0 to -5.0 by 0.2

### 2014-03-31 paper.tex
minor data

@@
     \selectcolormodel{gray} 
-    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.19,0.28)}},xlabel style={at={(0.5,0.14)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=6, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
+    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.19,0.28)}},xlabel style={at={(0.5,0.14)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=7, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
       \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30geom1}; 
@@
       %\addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1100q3};
-      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q4.0geom10};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q5.0geom10};
  %     \legend{Periodogram Theory, Least squares theory, Quantised periodogram, Periodogram, Least squares}

### 2014-03-31 paper.tex
added to paper.

@@
 
-The problem of estimating $T_0$ is called the \emph{sparse, noisy period estimation problem}~\cite{Clarkson2007,McKilliam2007}.
-Solutions have applications to synchronisation for telecommunications devices~\cite{Fogel1988,Fogel1989_bit_synch_zero_crossings,Sidiropoulos2005,5621928}, estimation of the pulse repetition interval in electronic support~\cite{EltonGray_puilse_train_rader_1994,Gray_more_pri_1994,Clarkson_thesis,clarkson_estimate_period_pulse_train_1996,Hauochan_pri_2012}, and beat and tempo estimation in music~\cite{dixon_beat_extraction_2001}.
-The problem is closely related to spike interval estimation in neurology~\cite{Arnett_neuro_pri_1976,Brillinger_spike_trains_1988,Rossoni200630} and a similar model has been proposed for the detection of anomolies, such as denial-of-service attacks, in internet traffic~\cite{He_detecting_periodic_patterns_in_internet_2006,5585849,5947313}.
-The problem is also connected with classical problems in number theory such as Diophantine approximation, lattice reduction, and the study of prime numbers~\cite{Cassels_geom_numbers_1997,490557,Clarkson_thesis,Lenstra1982,Wubben_2011,Casey1995,726812,CaseySadler_primes_2013}.
-
-Least squares estimation of period has been studied in~\cite{Clarkson2007,McKilliam2007,Quinn_sparse_noisy_SSP_2012}.
-The least squares estimator can be closely approximated using $O(N^2)$ operations by using special purpose algorithms from computational number theory~\cite{McKilliam2009CoxeterLattices,McKilliam2008,McKilliam2008b}.
-The least squares estimator is also the maximum likelihood estimator in the case that the noise variables $w_1,\dots,w_N$ are independent and identically normally distributed.
-An alternative estimator, called the~\emph{periodogram estimator}, was studied by Fogel and Gavish~\cite{Fogel1988,Fogel1989_bit_synch_zero_crossings} and involves the maximisation of Bartlett's periodogram for point processes~\cite{Bartlest_periodgram_point_process_1963}.
-The statistical properties of the periodogram estimator were recently studied by Quinn~et.~al.~\cite{Quinn20013asilomar_period_est} who show the estimator to be strongly consistent and asymptotically normal under mild conditions on the noise $w_1,\dots,w_N$ and integers $s_1,\dots,s_N$.
-The fastest existing algorithms can closely approximate the periodogram estimator in $O(N^2)$ operations.
-We describe the periodogram estimator in Section~\ref{sec:peri-estim}.
-
-In Section~\ref{sec:quant-peri} we show that, by quantising the observations $y_1,\dots,y_N$, an approximation to the periodogram estimator can be computed using only $O(N\log N)$ operations.
-This is facilitated by either the chirp z-transform~\cite{Rabiner1969} or the fast Fourier transform.
-The quantisation has the adverse affect of decreasing the accuracy of the estimator.
-The estimator is accurate if a sufficiently fine quantiser is chosen, but this requires a larger transform.
-The trade off between quantisation level, accuracy, and computational complexity is investigated by computer simulation in Section~\ref{sec:simulations}.
-The simulations indicate that significant computational savings are possible with negligible loss in accuracy.
-Sadler and Casey~\cite{726812} have previously suggested use of the fast Fourier transform for period estimation.
-However, their estimator requires far larger transforms and produces less accurate results than the estimator described here (see Section~VI and Figures 1-4 in~\cite{726812}).
-In contrast, our estimator has accuracy similar to the least squares estimator, i.e., similar to the maximum likelihood estimator if the noise variables $w_1,\dots,w_N$ are assumed independent and identically normally distributed.
-
 {
@@
 
+The problem of estimating $T_0$ is called the \emph{sparse, noisy period estimation problem}~\cite{Clarkson2007,McKilliam2007}.
+Solutions have applications to synchronisation for telecommunications devices~\cite{Fogel1988,Fogel1989_bit_synch_zero_crossings,Sidiropoulos2005,5621928}, estimation of the pulse repetition interval in electronic support~\cite{EltonGray_puilse_train_rader_1994,Gray_more_pri_1994,Clarkson_thesis,clarkson_estimate_period_pulse_train_1996,Hauochan_pri_2012}, and beat and tempo estimation in music~\cite{dixon_beat_extraction_2001}.
+The problem is closely related to spike interval estimation in neurology~\cite{Arnett_neuro_pri_1976,Brillinger_spike_trains_1988,Rossoni200630} and a similar model has been proposed for the detection of anomolies, such as denial-of-service attacks, in internet traffic~\cite{He_detecting_periodic_patterns_in_internet_2006,5585849,5947313}.
+The problem is also connected with classical problems in number theory such as Diophantine approximation, lattice reduction, and the study of prime numbers~\cite{Cassels_geom_numbers_1997,490557,Clarkson_thesis,Lenstra1982,Wubben_2011,Casey1995,726812,CaseySadler_primes_2013}.
+
+Least squares estimation of period has been studied in~\cite{Clarkson2007,McKilliam2007,Quinn_sparse_noisy_SSP_2012}.
+The least squares estimator can be closely approximated using $O(N^2)$ operations by using special purpose algorithms from computational number theory~\cite{McKilliam2009CoxeterLattices,McKilliam2008,McKilliam2008b}.
+The least squares estimator is also the maximum likelihood estimator in the case that the noise variables $w_1,\dots,w_N$ are independent and identically normally distributed.
+An alternative estimator, called the~\emph{periodogram estimator}, was studied by Fogel and Gavish~\cite{Fogel1988,Fogel1989_bit_synch_zero_crossings} and involves the maximisation of Bartlett's periodogram for point processes~\cite{Bartlest_periodgram_point_process_1963}.
+The statistical properties of the periodogram estimator were recently studied by Quinn~et.~al.~\cite{Quinn20013asilomar_period_est} who show the estimator to be strongly consistent and asymptotically normal under mild conditions on the noise $w_1,\dots,w_N$ and integers $s_1,\dots,s_N$.
+The fastest existing algorithms can closely approximate the periodogram estimator in $O(N^2)$ operations.
+We describe the periodogram estimator in Section~\ref{sec:peri-estim}.
+
+In Section~\ref{sec:quant-peri} we show that, by quantising the observations $y_1,\dots,y_N$, an approximation to the periodogram estimator can be computed using only $O(N\log N)$ operations.
+This is facilitated by either the chirp z-transform~\cite{Rabiner1969} or the fast Fourier transform.
+The quantisation has the adverse affect of decreasing the accuracy of the estimator.
+The estimator is accurate if a sufficiently fine quantiser is chosen, but this requires a larger transform.
+The trade off between quantisation level, accuracy, and computational complexity is investigated by computer simulation in Section~\ref{sec:simulations}.
+The simulations indicate that significant computational savings are possible with negligible loss in accuracy.
+Sadler and Casey~\cite{726812} have previously suggested use of the fast Fourier transform for period estimation.
+However, their estimator requires far larger transforms and produces less accurate results than the estimator described here (see Section~VI and Figures 1-4 in~\cite{726812}).
+In contrast, our estimator has accuracy similar to the least squares estimator, i.e., similar to the maximum likelihood estimator if the noise variables $w_1,\dots,w_N$ are assumed independent and identically normally distributed.
+
 \section{The periodogram estimator}\label{sec:peri-estim}
 
-Before describing the periodogram estimator we state some required assumptions.  Observe that
+%Before describing the periodogram estimator we state some required assumptions.  
+Observe that
 \[
@@
 where $r_n = 2s_n \in \ints$.
-Thus, to ensure identifiability, it is necessary to assume that $T_0$ lies in some interval $(T_{\text{min}}, T_{\text{max}})$ where $T_{\text{max}} < 2T_{\text{min}}$.
-We also assume that $y_1,\dots,y_N$ are in ascending order.
+Thus, to ensure identifiability, it is necessary to assume that $T_0$ lies in some interval $(T_{\text{min}}, T_{\text{max}})$ where $T_{\text{max}} < 2T_{\text{min}}$.%  We also assume that $y_1,\dots,y_N$ are in ascending order.
 The periodogram estimator is unchanged by a permutation observations, and so, this assumption can be made without loss of generality because it amounts only to sorting the observations if required.
@@
 
+\newcommand{\ymax}{y_{\text{max}}}
+\newcommand{\ymin}{y_{\text{min}}}
+
 In~\cite{McKilliam2007} it was observed that $D$ must shrink as $N$ increases in order for $\widetilde{f}$ to be a sufficiently good approximation to $\hat{f}$.
@@
 \begin{equation}\label{eq:gridwidthrecommended}
-D = \frac{1}{2(y_N - y_1)}
+D = \frac{1}{2(\ymax - \ymin)}
 \end{equation}
-which is $O(N^{-1})$ under reasonable assumptions about the integers $s_1,\dots,s_N$ and the noise $w_1,\dots,w_N$.
+where $\ymax$ and $\ymin$ are the largest and smallest values from $y_1,\dots,y_N$.
 This choice is supported by the simulations in~\cite{Haohuan2013435,6492742} and also those in Section~\ref{sec:simulations}.
+With this choice $D = O(N^{-1})$ under reasonable assumptions about the integers $s_1,\dots,s_N$ and the noise $w_1,\dots,w_N$.
 The asymptotic standard deviation of $\hat{f}$ is of order $O(N^{-3/2})$ as shown in~\cite{Quinn20013asilomar_period_est} and this initially suggests that $D = O(N^{-1})$ is too large.
@@
 
+\newcommand{\ellmin}{\ell_{\text{min}}}
+\newcommand{\ellmax}{\ell_{\text{max}}}
+
 The maximiser $\widetilde{f}^\prime$ can be computed efficiently using the chirp z-transform.  Let $b_i$ be the number of integers from $\ell_1,\dots,\ell_N$ that are equal to $i$.  That is,
@@
 \]
-where $\#$ indicates the number of elements in a set.
-The integers $\ell_1,\dots,\ell_N$ are in ascending order because $y_1,\dots,y_N$ are in ascending order.
-So $b_i = 0$ if $i < \ell_1$ or $i > \ell_N$.
-Now
+where $\#$ indicates the number of elements in a set.  Put $\ellmax = \round{q \ymax}$ and $\ellmin = \round{q \ymin}$ so that $b_i = 0$ if $i < \ellmin$ or $i > \ellmax$.  Now
 \[
-I_z(f) = \babs{ \sum_{i = \ell_1}^{\ell_N} e^{ -2\pi j f i/q} b_i }^2 = \babs{ \sum_{m = 0}^{\ell_N-\ell_1} e^{ -2\pi j f m/q} b_{m+\ell_1} }^2.
+I_z(f) = \babs{ \sum_{i = \ellmin}^{\ellmax} e^{ -2\pi j f i/q} b_i }^2 = \babs{ \sum_{m = 0}^{\ellmax-\ellmin} e^{ -2\pi j f m/q} b_{m+\ell_1} }^2.
 \]
@@
 \[
-I_z(f_{\text{min}} + Dk) = \babs{ \sum_{m = 0}^{\ell_N-\ell_1} e^{ -2\pi j Dk m/q} v_m }^2 = \abs{V_k}^2
+I_z(f_{\text{min}} + Dk) = \babs{ \sum_{m = 0}^{\ellmax-\ellmin} e^{ -2\pi j Dk m/q} v_m }^2 = \abs{V_k}^2
 \]   
@@
 \[
-V_k = \sum_{m = 0}^{\ell_N-\ell_1} e^{ -2\pi j Dk m/q} v_m = \sum_{m=0}^{\ell_N-\ell_1} A^{-m} W^{km} v_m
+V_k = \sum_{m = 0}^{\ellmax-\ellmin} e^{ -2\pi j Dk m/q} v_m = \sum_{m=0}^{\ellmax-\ellmin} A^{-m} W^{km} v_m
 \]
@@
 \[
-M = \frac{q}{D} > \ell_N-\ell_1.
+M = \frac{q}{D} > \ellmax-\ellmin.
 \]
@@
 We now describe recommended values for the parameters $q, M$, and $D$.
-The simulations presented in Section~\ref{sec:simulations} suggest that a reasonable value for the quantisation parameter is $q = 4 f_{\text{max}}$.
+The simulations presented in Section~\ref{sec:simulations} suggest that a reasonable value for the quantisation parameter is $q = 5 f_{\text{max}}$.
 If the chirp z-transform is used then $D$ is chosen according to~\eqref{eq:gridwidthrecommended}.
@@
 \[
-L = \ell_N - \ell_1 + 1 = \round{qy_N} - \round{qy_1} + 1.
+L = \ellmax - \ellmin + 1 = \round{q\ymax} - \round{q\ymin} + 1.
 \]
@@
 \[ 
-D = \frac{q}{M} = \frac{q}{2(\round{qy_N} - \round{qy_1} + 1)} \approx  \frac{1}{2(y_N - y_1)}
+D = \frac{q}{M} = \frac{q}{2(\round{q\ymax} - \round{q\ymax} + 1)} \approx  \frac{1}{2(\ymax - \ymin)}
 \]
@@
 
-Observe that $L$ grows proportionally with $q$ and, under reasonable assumptions about the integers $s_1,\dots,s_N$ and the noise variables $w_1,\dots,w_N$, we have $L = O(qN)$.
+Observe that $L$ grows proportionally with $q$ and that $L = O(qN)$ under reasonable assumptions about the integers $s_1,\dots,s_N$ and the noise variables $w_1,\dots,w_N$.
 The most computationally expensive part of the chirp z-transform is the computation of three fast Fourier transforms of length $L+K-1$.
@@
  %     \legend{Periodogram Theory, Least squares theory, Quantised periodogram, Periodogram, Least squares}
-     \legend{Periodogram theory, Least squares theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=4 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
-     \node[rotate=15,font=\footnotesize] at (axis cs:6e-4,1e-8) {$N=30$};
-     \node[rotate=15,font=\footnotesize] at (axis cs:7e-4,2e-13) {$N=1200$}; 
+     \legend{Periodogram theory, Least squares theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
+     \node[rotate=17,font=\footnotesize] at (axis cs:6e-4,9e-8) {$N=30$};
+     \node[rotate=17,font=\footnotesize] at (axis cs:7e-4,1.5e-12) {$N=1200$}; 
    \end{axis} 
@@
      %\legend{Periodogram theory, $q=\tfrac{3}{2}T_{\text{max}}$, $q=4T_{\text{max}}$, Periodogram, Least squares, SLS2-adj}
-     \node[rotate=15,font=\footnotesize] at (axis cs:6e-4,4e-10) {$N=30$};
-     \node[rotate=15,font=\footnotesize] at (axis cs:7e-4,0.6e-14) {$N=1200$}; 
+     \node[rotate=16,font=\footnotesize] at (axis cs:6e-4,9e-10) {$N=30$};
+     \node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=1200$}; 
    \end{axis}
@@
 The grid distance $D$ used for the modified least squares, periodogram, and SLS2-new estimators is given by~\eqref{eq:gridwidthrecommended}.
-The quantised periodogram estimator is computed using a single fast Fourier transform of length $M = 2L = 2(\ell_N - \ell_N + 1)$.
+The quantised periodogram estimator is computed using a single fast Fourier transform of length $M = 2L = 2(\ellmax - \ellmin + 1)$.
 Both the periodogram and quanitised periodogram estimators use the Newton-Raphson method.
@@
 Four thousand Monte-Carlo trials are run for each value of $\sigma^2$.
-The plots show the mean square error of the quantised periodogram estimator for two quantisation parameters $q = \tfrac{3}{2}f_{\text{max}} = \tfrac{15}{8}$ and $q = 4 f_{\text{max}} = 5$.
-The estimator performs poorly when $q = \tfrac{3}{2}f_{\text{max}}$ (triangles) but performs almost identically to the original periodogram estimator (circles) when $q = 4 f_{\text{max}}$ (dots).
+The plots show the mean square error of the quantised periodogram estimator for two quantisation parameters $q = \tfrac{3}{2}f_{\text{max}} = \tfrac{15}{8}$ and $q = 5 f_{\text{max}} = \tfrac{25}{4}$.
+The estimator performs poorly when $q = \tfrac{3}{2}f_{\text{max}}$ (triangles) but performs almost identically to the original periodogram estimator (circles) when $q = 5 f_{\text{max}}$ (dots).
 For comparison purposes, the plot also shows the mean square error of the modified least squares estimator~\cite{Clarkson2007,McKilliam2007} and the SLS2-new estimator~\cite{Clarkson2007,Sidiropoulos2005}.
@@
 
-Figure~\ref{plot:benchmark} shows the average running time for the estimators for $N$ in the range $10$ to $25000$ when $\mu=1$.
-The running times when $\mu=10$ lead to similar conclusions and are omitted.
-The quantisation parameter is $q = 4 f_{\text{max}}$ so that the performance of the quantised periodogram estimator is similar to that of the original periodogram estimator as observed in Figure~\ref{plot:multipleN}.
+Figure~\ref{plot:benchmark} shows the average running time for the estimators for $N$ in the range $10$ to $25000$ when $\mu=3$.
+Running times for other values of $\mu$ lead to similar conclusions.
+The quantisation parameter is $q = 5 f_{\text{max}}$ so that the performance of the quantised periodogram estimator is similar to that of the original periodogram estimator as observed in Figure~\ref{plot:multipleN}.
 The average running times for the quantised periodogram estimator computed using the chirp z-transform and using a single fast Fourier transform are shown.

### 2014-03-31 code/montecarlo.scala

@@ -35,10 +35,10 @@ for( N <- List(30,1200) ) {
     def discretedist() = sparsedist(m)
     val varbs = -38.0 to -5.0 by 1.5 //variances we use in dB
     //runsim(varbs, N, () => discretedist, () => new YeSampling(new PeriodogramEstimator(N,T_min,T_max)), "PeriodogramN" + N + "geom" + m)
-    //runsim(varbs, N, () => discretedist, () => new YeSampling(new NormalisedSamplingLLS(N,T_min,T_max)), "NormalisedSamplingLLSN" + N + "geom" + m)
+    runsim(varbs, N, () => discretedist, () => new YeSampling(new NormalisedSamplingLLS(N,T_min,T_max),10.0), "NormalisedSamplingLLSNsamp10" + N + "geom" + m)
     //runsim(varbs, N, () => discretedist, () => new YeSampling(new SLS2novlp(N,T_min,T_max)), "SLS2novlpN" + N + "geom" + m)
     //runsim(varbs, N, () => discretedist, () => new YeSampling(new SLS2new(N,T_min,T_max)), "SLS2newN" + N + "geom" + m)
-    for( q <- List(1.5,2.0,4.0,5.0) ) runsim(varbs, N, () => discretedist, () => new QuantisedPeriodogramFFT(N,T_min,T_max,q*f_max), "QuantisedPeriodogramFFTN" + N + "q" + q + "geom" + m)
+    //for( q <- List(1.5,2.0,4.0,5.0) ) runsim(varbs, N, () => discretedist, () => new QuantisedPeriodogramFFT(N,T_min,T_max,q*f_max), "QuantisedPeriodogramFFTN" + N + "q" + q + "geom" + m)
 
     //now compute CLTs
     val vardbsfiner = -38.0 to -5.0 by 0.2

### 2014-03-31 paper.tex
minor

@@
 \[
-V_k = \sum_{m = 0}^{\ellmax-\ellmin} e^{ -2\pi j Dk m/q} v_m = \sum_{m=0}^{\ellmax-\ellmin} A^{-m} W^{km} v_m
+V_k = \sum_{m = 0}^{\ellmax-\ellmin} e^{ -2\pi j Dk m/q} v_m = \sum_{m=0}^{\ellmax-\ellmin} A^{-m} W^{km} v_m,
 \]
-where $A = 1$ and $W = e^{ -2\pi j D/q}$.  Now $V_0,\dots,V_{K-1}$ can be efficiently computed by the chirp z-transform~\cite{Rabiner1969}.  Putting
+$A = 1$, $W = e^{ -2\pi j D/q}$ and $V_0,\dots,V_{K-1}$ can be efficiently computed by the chirp z-transform~\cite{Rabiner1969}.  Putting
 \[
@@
     \selectcolormodel{gray} 
-    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.19,0.28)}},xlabel style={at={(0.5,0.14)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=7, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
+    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.19,0.24)}},xlabel style={at={(0.5,0.14)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=7, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
       \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30geom1}; 
@@
    \end{axis} 
- \begin{axis}[name=plot2,at={(11,0)},font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.19,0.28)}},xlabel style={at={(0.5,0.14)}}]
+ \begin{axis}[name=plot2,at={(11,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.19,0.23)}},xlabel style={at={(0.5,0.14)}}]
       \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30geom10}; 
@@
 
-This section presents the results of Monte-Carlo simulations with the periodogram estimator~\cite{Haohuan2013435,Fogel1989_bit_synch_zero_crossings}, the modified least squares estimator~\cite{Clarkson2007,McKilliam2007}, Clarkson's modification of the separable least squares line search (SLS2-new) estimator~\cite{Clarkson2007,Sidiropoulos2005}, and the quantised periodogram estimator from Section~\ref{sec:quant-peri}.
+This section presents the results of Monte-Carlo simulations with the periodogram estimator~\cite{Haohuan2013435,Fogel1989_bit_synch_zero_crossings}, the normalised least squares estimator~\cite{Clarkson2007,McKilliam2007,Quinn_sparse_noisy_SSP_2012}, Clarkson's modification of the separable least squares line search (SLS2-new) estimator~\cite{Clarkson2007,Sidiropoulos2005}, and the quantised periodogram estimator from Section~\ref{sec:quant-peri}.
 The grid distance $D$ used for the modified least squares, periodogram, and SLS2-new estimators is given by~\eqref{eq:gridwidthrecommended}.
@@
 %Thus, on average $\tfrac{1}{3}$ of the elements of the point process~\eqref{eq:periodicprocess} are observed in the left hand  plot and only $\tfrac{1}{15}$ in the right hand plot.
-Four thousand Monte-Carlo trials are run for each value of $\sigma^2$.
+Three thousand Monte-Carlo trials are run for each value of $\sigma^2$.
 The plots show the mean square error of the quantised periodogram estimator for two quantisation parameters $q = \tfrac{3}{2}f_{\text{max}} = \tfrac{15}{8}$ and $q = 5 f_{\text{max}} = \tfrac{25}{4}$.
@@
 %For small $\sigma^2$ the mean square error of the SLS2 estimator is almost three orders of magnitude less accurate when $N = 50$ and almost five orders of magnitude less accurate when $N=1000$.
-The solid line shows the mean square error of the periodogram estimator predicted by Theorem~2 from~\cite{Quinn20013asilomar_period_est}.
+The solid line shows the mean square error of the periodogram estimator predicted by Theorem~2 from~\cite{Quinn20013asilomar_period_est} and the dashed line the mean square error of the modified least squares estimator predicted by Theorem~3 from~\cite{Quinn_sparse_noisy_SSP_2012}.
 
@@
 The quantised periodogram estimator is significantly faster than the other estimator for large $N$.
-For example, when $N = 25000$ the periodogram estimator requires approximately 530 seconds, whereas the quantised periodogram estimator requires approximately 960 milliseconds.
+For example, when $N = 25000$ the periodogram estimator requires approximately 530 seconds, whereas the quantised periodogram estimator requires approximately 990 milliseconds.
 The software is written in Java and the computer used is an Intel Core2 Q6600 running at \unit[2.4]{GHz}.

### 2014-03-31 code/benchmark.scala

@@ -37,8 +37,8 @@ val Ns = List(10,14,20,30,40,50,60,80,100,120,150,200,250,300,400,600,900,1200,1
 //runbench(Ns, (N : Int) => new YeSampling(new NormalisedSamplingLLS(N,T_min,T_max)), "LeastSquares")
 //runbench(Ns, (N : Int) => new YeSampling(new SLS2novlp(N,T_min,T_max)), "SLS2novlp")
 //runbench(Ns, (N : Int) => new YeSampling(new SLS2new(N,T_min,T_max)), "SLS2new")
-for( q <- List(4.0) ) runbench(Ns, (N : Int) => new QuantisedPeriodogramFFT(N,T_min,T_max,q*f_max), "QuantisedPeriodogramq" + q)
-for( q <- List(4.0) ) runbench(Ns, (N : Int) => new QuantisedPeriodogramChirpZ(N,T_min,T_max,q*T_max), "QuantisedPeriodogramChirpZq" + q)
+for( q <- List(5.0) ) runbench(Ns, (N : Int) => new QuantisedPeriodogramFFT(N,T_min,T_max,q*f_max), "QuantisedPeriodogramq" + q)
+for( q <- List(5.0) ) runbench(Ns, (N : Int) => new QuantisedPeriodogramChirpZ(N,T_min,T_max,q*f_max), "QuantisedPeriodogramChirpZq" + q)
 
 //benchmarks for Barry's chirpz tests.
 //for( q <- List(4.0) ) runbench(Ns, (N : Int) => new QuantisedPeriodogramFFT(N,T_min,T_max,q*f_max), "QuantisedPeriodogramFFTq" + q)

### 2014-03-31 code/montecarlo.scala

@@ -34,11 +34,11 @@ for( N <- List(30,1200) ) {
   for( m <- List( 1, 10 ) ) {
     def discretedist() = sparsedist(m)
     val varbs = -38.0 to -5.0 by 1.5 //variances we use in dB
-    //runsim(varbs, N, () => discretedist, () => new YeSampling(new PeriodogramEstimator(N,T_min,T_max)), "PeriodogramN" + N + "geom" + m)
-    runsim(varbs, N, () => discretedist, () => new YeSampling(new NormalisedSamplingLLS(N,T_min,T_max),10.0), "NormalisedSamplingLLSNsamp10" + N + "geom" + m)
+    runsim(varbs, N, () => discretedist, () => new YeSampling(new PeriodogramEstimator(N,T_min,T_max)), "PeriodogramN" + N + "geom" + m)
+    runsim(varbs, N, () => discretedist, () => new YeSampling(new NormalisedSamplingLLS(N,T_min,T_max),10.0), "NormalisedSamplingLLSN" + N + "geom" + m)
     //runsim(varbs, N, () => discretedist, () => new YeSampling(new SLS2novlp(N,T_min,T_max)), "SLS2novlpN" + N + "geom" + m)
-    //runsim(varbs, N, () => discretedist, () => new YeSampling(new SLS2new(N,T_min,T_max)), "SLS2newN" + N + "geom" + m)
-    //for( q <- List(1.5,2.0,4.0,5.0) ) runsim(varbs, N, () => discretedist, () => new QuantisedPeriodogramFFT(N,T_min,T_max,q*f_max), "QuantisedPeriodogramFFTN" + N + "q" + q + "geom" + m)
+    runsim(varbs, N, () => discretedist, () => new YeSampling(new SLS2new(N,T_min,T_max)), "SLS2newN" + N + "geom" + m)
+    for( q <- List(1.5,2.0,4.0,5.0) ) runsim(varbs, N, () => discretedist, () => new QuantisedPeriodogramFFT(N,T_min,T_max,q*f_max), "QuantisedPeriodogramFFTN" + N + "q" + q + "geom" + m)
 
     //now compute CLTs
     val vardbsfiner = -38.0 to -5.0 by 0.2

### 2014-03-31 paper.tex
added new data minor

@@
 \usepackage{pgfplots}
+\pgfplotsset{compat=1.8}
 
@@
     \selectcolormodel{gray} 
-    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.19,0.24)}},xlabel style={at={(0.5,0.14)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=7, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
+    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.24)}},xlabel style={at={(0.5,0.07)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=7, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
       \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30geom1}; 
@@
       \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN30geom1};
-       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/NormalisedSamplingLLSN30geom1}; 
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN30geom1}; 
  %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN30geom1};
@@
        \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN1200geom1};
-       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/NormalisedSamplingLLSN1200geom1};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSNsamp101200geom1};
  %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1200geom1};
@@
      \legend{Periodogram theory, Least squares theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
-     \node[rotate=17,font=\footnotesize] at (axis cs:6e-4,9e-8) {$N=30$};
+     \node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=30$};
      \node[rotate=17,font=\footnotesize] at (axis cs:7e-4,1.5e-12) {$N=1200$}; 
    \end{axis} 
- \begin{axis}[name=plot2,at={(11,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.19,0.23)}},xlabel style={at={(0.5,0.14)}}]
+ \begin{axis}[name=plot2,at={(11,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.23)}},xlabel style={at={(0.5,0.07)}}]
       \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30geom10}; 
@@
       \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN30geom10};
-       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/NormalisedSamplingLLSN30geom10}; 
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN30geom10}; 
   %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN30geom10};
@@
        \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN1200geom10};
-       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/NormalisedSamplingLLSN1200geom10};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN1200geom10};
  %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1100geom10};
@@
 
-This section presents the results of Monte-Carlo simulations with the periodogram estimator~\cite{Haohuan2013435,Fogel1989_bit_synch_zero_crossings}, the normalised least squares estimator~\cite{Clarkson2007,McKilliam2007,Quinn_sparse_noisy_SSP_2012}, Clarkson's modification of the separable least squares line search (SLS2-new) estimator~\cite{Clarkson2007,Sidiropoulos2005}, and the quantised periodogram estimator from Section~\ref{sec:quant-peri}.
-The grid distance $D$ used for the modified least squares, periodogram, and SLS2-new estimators is given by~\eqref{eq:gridwidthrecommended}.
+This section presents the results of Monte-Carlo simulations with the periodogram estimator~\cite{Haohuan2013435,Fogel1989_bit_synch_zero_crossings}, the modified least squares estimator~\cite{Clarkson2007,McKilliam2007,Quinn_sparse_noisy_SSP_2012}, Clarkson's modification of the separable least squares line search (SLS2-new) estimator~\cite{Clarkson2007,Sidiropoulos2005}, and the quantised periodogram estimator from Section~\ref{sec:quant-peri}.
+The grid distance $D$ used for the periodogram and SLS2-new estimators is given by~\eqref{eq:gridwidthrecommended}.
+The modified least squares estimator requires a finer grid to work accurately and we choose~\eqref{eq:gridwidthrecommended} divided by three in our simulations.
 The quantised periodogram estimator is computed using a single fast Fourier transform of length $M = 2L = 2(\ellmax - \ellmin + 1)$.

### 2014-03-31 code/benchmark.scala

@@ -34,11 +34,11 @@ val Ns = List(10,14,20,30,40,50,60,80,100,120,150,200,250,300,400,600,900,1200,1
 //val Ns = (1 to 40).map(i => (5*scala.math.pow(1.1,i)).toInt).distinct //number of symbols
 
 //runbench(Ns, (N : Int) => new YeSampling(new PeriodogramEstimator(N,T_min,T_max)), "Periodogram")
-//runbench(Ns, (N : Int) => new YeSampling(new NormalisedSamplingLLS(N,T_min,T_max)), "LeastSquares")
+runbench(Ns, (N : Int) => new YeSampling(new NormalisedSamplingLLS(N,T_min,T_max),6.0), "LeastSquares")
 //runbench(Ns, (N : Int) => new YeSampling(new SLS2novlp(N,T_min,T_max)), "SLS2novlp")
 //runbench(Ns, (N : Int) => new YeSampling(new SLS2new(N,T_min,T_max)), "SLS2new")
-for( q <- List(5.0) ) runbench(Ns, (N : Int) => new QuantisedPeriodogramFFT(N,T_min,T_max,q*f_max), "QuantisedPeriodogramq" + q)
-for( q <- List(5.0) ) runbench(Ns, (N : Int) => new QuantisedPeriodogramChirpZ(N,T_min,T_max,q*f_max), "QuantisedPeriodogramChirpZq" + q)
+//for( q <- List(5.0) ) runbench(Ns, (N : Int) => new QuantisedPeriodogramFFT(N,T_min,T_max,q*f_max), "QuantisedPeriodogramq" + q)
+//for( q <- List(5.0) ) runbench(Ns, (N : Int) => new QuantisedPeriodogramChirpZ(N,T_min,T_max,q*f_max), "QuantisedPeriodogramChirpZq" + q)
 
 //benchmarks for Barry's chirpz tests.
 //for( q <- List(4.0) ) runbench(Ns, (N : Int) => new QuantisedPeriodogramFFT(N,T_min,T_max,q*f_max), "QuantisedPeriodogramFFTq" + q)

### 2014-03-31 code/montecarlo.scala

@@ -34,11 +34,11 @@ for( N <- List(30,1200) ) {
   for( m <- List( 1, 10 ) ) {
     def discretedist() = sparsedist(m)
     val varbs = -38.0 to -5.0 by 1.5 //variances we use in dB
-    runsim(varbs, N, () => discretedist, () => new YeSampling(new PeriodogramEstimator(N,T_min,T_max)), "PeriodogramN" + N + "geom" + m)
-    runsim(varbs, N, () => discretedist, () => new YeSampling(new NormalisedSamplingLLS(N,T_min,T_max),10.0), "NormalisedSamplingLLSN" + N + "geom" + m)
+    //runsim(varbs, N, () => discretedist, () => new YeSampling(new PeriodogramEstimator(N,T_min,T_max)), "PeriodogramN" + N + "geom" + m)
+    runsim(varbs, N, () => discretedist, () => new YeSampling(new NormalisedSamplingLLS(N,T_min,T_max),6.0), "NormalisedSamplingLLSN" + N + "geom" + m)
     //runsim(varbs, N, () => discretedist, () => new YeSampling(new SLS2novlp(N,T_min,T_max)), "SLS2novlpN" + N + "geom" + m)
-    runsim(varbs, N, () => discretedist, () => new YeSampling(new SLS2new(N,T_min,T_max)), "SLS2newN" + N + "geom" + m)
-    for( q <- List(1.5,2.0,4.0,5.0) ) runsim(varbs, N, () => discretedist, () => new QuantisedPeriodogramFFT(N,T_min,T_max,q*f_max), "QuantisedPeriodogramFFTN" + N + "q" + q + "geom" + m)
+    //runsim(varbs, N, () => discretedist, () => new YeSampling(new SLS2new(N,T_min,T_max)), "SLS2newN" + N + "geom" + m)
+    //for( q <- List(1.5,2.0,4.0,5.0) ) runsim(varbs, N, () => discretedist, () => new QuantisedPeriodogramFFT(N,T_min,T_max,q*f_max), "QuantisedPeriodogramFFTN" + N + "q" + q + "geom" + m)
 
     //now compute CLTs
     val vardbsfiner = -38.0 to -5.0 by 0.2

### 2014-03-31 paper.tex

@@
     \selectcolormodel{gray} 
-    \begin{axis}[font=\footnotesize,xmode=log,ymode=log,height=7cm,width=9cm,xlabel={$N$},ylabel={time (ms)},ylabel style={at={(0.215,0.41)}},xlabel style={at={(0.5,0.205)}}, legend style={draw=none,fill=none,legend pos=north west,cells={anchor=west},font=\footnotesize}]
+    \begin{axis}[font=\footnotesize,xmode=log,ymode=log,height=7cm,width=9cm,xlabel={$N$},ylabel={time (ms)},ylabel style={at={(0.08,0.41)}},xlabel style={at={(0.5,0.1)}}, legend style={draw=none,fill=none,legend pos=north west,cells={anchor=west},font=\footnotesize}]
       \addplot[mark=o,mark options={scale=1}] table {code/data/PeriodogramBenchmark};
-      \addplot[mark=x,mark options={solid,fill=black,scale=1}] table {code/data/LeastSquaresBenchmark};
+      \addplot[mark=x,mark options={solid,fill=black,scale=1.3}] table {code/data/LeastSquaresBenchmark};
       \addplot[mark=*,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramq4.0Benchmark};
@@
       \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN30q5.0geom10};
-      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN30geom10};
-       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN30geom10}; 
+      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN30geom10};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN30geom10}; 
   %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN30geom10};
@@
       \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1200geom10};
-       \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN1200geom10};
-       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN1200geom10};
+       \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN1200geom10};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN1200geom10};
  %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1100geom10};
@@
 The grid distance $D$ used for the periodogram and SLS2-new estimators is given by~\eqref{eq:gridwidthrecommended}.
-The modified least squares estimator requires a finer grid to work accurately and we choose~\eqref{eq:gridwidthrecommended} divided by three in our simulations.
+The modified least squares estimator requires a smaller grid distance to work accurately for smaller noise variance and we choose~\eqref{eq:gridwidthrecommended} divided by three in our simulations.
 The quantised periodogram estimator is computed using a single fast Fourier transform of length $M = 2L = 2(\ellmax - \ellmin + 1)$.
@@
 %For small $\sigma^2$ the mean square error of the SLS2 estimator is almost three orders of magnitude less accurate when $N = 50$ and almost five orders of magnitude less accurate when $N=1000$.
-The solid line shows the mean square error of the periodogram estimator predicted by Theorem~2 from~\cite{Quinn20013asilomar_period_est} and the dashed line the mean square error of the modified least squares estimator predicted by Theorem~3 from~\cite{Quinn_sparse_noisy_SSP_2012}.
+The solid line shows the mean square error of the periodogram estimator predicted by Theorem~2 from~\cite{Quinn20013asilomar_period_est}.
+The dashed line shows the mean square error of the modified least squares estimator predicted by Theorem~3 from~\cite{Quinn_sparse_noisy_SSP_2012}.
 
@@
 The chirp z-transform is slower in these simulations, but is expected to be faster when $N$ is exceptionally large.
-The quantised periodogram estimator computed using a single fast Fourier transform is faster than the periodogram estimator when $N \geq 20$.
+%The quantised periodogram estimator computed using a single fast Fourier transform is faster than the periodogram estimator when $N \geq 20$.
 It is faster than the least squares estimator when $N \geq 80$ and it is faster than the SLS2-new estimator when $N \geq 400$.
-The quantised periodogram estimator is significantly faster than the other estimator for large $N$.
+The quantised periodogram estimator is significantly faster than the other estimators for large $N$.
 For example, when $N = 25000$ the periodogram estimator requires approximately 530 seconds, whereas the quantised periodogram estimator requires approximately 990 milliseconds.

### 2014-04-01 paper.tex
minor

@@
 The grid distance $D$ used for the periodogram and SLS2-new estimators is given by~\eqref{eq:gridwidthrecommended}.
-The modified least squares estimator requires a smaller grid distance to work accurately for smaller noise variance and we choose~\eqref{eq:gridwidthrecommended} divided by three in our simulations.
+The modified least squares estimator requires a smaller grid distance to work accurately and we choose~\eqref{eq:gridwidthrecommended} divided by three in our simulations.
 The quantised periodogram estimator is computed using a single fast Fourier transform of length $M = 2L = 2(\ellmax - \ellmin + 1)$.
@@
 The quantised periodogram estimator is significantly faster than the other estimators for large $N$.
-For example, when $N = 25000$ the periodogram estimator requires approximately 530 seconds, whereas the quantised periodogram estimator requires approximately 990 milliseconds.
+For example, when $N = 25000$ the periodogram estimator requires approximately 530 seconds, whereas the quantised periodogram estimator requires approximately 1 second.
 The software is written in Java and the computer used is an Intel Core2 Q6600 running at \unit[2.4]{GHz}.

### 2014-06-11 response/response.tex
started added to responses

\documentclass[a4paper,10pt]{article}
%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
%\documentclass[journal]{IEEEtran}

\usepackage{../mathbf-abbrevs}
\input{../defs.tex}

\usepackage{zref-xr,zref-user}
\zexternaldocument*{paper}

\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\usepackage{amsmath,amsfonts,amssymb, amsthm, bm}

%\usepackage[square,comma,numbers,sort&compress]{natbib}

\usepackage{color}
\newcommand{\comment}[1]{\textcolor{red}{#1}}

\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\sinc}{\operatorname{sinc}}
\newcommand{\rect}[1]{\operatorname{rect}\left(#1\right)}

%opening
\title{Fast sparse period estimation: Reviewer responses}
\author{R.~G.~McKilliam, I.~V.~L.~Clarkson and B.~G.~Quinn 
%\thanks{}
}

\renewcommand{\theenumii}{(\alph{enumii})}
\renewcommand{\labelenumii}{\theenumii}

\begin{document}

\maketitle

\section*{Reviewer 1}\label{sec:reviewer-1}

\begin{enumerate}

\item\textbf{Comment:}
This paper proposes a new fast method for estimating the period of sparse and noisy
observations of a point process. The new idea is to used the periodogram of quantized
observations that can be computed faster than the standard periodogram by using
a chirp z-transform or a fast Fourier transform. The reviewer is not convinced that
there is enough material deserving publication in the IEEE signal processing letters.
Moreover, there are several important comments that should be addressed before a
possible publication of this letter.
\\
\textbf{Response:}
BLERG: Do I want to say something it being simple, but novel, and potentially really useful!

\item\textbf{Comment:}
The main reviewer concern is about the choice of the quantization parameter
q that should be explained more carefully. It is not suﬃcient to display two sets
of curves obtained for two diﬀerent values of q and to choose the best one without
explanations. For instance, it is not clear whether the value of q chosen in this paper
will be appropriate for other noise distributions or for other ways of generating the
integers sn. This point should be addressed carefully before a possible resubmission
of the paper. The other reviewer comments are summarized below.
\\
\textbf{Response:}
BLERG: Run some extra simulations for them

\item\textbf{Comment:}
p. 3, the authors mention “Sadler and Casey [24]... However, their estimator
requires far larger transforms and produces less accurate results than the estimator
described here (see Section VI)". Unfortunately, there is nothing in
Section VI (which is the conclusion) supporting this claim. The authors should
compare their results in term of accuracy and of computational complexity to
the results of [24]. Including the Cramér-Rao bounds for the estimation of T0
(as in [24]) would also be appreciated.
\\
\textbf{Response:}
BLERG:

\item\textbf{Comment:}
p. 7, the authors indicate that they have used the Newton-Raphson (NR)
method. More details should be provided here. For instance, what is the exact
NR scheme used in this paper (there are several options)? How have the
authors selected the total number of iterations? Is the quantized periodogram
just used to initialize a Newton-Raphson procedure? Do the MSEs displayed
in Fig. 3 correspond to the estimator obtained after applying the NR algorithm?
Also, do the results displayed in Fig. 2 correspond to the total time
required to compute the estimator (including the NR procedure) or just to the
initialization of the NR procedure?
\\
\textbf{Response:}
BLERG:

\item\textbf{Comment:}
p. 5, in the definition of $v_m$, $b_{m+\ell_1}$ should be replaced by $b_{m+\ell_{\text{min}}}$.
\\
\textbf{Response:}
Thankyou for spotting this error.  A similar error was fixed in equation for $I_z$ directly above this.

\item\textbf{Comment:}
p. 6, ``each of length $L+K-1$'' might be replaced by ``each of length $M = L + K - 1$"
\\
\textbf{Response:}
We have not done this.  The parameter $M$ is reserved for the length of the FFT used for the single Fourier transform method.  We feel it will lead to confusion if $M$ is also used for the length of the transforms required for the chirp-z transform.

\item\textbf{Comment:}
p. 8, $n = 2,\dots,N$ should be $n = 1, \dots, N - 1$
\\
\textbf{Response:}

\item\textbf{Comment:}
p. 8, it is not completely clear to identify the curves associated with $N = 30$ and $N = 1200$. Can the authors display 4 curves associated with $\mu= 1$, $\mu = 10$, $N = 30$ and $N = 1200$?
\\
\textbf{Response:}
BLERG:

\item\textbf{Comment:}
Typos: p. 2, “anomolies" should be “anomalies", p. 7, “quanitised" should be
“quantized".
\\
\textbf{Response:}
BLERG:

\item\textbf{Comment:}
Is it necessary to include 11 self citations in this letter? Removing some of
these references would allow to save some space, which could be used to better
explain some points mentioned above.
\\
\textbf{Response:}
IEEE Signal Processing Letters allows a 5th page that can contain only references.  Therefore, we cannot save space by removing references.   We have removed reference [29] from the original manuscript because the material in that paper is also covered in [27].  We feel that all 36 other references are strongly relevant and that the paper would be weakened without them.

\end{enumerate}

\section*{Reviewer 2}

\begin{enumerate}

\item\textbf{Comment:}
 The identifiability claim at the beginning of sec. II is not obvious and a reference is mandatory. Indeed, referring to eq. (2), let us assume that $s_n = 2 i_n$ where $i_n$ are all non-negative integer. In this case the point process is periodic with period $2 T_0$. This referee does not believe that maximizing the periodogram over $(f_{\text{min}},f_{\text{max}})$ will identify the true frequency (or period). LS methods might have the capabilities to identify the period, but, in any case the claim of idenitifiability needs to be further supported.
\\
\textbf{Response:}

\item\textbf{Comment:}
 Is the equality in the second (un-numbered) equation of page 2 correct?
\\
\textbf{Response:}

\item\textbf{Comment:}
Newton-Rapson iteration to further improve the frequency estimation is advocated in Sec. II. Can a reference supporting this statement be added?
\\
\textbf{Response:}

\item\textbf{Comment:}
Referring to Fig. 3, in case $N=30$, the novel method incurs in a significant loss of performance with respect to classical LS and periodogram for low-signal-to-noise ratios. In this scenario (low $N$ and low signal-to-noise-ration) relevant for real applications?
\\
\textbf{Response:}

\end{enumerate}

\bibstyle{IEEEtran}
\bibliography{bib}

\end{document}

### 2014-06-11 paper.tex
minor?
started added to responses

@@
 Least squares estimation of period has been studied in~\cite{Clarkson2007,McKilliam2007,Quinn_sparse_noisy_SSP_2012}.
-The least squares estimator can be closely approximated using $O(N^2)$ operations by using special purpose algorithms from computational number theory~\cite{McKilliam2009CoxeterLattices,McKilliam2008,McKilliam2008b}.
+The least squares estimator can be closely approximated using $O(N^2)$ operations by using special purpose algorithms from computational number theory~\cite{McKilliam2009CoxeterLattices,McKilliam2008}.
 The least squares estimator is also the maximum likelihood estimator in the case that the noise variables $w_1,\dots,w_N$ are independent and identically normally distributed.
@@
 \[
-I_z(f) = \babs{ \sum_{i = \ellmin}^{\ellmax} e^{ -2\pi j f i/q} b_i }^2 = \babs{ \sum_{m = 0}^{\ellmax-\ellmin} e^{ -2\pi j f m/q} b_{m+\ell_1} }^2.
+I_z(f) = \babs{ \sum_{i = \ellmin}^{\ellmax} e^{ -2\pi j f i/q} b_i }^2 = \babs{ \sum_{m = 0}^{\ellmax-\ellmin} e^{ -2\pi j f m/q} b_{m+\ellmin} }^2.
 \]
@@
 \[
-v_m = e^{ -2\pi j f_{\text{min}}m/q} b_{m+\ell_1}
+v_m = e^{ -2\pi j f_{\text{min}}m/q} b_{m+\ellmin}
 \]
@@
 However, it is apparent that the length of the transforms, and hence the complexity, increases as the quantisation becomes finer, i.e., as $q$ increases.
-This presents a trade off between computationally complexity and quantisation accuracy.
+This presents a trade off between computational complexity and quantisation accuracy.
 This trade off is investigated in the next section.

### 2014-06-11 response/response.tex
added to response

@@
 \usepackage{zref-xr,zref-user}
-\zexternaldocument*{paper}
+\zexternaldocument*{../paper}
 
@@
 \textbf{Response:}
-BLERG:
+Section VI in~\cite{726812}, i.e., the paper is not the conclusion.
+It is the section entitled ``Numerical Results''.
+It is clear from Figure 1 to 4 in~\ref{726812} that the periogram estimator implemented by Casey and Sadler is inferior to their MEA estimator.
+Casey and Sadler do use an FFT to compute the periodgram, but fail to implement an optimisation proceedure.
+We used the Newton-Raphson proceedure in the old manuscript, but have used Brent's method~\cite{Brent_opt_no_derivs_1973} is used as dicussed in our response.
+This leads Casey and Sadler to the conclusion that very large FFTs are required and that the periodgram estimator is not particularly accurate.
+In the conclusion of~\cite{726812} it is stated that:
+
+``Comparisons with the periodogram show the limitations of
+periodogram performance.''
+
+Our paper clearly shows this is not the case.
+The periodgram estimator, computed by quanitisation and FFT and combined with an appropriate optimisation proceedure, is both compuationally efficient and highly accurate.
+Our paper shows that (for moderately large $N$) this is the most computationally efficient proceedure currently available for this problem.
 
@@
 \textbf{Response:}
-BLERG:
+BLERG: In Section~\zref{sec:peri-estim} the text has been modifed to read:
+
+``An iterative procedure, such as the Newton-Raphson method, or Brent's method~\cite{Brent_opt_no_derivs_1973} 
+can then be used to compute $\hat{f}$ starting from $\widetilde{f}$.''
 
@@
 \textbf{Response:}
+This error has be fixed.  The same error in the caption of Figure~\zref{plot:multipleN} has been fixed.
 
@@
 \textbf{Response:}
-BLERG:
+We have connected the data points for the $N=1200$ case by gray lines.  We believe this make the plot clear.  We are happy to make further adjustments to this plot should the reviewer require them.
 
@@
 We have removed reference [29] from the original manuscript because the material in that paper is also covered in [27].
-We feel that all 36 other references are strongly relevant and that the paper would be weakened without them.
+We feel that the all 37 references in the new version of the paper are highly relevant.
+The paper would be weakened without them.
 
@@
 \textbf{Response:}
+Yes.  The terms inside the sum on the left and right are conjugates of one another.  That is,
+\[
+\sum_{n=1}^N e^{ 2\pi j f y_n} = \sum_{n=1}^N (e^{-2\pi j f y_n})^* = \left( \sum_{n=1}^N e^{ -2\pi j f y_n} \right)^*
+\]
+where $^*$ denotes the complex conjugate.  So,
+\[
+I_y(f) = \babs{ \sum_{n=1}^N e^{ 2\pi j f y_n} }^2 = \babs{ \sum_{n=1}^N e^{ -2\pi j f y_n} }^2.
+\]
 
@@
 \textbf{Response:}
+The quantised periodgram estimator does not suffer a significant performance loss with respect to the classical LS and periodogram estimators in any scenario we have found.
+Observe in Figure 3 that the solid dots ($q=5 f_{\text{max}}$) the crosses (LS), the circles (periodogram) are all close for $N=30,1200$ and $\mu=1,10$.
+Be aware that the solid lines indicates the asymptotic theory ($N\to\infty$)~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est} for the periodogram and LS estimators and not the measured accuracy for finite $N$.
+As $N\to\infty$, the MSE of the periogram and LS estimators will converge to the solid and dashed curves.
 
@@
 
-\bibstyle{IEEEtran}
-\bibliography{bib}
+{\small
+\input{../paper.bbl}
+}
 

### 2014-06-11 paper.tex
added to response

@@
 If the distance $D$ is sufficiently small, then $\widetilde{f}$ is a good approximation to the maximiser $\hat{f}$.
-An iterative procedure, such as the Newton-Raphson method, %, or Brent's method~\cite{Brent_opt_no_derivs_1973}
+An iterative procedure, such as the Newton-Raphson method, or Brent's method~\cite{Brent_opt_no_derivs_1973}
 can then be used to compute $\hat{f}$ starting from $\widetilde{f}$.
@@
 %      \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/NormalisedSamplingLLSN200};
-      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN1200geom1};
-      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1200geom1};
-       \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN1200geom1};
-       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSNsamp101200geom1};
+      \addplot[color=gray,mark=none,color=black] table {code/data/PeriodogramCLTN1200geom1};
+      \addplot[color=gray,mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1200geom1};
+       \addplot[color=gray,mark=o,mark options={color=black,solid,fill=black,scale=1}] table {code/data/PeriodogramN1200geom1};
+       \addplot[color=gray,mark=x,mark options={color=black,solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSNsamp101200geom1};
  %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1200geom1};
-       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN1200geom1};
-       \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q1.5geom1};
+       \addplot[color=gray,mark=square,mark options={solid,color=black,fill=black,scale=1}] table {code/data/SLS2newN1200geom1};
+       \addplot[color=gray,mark=triangle,mark options={solid,color=black,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q1.5geom1};
       %\addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q3};
-      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q5.0geom1};
+      \addplot[color=gray,mark=*,mark options={solid,color=black,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q5.0geom1};
  %     \legend{Periodogram Theory, Least squares theory, Quantised periodogram, Periodogram, Least squares}
@@
       \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1200geom10};
-       \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN1200geom10};
-       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN1200geom10};
+       \addplot[color=gray,mark=o,mark options={solid,color=black,fill=black,scale=1.1}] table {code/data/PeriodogramN1200geom10};
+       \addplot[color=gray,mark=x,mark options={solid,color=black,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN1200geom10};
  %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1100geom10};
-       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN1200geom10};
-       \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q1.5geom10};
+       \addplot[color=gray,mark=square,mark options={solid,color=black,fill=black,scale=1}] table {code/data/SLS2newN1200geom10};
+       \addplot[color=gray,mark=triangle,mark options={solid,color=black,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q1.5geom10};
       %\addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1100q3};
-      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q5.0geom10};
+      \addplot[color=gray,mark=*,mark options={solid,color=black,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q5.0geom10};
  %     \legend{Periodogram Theory, Least squares theory, Quantised periodogram, Periodogram, Least squares}
@@
   \caption{Mean square period error versus noise variance $\sigma^2$.
-The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=2,\dots,N$ are independent and identically geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right.}\label{plot:multipleN}
+The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right.
+Data points for $N=1200$ are connected by gray lines.}\label{plot:multipleN}
 \end{figure*}
@@
 In all simulations the noise variables $w_1,\dots,w_N$ are independent and identically normally distributed with variance $\sigma^2$.
-The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=2,\dots,N$ are independent and identically geometrically distributed starting from zero with means $\mu=1$ and $10$.
+The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed starting from zero with means $\mu=1$ and $10$.
 The true period is $T_0 = \pi/3$, the true phase is $\theta_0 = 0.2$, and $T_{\text{min}} = 0.8$ and $T_{\text{max}} = 1.5$.

### 2014-06-11 paper.tex
added more the response and paper

@@
 Existing accurate algorithms for estimating the period require $O(N^2)$ operations where $N$ is the number of observations.
-By quantising the observations we produce an estimator that requires only $O(N\log N)$ operations by use of the chirp z-transform or the fast Fourier transform.
-The quantisation has the adverse effect of decreasing the accuracy of the estimator.
+By quantizing the observations we produce an estimator that requires only $O(N\log N)$ operations by use of the chirp z-transform or the fast Fourier transform.
+The quantization has the adverse effect of decreasing the accuracy of the estimator.
 This is investigated by Monte-Carlo simulation.
@@
 Solutions have applications to synchronisation for telecommunications devices~\cite{Fogel1988,Fogel1989_bit_synch_zero_crossings,Sidiropoulos2005,5621928}, estimation of the pulse repetition interval in electronic support~\cite{EltonGray_puilse_train_rader_1994,Gray_more_pri_1994,Clarkson_thesis,clarkson_estimate_period_pulse_train_1996,Hauochan_pri_2012}, and beat and tempo estimation in music~\cite{dixon_beat_extraction_2001}.
-The problem is closely related to spike interval estimation in neurology~\cite{Arnett_neuro_pri_1976,Brillinger_spike_trains_1988,Rossoni200630} and a similar model has been proposed for the detection of anomolies, such as denial-of-service attacks, in internet traffic~\cite{He_detecting_periodic_patterns_in_internet_2006,5585849,5947313}.
+The problem is closely related to spike interval estimation in neurology~\cite{Arnett_neuro_pri_1976,Brillinger_spike_trains_1988,Rossoni200630} and a similar model has been proposed for the detection of anomalies, such as denial-of-service attacks, in internet traffic~\cite{He_detecting_periodic_patterns_in_internet_2006,5585849,5947313}.
 The problem is also connected with classical problems in number theory such as Diophantine approximation, lattice reduction, and the study of prime numbers~\cite{Cassels_geom_numbers_1997,490557,Clarkson_thesis,Lenstra1982,Wubben_2011,Casey1995,726812,CaseySadler_primes_2013}.
@@
 Least squares estimation of period has been studied in~\cite{Clarkson2007,McKilliam2007,Quinn_sparse_noisy_SSP_2012}.
-The least squares estimator can be closely approximated using $O(N^2)$ operations by using special purpose algorithms from computational number theory~\cite{McKilliam2009CoxeterLattices,McKilliam2008}.
+The least squares estimator can be closely approximated using $O(N^2)$ operations using special purpose algorithms~\cite{McKilliam2009CoxeterLattices,McKilliam2008}.
 The least squares estimator is also the maximum likelihood estimator in the case that the noise variables $w_1,\dots,w_N$ are independent and identically normally distributed.
@@
 
-In Section~\ref{sec:quant-peri} we show that, by quantising the observations $y_1,\dots,y_N$, an approximation to the periodogram estimator can be computed using only $O(N\log N)$ operations.
+In Section~\ref{sec:quant-peri} we show that, by quantizing the observations $y_1,\dots,y_N$, an approximation to the periodogram estimator can be computed using only $O(N\log N)$ operations.
 This is facilitated by either the chirp z-transform~\cite{Rabiner1969} or the fast Fourier transform.
-The quantisation has the adverse affect of decreasing the accuracy of the estimator.
-The estimator is accurate if a sufficiently fine quantiser is chosen, but this requires a larger transform.
-The trade off between quantisation level, accuracy, and computational complexity is investigated by computer simulation in Section~\ref{sec:simulations}.
+The quantization has the adverse affect of decreasing the accuracy of the estimator.
+The estimator is accurate if a sufficiently fine quantizer is chosen, but this requires a larger transform.
+The trade off between quantization level, accuracy, and computational complexity is investigated by computer simulation in Section~\ref{sec:simulations}.
 The simulations indicate that significant computational savings are possible with negligible loss in accuracy.
@@
 \]
-over the interval $(f_{\text{min}}, f_{\text{max}}) = (1/T_{\text{max}}, 1/T_{\text{min}})$ and $j = \sqrt{-1}$.  That is,
+over the interval $(f_{\text{min}}, f_{\text{max}}) = (1/T_{\text{max}}, 1/T_{\text{min}})$.  That is,
 \[
@@
 \[
-G = \{ f_{\text{min}} + Dk \mid k = 0, \dots K-1 \},
+G = \{ f_{\text{min}} + Dk \mid k = 0, \dots K-1 \}
 \]
@@
 If the distance $D$ is sufficiently small, then $\widetilde{f}$ is a good approximation to the maximiser $\hat{f}$.
-An iterative procedure, such as the Newton-Raphson method, or Brent's method~\cite{Brent_opt_no_derivs_1973}
+An iterative procedure, such as the Newton-Raphson method, or Brent's method~\cite[Ch.~5]{Brent_opt_no_derivs_1973}
 can then be used to compute $\hat{f}$ starting from $\widetilde{f}$.
@@
 
-\section{Quantisation and the periodogram}\label{sec:quant-peri}
+\section{Quantization and the periodogram}\label{sec:quant-peri}
 
@@
 Half integers are rounded up in our implementation.
-Let $q$ be a real number and define the quantisation function
+Let $q$ be a real number and define the quantization function
 \[
@@
 \]
-that rounds its argument to a nearest multiple of $1/q$.  Large $q$ corresponds with fine quantisation and small $q$ corresponds with coarse quantisation.  Compute quantised observations
+that rounds its argument to a nearest multiple of $1/q$.  Large $q$ corresponds with fine quantization and small $q$ corresponds with coarse quantization.  Compute quantized observations
 \[
@@
 \]
-where $\ell_n = \round{qy_n} \in \ints$.  The periodogram of the quantised observations $z_1,\dots,z_N$ is
+where $\ell_n = \round{qy_n} \in \ints$.  The periodogram of the quantized observations $z_1,\dots,z_N$ is
 \begin{align*}
@@
 \]
-If the quantisation parameter $q$ is sufficiently large we expect that $I_z$ is close to $I_y$ and so $\widetilde{f}^\prime$ will be close to $\widetilde{f}$.
+If the quantization parameter $q$ is sufficiently large we expect that $I_z$ is close to $I_y$ and so $\widetilde{f}^\prime$ will be close to $\widetilde{f}$.
 Thus, if $q$ is sufficiently large and the distance $D$ is sufficiently small we expect $\widetilde{f}^\prime$ to be a good approximation to $\hat{f}$.
@@
 can then be used to compute $\hat{f}$ starting from $\widetilde{f}^\prime$.  
-Observe that $I_z$ is periodic with period $q$.  This immediately indicates that we should choose the quantisation fine enough that 
+Observe that $I_z$ is periodic with period $q$.  This indicates that quantization should be fine enough that 
 \begin{equation}\label{eq:boundaliasingP}
@@
 \]
-$A = 1$, $W = e^{ -2\pi j D/q}$ and $V_0,\dots,V_{K-1}$ can be efficiently computed by the chirp z-transform~\cite{Rabiner1969}.  Putting
+where $A = 1$, and $W = e^{ -2\pi j D/q}$.  All of $V_0,\dots,V_{K-1}$ can be efficiently computed by the chirp z-transform~\cite{Rabiner1969}.  Putting
 \[
@@
 \]
-The frequency $\widetilde{f}^\prime$ can be used to initialize an iterative procedure to maximise $I_y$.
+%The frequency $\widetilde{f}^\prime$ can be used to initialize an iterative procedure to maximise $I_y$.
 The procedure will converge to $\hat{f}$ if $\widetilde{f}^\prime$ is sufficiently close to $\hat{f}$.
@@
 We have found this to be faster than the chirp z-transform for all but exceptionally large $N$.
-Suppose that we choose the distance $D$ and quantisation parameter $q$ so that $M = q/D$ is an integer and
+Suppose that we choose the distance $D$ and quantization parameter $q$ so that $M = q/D$ is an integer and
 \[
@@
 We now describe recommended values for the parameters $q, M$, and $D$.
-The simulations presented in Section~\ref{sec:simulations} suggest that a reasonable value for the quantisation parameter is $q = 5 f_{\text{max}}$.
+The simulations presented in Section~\ref{sec:simulations} suggest that a reasonable value for the quantization parameter is $q = 5 f_{\text{max}}$.
 If the chirp z-transform is used then $D$ is chosen according to~\eqref{eq:gridwidthrecommended}.
@@
 Both approaches require $O(qN\log(qN)) = O(N\log N)$ operations because $q$ does not depend on $N$.
-However, it is apparent that the length of the transforms, and hence the complexity, increases as the quantisation becomes finer, i.e., as $q$ increases.
-This presents a trade off between computational complexity and quantisation accuracy.
+However, it is apparent that the length of the transforms, and hence the complexity, increases as the quantization becomes finer, i.e., as $q$ increases.
+This presents a trade off between computational complexity and quantization accuracy.
 This trade off is investigated in the next section.
@@
   \end{tikzpicture}  
-  \caption{Computation time in milliseconds versus $N$ for the periodogram, least squares, and quantised periodogram estimators computed using the chirp z-transform and a single fast Fourier transform.}\label{plot:benchmark}
+  \caption{Computation time in milliseconds versus $N$ for the periodogram, least squares, and quantized periodogram estimators computed using the chirp z-transform and a single fast Fourier transform.}\label{plot:benchmark}
 \end{figure}
@@
 
-This section presents the results of Monte-Carlo simulations with the periodogram estimator~\cite{Haohuan2013435,Fogel1989_bit_synch_zero_crossings}, the modified least squares estimator~\cite{Clarkson2007,McKilliam2007,Quinn_sparse_noisy_SSP_2012}, Clarkson's modification of the separable least squares line search (SLS2-new) estimator~\cite{Clarkson2007,Sidiropoulos2005}, and the quantised periodogram estimator from Section~\ref{sec:quant-peri}.
+This section presents the results of Monte-Carlo simulations with the periodogram estimator~\cite{Haohuan2013435,Fogel1989_bit_synch_zero_crossings}, the modified least squares estimator~\cite{Clarkson2007,McKilliam2007,Quinn_sparse_noisy_SSP_2012}, Clarkson's modification of the separable least squares line search (SLS2-new) estimator~\cite{Clarkson2007,Sidiropoulos2005}, and the quantized periodogram estimator from Section~\ref{sec:quant-peri}.
 The grid distance $D$ used for the periodogram and SLS2-new estimators is given by~\eqref{eq:gridwidthrecommended}.
 The modified least squares estimator requires a smaller grid distance to work accurately and we choose~\eqref{eq:gridwidthrecommended} divided by three in our simulations.
-The quantised periodogram estimator is computed using a single fast Fourier transform of length $M = 2L = 2(\ellmax - \ellmin + 1)$.
+%The quantized periodogram estimator is computed using a single fast Fourier transform of length $M = 2L = 2(\ellmax - \ellmin + 1)$.
 Both the periodogram and quanitised periodogram estimators use the Newton-Raphson method.
@@
 Three thousand Monte-Carlo trials are run for each value of $\sigma^2$.
-The plots show the mean square error of the quantised periodogram estimator for two quantisation parameters $q = \tfrac{3}{2}f_{\text{max}} = \tfrac{15}{8}$ and $q = 5 f_{\text{max}} = \tfrac{25}{4}$.
+The plots show the mean square error of the quantized periodogram estimator for two quantization parameters $q = \tfrac{3}{2}f_{\text{max}} = \tfrac{15}{8}$ and $q = 5 f_{\text{max}} = \tfrac{25}{4}$.
 The estimator performs poorly when $q = \tfrac{3}{2}f_{\text{max}}$ (triangles) but performs almost identically to the original periodogram estimator (circles) when $q = 5 f_{\text{max}}$ (dots).
-For comparison purposes, the plot also shows the mean square error of the modified least squares estimator~\cite{Clarkson2007,McKilliam2007} and the SLS2-new estimator~\cite{Clarkson2007,Sidiropoulos2005}.
+The plot also shows the mean square error of the modified least squares estimator~\cite{Clarkson2007,McKilliam2007} and the SLS2-new estimator~\cite{Clarkson2007,Sidiropoulos2005}.
 %The statistical performance of the SLS2-new estimator is poor when compared with the other estimators.
@@
 Running times for other values of $\mu$ lead to similar conclusions.
-The quantisation parameter is $q = 5 f_{\text{max}}$ so that the performance of the quantised periodogram estimator is similar to that of the original periodogram estimator as observed in Figure~\ref{plot:multipleN}.
-The average running times for the quantised periodogram estimator computed using the chirp z-transform and using a single fast Fourier transform are shown.
+The quantization parameter is $q = 5 f_{\text{max}}$ so that the performance of the quantized periodogram estimator is similar to that of the original periodogram estimator as observed in Figure~\ref{plot:multipleN}.
+The average running times for the quantized periodogram estimator computed using the chirp z-transform and using a single fast Fourier transform are shown.
 The chirp z-transform is slower in these simulations, but is expected to be faster when $N$ is exceptionally large.
-%The quantised periodogram estimator computed using a single fast Fourier transform is faster than the periodogram estimator when $N \geq 20$.
+%The quantized periodogram estimator computed using a single fast Fourier transform is faster than the periodogram estimator when $N \geq 20$.
 It is faster than the least squares estimator when $N \geq 80$ and it is faster than the SLS2-new estimator when $N \geq 400$.
-The quantised periodogram estimator is significantly faster than the other estimators for large $N$.
-For example, when $N = 25000$ the periodogram estimator requires approximately 530 seconds, whereas the quantised periodogram estimator requires approximately 1 second.
+The quantized periodogram estimator is significantly faster than the other estimators for large $N$.
+For example, when $N = 25000$ the periodogram estimator requires approximately 530 seconds, whereas the quantized periodogram estimator requires approximately 1 second.
 The software is written in Java and the computer used is an Intel Core2 Q6600 running at \unit[2.4]{GHz}.
@@
 We have considered the problem of estimating the period of a point process from $N$ observations that are both sparse and noisy.
-By quantising the observations we describe an estimator that can be computed in $O(N\log N)$ operations by use of the fast Fourier transform or chirp z-transform.
-This is in contrast to existing accurate estimators that require at least $O(N^2)$ operations.
-The trade off between computational complexity and quantisation accuracy was examined by simulation.
+By quantizing the observations we describe an estimator that can be computed in $O(N\log N)$ operations by use of the fast Fourier transform or chirp z-transform.
+This improved upon existing accurate estimators that require at least $O(N^2)$ operations.
+The trade off between computational complexity and quantization accuracy was examined by simulation.
 The simulations indicate that significant computational savings are possible with negligible loss in statistical accuracy.

### 2014-06-11 response/response.tex

@@
 \item\textbf{Comment:}
-p. 3, the authors mention “Sadler and Casey [24]... However, their estimator
+p. 3, the authors mention “Sadler and Casey [24] \dots However, their estimator
 requires far larger transforms and produces less accurate results than the estimator
@@
 \textbf{Response:}
-Section VI in~\cite{726812}, i.e., the paper is not the conclusion.
+Section VI in~\cite{726812} is not the conclusion.
 It is the section entitled ``Numerical Results''.
-It is clear from Figure 1 to 4 in~\ref{726812} that the periogram estimator implemented by Casey and Sadler is inferior to their MEA estimator.
+It is clear from Figure 1 to 4 in~\cite{726812} that the periogram estimator implemented by Casey and Sadler is inferior to their MEA estimator.
 Casey and Sadler do use an FFT to compute the periodgram, but fail to implement an optimisation proceedure.
-We used the Newton-Raphson proceedure in the old manuscript, but have used Brent's method~\cite{Brent_opt_no_derivs_1973} is used as dicussed in our response.
+We used the Newton-Raphson proceedure in the old manuscript, but have used Brent's method~\cite{Brent_opt_no_derivs_1973} in the new manuscript as dicussed in our response to comment~\ref{com:rev1:nr} below.
 This leads Casey and Sadler to the conclusion that very large FFTs are required and that the periodgram estimator is not particularly accurate.
@@
 
-``Comparisons with the periodogram show the limitations of
-periodogram performance.''
+``Comparisons with the periodogram show the limitations of periodogram performance.''
 
-Our paper clearly shows this is not the case.
+Our paper clearly shows that such limitation do not exist.
 The periodgram estimator, computed by quanitisation and FFT and combined with an appropriate optimisation proceedure, is both compuationally efficient and highly accurate.
 Our paper shows that (for moderately large $N$) this is the most computationally efficient proceedure currently available for this problem.
+Simultaneously, it is as accurate at the most accuraate estimators available, namely the least squares estimator, and the periodgram estimator computed directly without FFT.
 
-\item\textbf{Comment:}
+BLERG: Cramer-Rao bound not available for randomly chosen $s_1,\dots,s_N$.
+
+\item\textbf{Comment:}\label{com:rev1:nr}
 p. 7, the authors indicate that they have used the Newton-Raphson (NR)
@@
 \textbf{Response:}
+We are only aware of a single Netwon-Raphson method that proceeds by the iteration
+\[
+x_{n+1} = x_n - \frac{f^\prime(x_{n})}{f^{\prime\prime}(x_n)}
+\]
+to find a stationary point of the function $f$ where $f^\prime$ and $f^{\prime\prime}$ are the first and second derivatives of $f$.
+As was stated Section~\zref{sec:simulations}: ``Both the periodogram and quanitised periodogram estimators use the Newton-Raphson method.''  This was the case both for the simulations presented in Figure~\zref{plot:multipleN} and for the computational benchmarks presented in Figure~\zref{plot:benchmark}.
+
+To avoid the potential ambguity regarding the definition of the Newton-Raphson method we have, in the new manuscipt, replaced the Newton-Raphson method with the optimisation proceedure due to Brent~\cite[Ch.~5]{Brent_opt_no_derivs_1973}.
+This optimiation method is likely to be more familiar with the target audience because it is used internally by the \texttt{fminbnd} function in the popular MATLAB environment.
+
 BLERG: In Section~\zref{sec:peri-estim} the text has been modifed to read:
@@
 \textbf{Response:}
-BLERG:
+Done.
 
@@
 \textbf{Response:}
+BLERG
 

### 2014-06-11 response/response.tex
added minor

@@
 \textbf{Response:}
-BLERG: Do I want to say something it being simple, but novel, and potentially really useful!
+We agree with the reviewer that the results in this paper are simple.
+We do not believe this to be a weakness.
+Despite being simple, the results are novel and useful.
+The quantisated periodogram estimator described in this paper is as accurate as the most accurate estimators in the literature (the least squares and periodogram estimator),  but, for moderate and large data sets our estimator is substantially more computationally simple.
+For these reasons, we firmly believe this will become the estimator of choice for periodic point processes.
+It enables period estimation for extremely large data sets in reasonable time.
+This is not possible with previous estimators that require a number of operations growing quadratically with data size.
+We have provided all the simulation code as supplementary material.
+This will enable simple implementation for future research.
 
@@
 The main reviewer concern is about the choice of the quantization parameter
-q that should be explained more carefully. It is not suﬃcient to display two sets
-of curves obtained for two diﬀerent values of q and to choose the best one without
-explanations. For instance, it is not clear whether the value of q chosen in this paper
+$q$ that should be explained more carefully. It is not suﬃcient to display two sets
+of curves obtained for two diﬀerent values of $q$ and to choose the best one without
+explanations. For instance, it is not clear whether the value of $q$ chosen in this paper
 will be appropriate for other noise distributions or for other ways of generating the
-integers sn. This point should be addressed carefully before a possible resubmission
+integers $s_n$. This point should be addressed carefully before a possible resubmission
 of the paper. The other reviewer comments are summarized below.
@@
 \textbf{Response:}
-BLERG: Run some extra simulations for them
+We anticipate that developing a rigorous theoretical method for choosing $q$ will be extremely complicated.
+The properties of many optimisation proceedures, such as the Newton-Raphson method, and Brent's method~\cite[Ch.~5]{Brent_opt_no_derivs_1973} are reasonably well understood for the purpose of maximising fixed known functions.
+However, very few theorectical statistical analyses for these proceedures exist when the functions to be maximised are themselves random variables, i.e., when noise is involved.
+Add to this the existence of quantisation (itself typically very tricky to analyse) and it appears that a theorectical analysis of this estimator will be extremely difficult.
+
+These theoretical difficulties do not negate the immediate practical applicability of our estimator.
+We agree with the reviewer that the simulations presented are limited.
+This is entirely due to the 4 page limit required IEEE Signal Processing Letters.
+In the paper we use our simulations to motivate the choice $q = 5 f_{\text{max}}$ for the quantisation parameter.
+It is natural to wonder whether this choice also works in other scenarios, for example (as the reviewer suggests), for different methods of generating the integers $s_1,\dots,s_N$.
 

### 2014-06-11 code/java/src/snpe/NormalisedPeriodogram.java

@@ -36,7 +36,8 @@ public class NormalisedPeriodogram extends PeriodogramEstimator  {
                 fhat = f;
             }
         }
-        fhat = refine(fhat, y); //refine coarse frequency estimate to local maximiser of normalised periodogram
+        //fhat = refine(fhat, y); //refine coarse frequencey estimate using Newton's method
+        fhat = refine(fhat, fstep, y); //refine coarse frequencey estimate using Brent's method
         That = 1.0 / fhat;
         phat = phasestor.getPhase(y, That); //now compute the phase estimate
     }

### 2014-06-11 code/java/src/snpe/PeriodogramEstimator.java

@@ -1,5 +1,8 @@
 package snpe;
 
+import pubsim.optimisation.Brent;
+import pubsim.optimisation.SingleVariateFunction;
+
 /**
  * Written by Vaughan Clarkson, 05-Jan-07. New method calculatePeriodogram and
  * fixed the Newton iteration steps, 08-Jan-07.
@@ -48,7 +51,8 @@ public class PeriodogramEstimator extends AbstractPRIEstimator.Sampled {
                 fhat = f;
             }
         }
-        fhat = refine(fhat, y); //refine coarse frequencey estimate
+        //fhat = refine(fhat, y); //refine coarse frequencey estimate using Newton's method
+        fhat = refine(fhat, fstep, y); //refine coarse frequencey estimate using Brent's method
         That = 1.0 / fhat;
         phat = phasestor.getPhase(y, That); //now compute the phase estimate
     }
@@ -94,6 +98,17 @@ public class PeriodogramEstimator extends AbstractPRIEstimator.Sampled {
             numIter++;
         }
         return fhat;
+    }
+    
+    ///Brent's method
+    protected double refine(double fhat, double fstep, final Double[] y) {
+        SingleVariateFunction periodogram = new SingleVariateFunction() {
+            @Override
+            public double value(double x) {
+                return -calculatePeriodogram(y,x);
+            }
+        };
+        return new Brent(periodogram,fhat - fstep/2, fhat, fhat + fstep/2).xmin();
     }
 
     @Override

### 2014-06-11 code/java/src/snpe/QuantisedPeriodogramChirpZ.java

@@ -46,7 +46,8 @@ public class QuantisedPeriodogramChirpZ extends PeriodogramEstimator {
             }
         }        
         double ftilde = fmin + D*khat;
-        double fhat = refine(ftilde, y); //refine coarse frequency estimate
+        //double fhat = refine(fhat, y); //refine coarse frequencey estimate using Newton's method
+        double fhat = refine(ftilde, D, y); //refine coarse frequencey estimate using Brent's method
         That = 1.0 / fhat;
         phat = phasestor.getPhase(y, That); //now compute the phase estimate        
     }

### 2014-06-11 code/java/src/snpe/QuantisedPeriodogramFFT.java

@@ -48,7 +48,8 @@ public class QuantisedPeriodogramFFT extends PeriodogramEstimator {
             }
         }        
         double ftilde = fmin + D*khat;
-        double fhat = refine(ftilde, y); //refine coarse frequency estimate
+        //double fhat = refine(fhat, y); //refine coarse frequencey estimate using Newton's method
+        double fhat = refine(ftilde, D, y); //refine coarse frequencey estimate using Brent's method
         That = 1.0 / fhat;
         phat = phasestor.getPhase(y, That); //now compute the phase estimate        
     }

### 2014-06-12 response/response.tex
minor

@@
 \zexternaldocument*{../paper}
+
+\usepackage{tikz}
+\usetikzlibrary{calc}
+\usepackage{pgfplots}
+\pgfplotsset{compat=1.8}
 
@@
 These theoretical difficulties do not negate the immediate practical applicability of our estimator.
+We believe it would be unwise to halt dissemination of these useful and practical results simply because a difficult theoretical problem has not yet been solved.
 We agree with the reviewer that the simulations presented are limited.
@@
 It is natural to wonder whether this choice also works in other scenarios, for example (as the reviewer suggests), for different methods of generating the integers $s_1,\dots,s_N$.
+We have conducted such simulations and found the choice $q=5 f_{\text{max}}$ appears to be a good choice in all of these settings.
+Figures~\ref{plot:poisson} to~\ref{?} in this document display simulation results for the following methods of generating the integers $s_1,\dots,s_N$:
+
+\begin{enumerate}
+\item In Figure~\ref{plot:poisson} the integers are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Poisson distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right.
+\item 
+\end{enumerate}
+
+\begin{figure*}[t] 
+  \centering 
+  \begin{tikzpicture} 
+    \selectcolormodel{gray} 
+    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.24)}},xlabel style={at={(0.5,0.07)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=7, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
+      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30poisson1}; 
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN30poisson1};
+            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q1.5poisson1};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN30q5.0poisson1};
+      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN30poisson1};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN30poisson1}; 
+ %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN30poisson1};
+      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN30poisson1};
+%      \addplot[mark=none,dashed,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q2};
+ %     \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200};
+ %     \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200}; 
+ %     \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/PeriodogramN200};
+ %     \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q3};
+ %     \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q10};
+%      \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/NormalisedSamplingLLSN200};
+      \addplot[color=gray,mark=none,color=black] table {code/data/PeriodogramCLTN500poisson1};
+      \addplot[color=gray,mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN500poisson1};
+       \addplot[color=gray,mark=o,mark options={color=black,solid,fill=black,scale=1}] table {code/data/PeriodogramN500poisson1};
+       \addplot[color=gray,mark=x,mark options={color=black,solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN500poisson1};
+ %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN500poisson1};
+       \addplot[color=gray,mark=square,mark options={solid,color=black,fill=black,scale=1}] table {code/data/SLS2newN500poisson1};
+       \addplot[color=gray,mark=triangle,mark options={solid,color=black,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN500q1.5poisson1};
+      %\addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN500q3};
+      \addplot[color=gray,mark=*,mark options={solid,color=black,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN500q5.0poisson1};
+ %     \legend{Periodogram Theory, Least squares theory, Quantised periodogram, Periodogram, Least squares}
+     \legend{Periodogram theory, Least squares theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
+     \node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=30$};
+     \node[rotate=17,font=\footnotesize] at (axis cs:7e-4,1.5e-12) {$N=500$}; 
+   \end{axis} 
+ \begin{axis}[name=plot2,at={(11,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.23)}},xlabel style={at={(0.5,0.07)}}]
+      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30poisson10}; 
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN30poisson10};
+            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q1.5poisson10};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN30q5.0poisson10};
+      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN30poisson10};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN30poisson10}; 
+  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN30poisson10};
+       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN30poisson10};
+%      \addplot[mark=none,dashed,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q2};
+ %     \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200};
+ %     \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200}; 
+ %     \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/PeriodogramN200};
+ %     \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q3};
+ %     \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q10};
+%      \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/NormalisedSamplingLLSN200}; 
+      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN500poisson10};
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN500poisson10};
+       \addplot[color=gray,mark=o,mark options={solid,color=black,fill=black,scale=1.1}] table {code/data/PeriodogramN500poisson10};
+       \addplot[color=gray,mark=x,mark options={solid,color=black,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN500poisson10};
+ %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1100poisson10};
+       \addplot[color=gray,mark=square,mark options={solid,color=black,fill=black,scale=1}] table {code/data/SLS2newN500poisson10};
+       \addplot[color=gray,mark=triangle,mark options={solid,color=black,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN500q1.5poisson10};
+      %\addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1100q3};
+      \addplot[color=gray,mark=*,mark options={solid,color=black,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN500q5.0poisson10};
+ %     \legend{Periodogram Theory, Least squares theory, Quantised periodogram, Periodogram, Least squares}
+     %\legend{Periodogram theory, $q=\tfrac{3}{2}T_{\text{max}}$, $q=4T_{\text{max}}$, Periodogram, Least squares, SLS2-adj}
+     \node[rotate=16,font=\footnotesize] at (axis cs:6e-4,9e-10) {$N=30$};
+     \node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=500$}; 
+   \end{axis}
+  \end{tikzpicture}  
+  \caption{Mean square period error versus noise variance $\sigma^2$.
+The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Poisson distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right.
+Data points for $N=500$ are connected by gray lines.}\label{plot:poisson}
+\end{figure*}
 
@@
 \textbf{Response:}
-Section VI in~\cite{726812} is not the conclusion.
+Section VI in the paper~\cite{726812} entitled ``On periodic pulse interval analysis with outliers and missing observations'' is not the conclusion.
 It is the section entitled ``Numerical Results''.
-It is clear from Figure 1 to 4 in~\cite{726812} that the periogram estimator implemented by Casey and Sadler is inferior to their MEA estimator.
+From Figure 1 to 4 in~\cite{726812} the periogram estimator implemented by Casey and Sadler is seen to be inferior to their MEA estimator.
 Casey and Sadler do use an FFT to compute the periodgram, but fail to implement an optimisation proceedure.
 We used the Newton-Raphson proceedure in the old manuscript, but have used Brent's method~\cite{Brent_opt_no_derivs_1973} in the new manuscript as dicussed in our response to comment~\ref{com:rev1:nr} below.
-This leads Casey and Sadler to the conclusion that very large FFTs are required and that the periodgram estimator is not particularly accurate.
+Casey and Sadler arrive at the conclusion that very large FFTs are required and that the periodgram estimator is not particularly accurate.
 In the conclusion of~\cite{726812} it is stated that:
@@
 
-BLERG: Cramer-Rao bound not available for randomly chosen $s_1,\dots,s_N$.
+In the simulations we regenerate the integers $s_1,\dots,s_N$ on each Monte-Carlo trial.
+To the authors knowledge, a Cramer-Rao bound is not known in this case.
+The \emph{clairvoyant} Cramer-Rao bound from~\cite{Sidiropoulos2005} (also in \cite{726812,Clarkson2007}) supposes a single fixed realisation for $s_1,\dots,s_N$.
+This bound is not valid for our simulations.
+However, the curves derived from the asymptotic theory for the least squares and periodogram estimators~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est} are valid in this case.
+The least squares estimator is the maximum likelihood estimator, and so, the dashed curve serves a similar type of purpose as the CRB often does in these types of simulations.
+Namely, to indicate whether an estimator is performing similarly to what might be considered `optimal', i.e.
+close the what might theoretically be expected from the maximum likelihood estimator.
 
@@
 To avoid the potential ambguity regarding the definition of the Newton-Raphson method we have, in the new manuscipt, replaced the Newton-Raphson method with the optimisation proceedure due to Brent~\cite[Ch.~5]{Brent_opt_no_derivs_1973}.
-This optimiation method is likely to be more familiar with the target audience because it is used internally by the \texttt{fminbnd} function in the popular MATLAB environment.
+This optimiation method is likely to be more familiar with readers because it is used internally by the \texttt{fminbnd} function in the popular MATLAB environment.
 

### 2014-06-12 paper.tex
minor

@@
 The problem is closely related to spike interval estimation in neurology~\cite{Arnett_neuro_pri_1976,Brillinger_spike_trains_1988,Rossoni200630} and a similar model has been proposed for the detection of anomalies, such as denial-of-service attacks, in internet traffic~\cite{He_detecting_periodic_patterns_in_internet_2006,5585849,5947313}.
-The problem is also connected with classical problems in number theory such as Diophantine approximation, lattice reduction, and the study of prime numbers~\cite{Cassels_geom_numbers_1997,490557,Clarkson_thesis,Lenstra1982,Wubben_2011,Casey1995,726812,CaseySadler_primes_2013}.
+The problem is also connected with classical problems in number theory such as Diophantine approximation, lattice reduction, and the study of prime numbers~\cite{Cassels_geom_numbers_1997,490557,Clarkson_thesis,Lenstra1982,Wubben_2011,536682,726812,CaseySadler_primes_2013}.
 
@@
        \addplot[color=gray,mark=o,mark options={color=black,solid,fill=black,scale=1}] table {code/data/PeriodogramN1200geom1};
-       \addplot[color=gray,mark=x,mark options={color=black,solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSNsamp101200geom1};
+       \addplot[color=gray,mark=x,mark options={color=black,solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN1200geom1};
  %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1200geom1};

### 2014-06-12 code/java/src/snpe/AbstractPRIEstimator.java

@@ -22,7 +22,7 @@ public abstract class AbstractPRIEstimator implements PRIEstimator {
     }
     
     @Override
-    public int getLength() {
+    public int length() {
         return N;
     }
     

### 2014-06-12 code/java/src/snpe/NormalisedPeriodogram.java

@@ -39,7 +39,7 @@ public class NormalisedPeriodogram extends PeriodogramEstimator  {
         //fhat = refine(fhat, y); //refine coarse frequencey estimate using Newton's method
         fhat = refine(fhat, fstep, y); //refine coarse frequencey estimate using Brent's method
         That = 1.0 / fhat;
-        phat = phasestor.getPhase(y, That); //now compute the phase estimate
+        phat = phasestor.phase(y, That); //now compute the phase estimate
     }
 
     ///Modified Newton step minimses the normalised periodogram

### 2014-06-12 code/java/src/snpe/NormalisedSamplingLLS.java

@@ -52,7 +52,7 @@ public class NormalisedSamplingLLS extends SamplingLLS {
         That = 1 / fhat;
 
         //now compute the phase estimate
-        phat = phasestor.getPhase(y, That);
+        phat = phasestor.phase(y, That);
     }
     
 }

### 2014-06-12 code/java/src/snpe/NormalisedZnLLS.java

@@ -71,7 +71,7 @@ public class NormalisedZnLLS extends ZnLLS implements PRIEstimator {
         That = 1.0/bestf;
 
         //now compute the phase estimate
-        phat = phasestor.getPhase(y, That);
+        phat = phasestor.phase(y, That);
 
 
     }

### 2014-06-12 code/java/src/snpe/PRIEstimator.java

@@ -15,22 +15,23 @@ public interface PRIEstimator extends Serializable{
 
     /**
      * Runs the estimator and return the estimated frequency.
+     * @param y the data
      */
     public void estimate(double[] y);
 
     /**
      * @return the period estimate
      */
-    public double getPeriod();
+    public double period();
 
     /**
      *
      * @return the phase estimate
      */
-    public double getPhase();
+    public double phase();
 
     /// @return the number of observations this estimator uses.
-    public int getLength();
+    public int length();
     
     /// Interface for period estimators that operate by sampling their objective functions
     public static interface Sampled extends PRIEstimator {
@@ -57,7 +58,7 @@ public interface PRIEstimator extends Serializable{
             fz = new double[N];
         }
 
-        public double getPhase(double[] y, double That){
+        public double phase(double[] y, double That){
             Anstar.project(y, z);
             for(int i = 0; i < N; i++) fz[i] =  z[i] / That;
             lattice.nearestPoint(fz);

### 2014-06-12 code/java/src/snpe/PeriodogramEstimator.java

@@ -54,7 +54,7 @@ public class PeriodogramEstimator extends AbstractPRIEstimator.Sampled {
         //fhat = refine(fhat, y); //refine coarse frequencey estimate using Newton's method
         fhat = refine(fhat, fstep, y); //refine coarse frequencey estimate using Brent's method
         That = 1.0 / fhat;
-        phat = phasestor.getPhase(y, That); //now compute the phase estimate
+        phat = phasestor.phase(y, That); //now compute the phase estimate
     }
 
     /** Modified Newton step */
@@ -112,12 +112,12 @@ public class PeriodogramEstimator extends AbstractPRIEstimator.Sampled {
     }
 
     @Override
-    public double getPeriod() {
+    public double period() {
         return That;
     }
 
     @Override
-    public double getPhase() {
+    public double phase() {
         return phat;
     }
     

### 2014-06-12 code/java/src/snpe/QuantisedPeriodogramChirpZ.java

@@ -49,7 +49,7 @@ public class QuantisedPeriodogramChirpZ extends PeriodogramEstimator {
         //double fhat = refine(fhat, y); //refine coarse frequencey estimate using Newton's method
         double fhat = refine(ftilde, D, y); //refine coarse frequencey estimate using Brent's method
         That = 1.0 / fhat;
-        phat = phasestor.getPhase(y, That); //now compute the phase estimate        
+        phat = phasestor.phase(y, That); //now compute the phase estimate        
     }
 
     private Complex[] computeV(final double D, final int K) {

### 2014-06-12 code/java/src/snpe/QuantisedPeriodogramFFT.java

@@ -51,7 +51,7 @@ public class QuantisedPeriodogramFFT extends PeriodogramEstimator {
         //double fhat = refine(fhat, y); //refine coarse frequencey estimate using Newton's method
         double fhat = refine(ftilde, D, y); //refine coarse frequencey estimate using Brent's method
         That = 1.0 / fhat;
-        phat = phasestor.getPhase(y, That); //now compute the phase estimate        
+        phat = phasestor.phase(y, That); //now compute the phase estimate        
     }
     
     /// Calculates the vector v and returns its Fourier transform computed by FFT

### 2014-06-12 code/java/src/snpe/SLS2all.java

@@ -64,16 +64,16 @@ public class SLS2all extends AbstractPRIEstimator.Sampled {
 	That = 1/fhat;
 
         //now compute the phase estimate
-        phat = phasestor.getPhase(y, That);
+        phat = phasestor.phase(y, That);
     }
 
     @Override
-    public double getPeriod() {
+    public double period() {
         return That;
     }
 
     @Override
-    public double getPhase() {
+    public double phase() {
         return phat;
     }
 

### 2014-06-12 code/java/src/snpe/SLS2new.java

@@ -62,16 +62,16 @@ public class SLS2new extends AbstractPRIEstimator.Sampled {
         That = 1 / fhat;
 
         //now compute the phase estimate
-        phat = phasestor.getPhase(y, That);
+        phat = phasestor.phase(y, That);
     }
 
     @Override
-    public double getPeriod() {
+    public double period() {
         return That;
     }
 
     @Override
-    public double getPhase() {
+    public double phase() {
         return phat;
     }
 }

### 2014-06-12 code/java/src/snpe/SLS2novlp.java

@@ -62,16 +62,16 @@ public class SLS2novlp extends AbstractPRIEstimator.Sampled {
         That = 1 / fhat;
 
         //now compute the phase estimate
-        phat = phasestor.getPhase(y, That);
+        phat = phasestor.phase(y, That);
     }
 
     @Override
-    public double getPeriod() {
+    public double period() {
         return That;
     }
 
     @Override
-    public double getPhase() {
+    public double phase() {
         return phat;
     }
 }

### 2014-06-12 code/java/src/snpe/SamplingLLS.java

@@ -65,16 +65,16 @@ public class SamplingLLS extends AbstractPRIEstimator.Sampled {
         That = 1 / fhat;
 
         //now compute the phase estimate
-        phat = phasestor.getPhase(y, That);
+        phat = phasestor.phase(y, That);
     }
 
     @Override
-    public double getPeriod() {
+    public double period() {
         return That;
     }
 
     @Override
-    public double getPhase() {
+    public double phase() {
         return phat;
     }
     

### 2014-06-12 code/java/src/snpe/YeSampling.java

@@ -36,18 +36,18 @@ public class YeSampling implements PRIEstimator {
     }
 
     @Override
-    public double getPeriod() {
-        return periodest.getPeriod();
+    public double period() {
+        return periodest.period();
     }
 
     @Override
-    public double getPhase() {
-        return periodest.getPhase();
+    public double phase() {
+        return periodest.phase();
     }
 
     @Override
-    public int getLength() {
-        return periodest.getLength();
+    public int length() {
+        return periodest.length();
     }
     
 }

### 2014-06-12 code/java/src/snpe/ZnLLS.java

@@ -102,18 +102,18 @@ public class ZnLLS extends AbstractPRIEstimator {
         That = 1.0/bestf;
 
         //now compute the phase estimate
-        phat = phasestor.getPhase(y, That);
+        phat = phasestor.phase(y, That);
 
         
     }
     
     @Override
-    public double getPeriod() {
+    public double period() {
         return That;
     }
 
     @Override
-    public double getPhase() {
+    public double phase() {
         return phat;
     }
     

### 2014-06-12 code/java/test/snpe/NormalisedSamplingLLSTest.java

@@ -54,9 +54,9 @@ public class NormalisedSamplingLLSTest {
         double[] y = sig.generate();
 
         instance.estimate(y);
-        System.out.println(instance.getPeriod() + "\t" + instance.getPhase());
-        assertEquals(T, instance.getPeriod(), 0.001);
-        assertEquals(phase, instance.getPhase(), 0.001);
+        System.out.println(instance.period() + "\t" + instance.phase());
+        assertEquals(T, instance.period(), 0.001);
+        assertEquals(phase, instance.phase(), 0.001);
     }
 
 }

### 2014-06-12 code/java/test/snpe/PeriodogramEstimatorTest.java

@@ -58,9 +58,9 @@ public class PeriodogramEstimatorTest {
         double[] y = sig.generate();
 
         instance.estimate(y);
-        System.out.println(instance.getPeriod() + "\t" + instance.getPhase());
-        assertEquals(T, instance.getPeriod(), 0.001);
-        assertEquals(phase, instance.getPhase(), 0.001);
+        System.out.println(instance.period() + "\t" + instance.phase());
+        assertEquals(T, instance.period(), 0.001);
+        assertEquals(phase, instance.phase(), 0.001);
     }
     
 }

### 2014-06-12 code/java/test/snpe/QuantisedPeriodogramChirpZTest.java

@@ -58,9 +58,9 @@ public class QuantisedPeriodogramChirpZTest {
         double[] y = sig.generate();
 
         instance.estimate(y);
-        System.out.println(instance.getPeriod() + "\t" + instance.getPhase());
-        assertEquals(T, instance.getPeriod(), 0.001);
-        assertEquals(phase, instance.getPhase(), 0.001);
+        System.out.println(instance.period() + "\t" + instance.phase());
+        assertEquals(T, instance.period(), 0.001);
+        assertEquals(phase, instance.phase(), 0.001);
     }
     
 }

### 2014-06-12 code/java/test/snpe/QuantisedPeriodogramFFTTest.java

@@ -100,9 +100,9 @@ public class QuantisedPeriodogramFFTTest {
         double[] y = sig.generate();
 
         instance.estimate(y);
-        System.out.println(instance.getPeriod() + "\t" + instance.getPhase());
-        assertEquals(T, instance.getPeriod(), 0.001);
-        assertEquals(phase, instance.getPhase(), 0.001);
+        System.out.println(instance.period() + "\t" + instance.phase());
+        assertEquals(T, instance.period(), 0.001);
+        assertEquals(phase, instance.phase(), 0.001);
     }
     
 }

### 2014-06-12 code/java/test/snpe/YeSamplingTest.java

@@ -58,9 +58,9 @@ public class YeSamplingTest {
         double[] y = sig.generate();
 
         instance.estimate(y);
-        System.out.println(instance.getPeriod() + "\t" + instance.getPhase());
-        assertEquals(T, instance.getPeriod(), 0.001);
-        assertEquals(phase, instance.getPhase(), 0.001);
+        System.out.println(instance.period() + "\t" + instance.phase());
+        assertEquals(T, instance.period(), 0.001);
+        assertEquals(phase, instance.phase(), 0.001);
     }
     
     
@@ -86,9 +86,9 @@ public class YeSamplingTest {
         double[] y = sig.generate();
 
         instance.estimate(y);
-        System.out.println(instance.getPeriod() + "\t" + instance.getPhase());
-        assertEquals(T, instance.getPeriod(), 0.001);
-        assertEquals(phase, instance.getPhase(), 0.001);
+        System.out.println(instance.period() + "\t" + instance.phase());
+        assertEquals(T, instance.period(), 0.001);
+        assertEquals(phase, instance.phase(), 0.001);
     }
     
 }

### 2014-06-12 code/java/test/snpe/ZnLLSTest.java

@@ -46,7 +46,7 @@ public class ZnLLSTest extends TestCase {
         
         double expResult = T;
         instance.estimate(y);
-        assertEquals(expResult, instance.getPeriod(), 0.001);
+        assertEquals(expResult, instance.period(), 0.001);
 
     }
 
@@ -74,8 +74,8 @@ public class ZnLLSTest extends TestCase {
         double[] y = sig.generate();
 
         instance.estimate(y);
-        double hatT = instance.getPeriod();
-        double hatp = instance.getPhase();
+        double hatT = instance.period();
+        double hatp = instance.phase();
 
         System.out.println(hatT + ", " + hatp);
 

### 2014-06-12 code/montecarlo.scala

@@ -1,11 +1,12 @@
 /**
  * Run simulations of various sparse noise period estimators.
  */
+import snpe.generators.DifferencesIID
 import pubsim.distributions.discrete.DiscreteRandomVariable
 import pubsim.distributions.discrete.GeometricRandomVariable
 import pubsim.distributions.GaussianNoise
 import pubsim.distributions.RealRandomVariable
-import snpe.SparseNoisyPeriodicSignal
+import snpe.generators.SparseNoisyPeriodicSignal
 import snpe.PRIEstimator
 import snpe.YeSampling
 import snpe.PeriodogramEstimator
@@ -16,29 +17,28 @@ import snpe.SLS2new
 import snpe.bounds.PeriodogramCLT
 import snpe.bounds.NormalisedLLSCLT
 
-val iters = 3000
+val iters = 100
 val T_0 = scala.math.Pi/3 //true period
 val T_min = 0.8 //minimum period
 val T_max = 1.5 //maximum period
 val f_min = 1.0/T_max //minimum frequency
 val f_max = 1.0/T_min //maximum frequency
 val theta_0 = 0.2 //true phase
-def sparsedist(m : Double) = new GeometricRandomVariable.StartingAtZero(1.0/(m+1.0))
 def noisedist(v : Double) = new GaussianNoise(0,v) //the noise distribution we use with variance v
 
 val starttime = (new java.util.Date).getTime
 
-for( N <- List(30,1200) ) {
-  
-  //for discrete distributions with mean 3 and 15
+for( N <- List(30,1200) ) {  
+  //for discrete distributions with mean 1 and 10
   for( m <- List( 1, 10 ) ) {
-    def discretedist() = sparsedist(m)
+    def discretedist() =  new GeometricRandomVariable.StartingAtZero(1.0/(m+1.0))
+    def sparsenoisygenerator(v : Double) = new SparseNoisyPeriodicSignal(N,T_0,theta_0,new DifferencesIID(N,discretedist),noisedist(v))
     val varbs = -38.0 to -5.0 by 1.5 //variances we use in dB
-    runsim(varbs, N, () => discretedist, () => new YeSampling(new PeriodogramEstimator(N,T_min,T_max)), "PeriodogramN" + N + "geom" + m)
[…]

### 2014-06-12 response/code/montecarlo.scala

@@ -1,0 +1,126 @@
+/**
+ * Run simulations of various sparse noise period estimators.
+ */
+import snpe.generators.DifferencesIID
+import pubsim.distributions.discrete.DiscreteRandomVariable
+import pubsim.distributions.discrete.GeometricRandomVariable
+import pubsim.distributions.discrete.PoissonRandomVariable
+import pubsim.distributions.GaussianNoise
+import pubsim.distributions.RealRandomVariable
+import snpe.generators.SparseNoisyPeriodicSignal
+import snpe.PRIEstimator
+import snpe.YeSampling
+import snpe.PeriodogramEstimator
+import snpe.NormalisedSamplingLLS
+import snpe.QuantisedPeriodogramFFT
+import snpe.SLS2novlp
+import snpe.SLS2new
+import snpe.bounds.PeriodogramCLT
+import snpe.bounds.NormalisedLLSCLT
+
+val iters = 100
+val T_0 = scala.math.Pi/3 //true period
+val T_min = 0.8 //minimum period
+val T_max = 1.5 //maximum period
+val f_min = 1.0/T_max //minimum frequency
+val f_max = 1.0/T_min //maximum frequency
+val theta_0 = 0.2 //true phase
+def noisedist(v : Double) = new GaussianNoise(0,v) //the noise distribution we use with variance v
+
+val starttime = (new java.util.Date).getTime
+
+//trials with Poisson discrete random variables
+for( N <- List(30,500) ) {  
+  //for discrete distributions with mean 1 and 10
+  for( m <- List( 1, 10 ) ) {
+    def discretedist() = new PoissonRandomVariable(m);
+    def sparsenoisygenerator(v : Double) = new SparseNoisyPeriodicSignal(N,T_0,theta_0,new DifferencesIID(N,discretedist),noisedist(v))
+    val varbs = -38.0 to -5.0 by 1.5 //variances we use in dB
+    runsim(varbs, sparsenoisygenerator, () => new YeSampling(new PeriodogramEstimator(N,T_min,T_max)), "PeriodogramN" + N + "poisson" + m)
+    runsim(varbs, sparsenoisygenerator, () => new YeSampling(new NormalisedSamplingLLS(N,T_min,T_max),6.0), "NormalisedSamplingLLSN" + N + "poisson" + m)
+    //runsim(varbs, sparsenoisygenerator, () => new YeSampling(new SLS2novlp(N,T_min,T_max)), "SLS2novlpN" + N + "poisson" + m)
[…]

### 2014-06-13 code/java/src/snpe/bounds/BoundCalculator.java
renamed bound package to theory

@@ -1,38 +1,0 @@
-package snpe.bounds;
-
-/**
- *
- * @author Robby McKilliam
- */
-public interface BoundCalculator {
-
-    /**
-     * Return the clairvoyant CRB for the set parameters
-     */
-    double getBound();
-
-    /**
-     *
-     * Set the parameter for the geometrically distributed
-     * index differences.
-     * @param p the geometric parameter in range [0,1]
-     */
-    void setGeometicParameter(double p);
-
-    /**
-     * Set the data length
-     */
-    void setLength(int N);
-
-    /**
-     * Set the vector of indicies
-     */
-    void setS(double[] s);
-    
-    /** Set the true period of the signal */
-    void setPeriod(double T);
-    
-    /** Set the variance of the noise */
-    void setVariance(double var);
-
-}

### 2014-06-13 code/java/src/snpe/bounds/ClairvoyantCRB.java

@@ -1,48 +1,0 @@
-package snpe.bounds;
-
-import pubsim.lattices.Anstar.Anstar;
-import pubsim.VectorFunctions;
-
-/**
- * This is Sirdopulous's clairvoyant Cramer-Rao bound.  It is
- * just the bound given by linear regression under the assumption
- * that the vector of indicies, s, is known.
- * @author Robby McKilliam
- */
-public class ClairvoyantCRB implements BoundCalculator {
-    
-    protected double[] s, x;
-    
-    protected double T, var, mag2x;
-    
-    /** Set the vector of indicies */
-    public void setS(double[] s){
-        this.s = s;
-        x = new double[s.length];
-        Anstar.project(s, x);
-        mag2x = VectorFunctions.sum2(x);
-    }
-    
-    /** Set the true period of the signal */
-    public void setPeriod(double T){
-        this.T = T;
-    }
-    
-    /** Set the variance of the noise */
-    public void setVariance(double var){
-        this.var = var;
-    }
-    
-    /** Return the clairvoyant CRB for the set parameters */
-    public double getBound(){
-        return var/mag2x;
-    }
-
-    public void setGeometicParameter(double p) {
-    }
-
-    public void setLength(int N) {
-    }
-    
-        
-}

### 2014-06-13 code/java/src/snpe/bounds/ClairvoyantCRBRandomS.java

@@ -1,50 +1,0 @@
-package snpe.bounds;
-
-/**
- *
- * @author Robby McKilliam
- */
-public class ClairvoyantCRBRandomS extends ClairvoyantCRB{
-    
-    double p;
-    int N;
-    
-    /** 
-     * Set the parameter for the geometrically distributed
-     * index differences.
-     * @param p the geometric parameter in range [0,1]
-     */
-    @Override
-    public void setGeometicParameter(double p){
-        this.p = p;
-    }
-    
-    /** Set the vector of indicies */
-    @Override
-    public void setS(double[] s){
-        setLength(s.length);
-    }
-    
-    /** Set the data length */
-    @Override
-    public void setLength(int N){
-        this.N = N-1;
-    }
-    
-    
-    /** Return the clairvoyant CRB for the set parameters */
-    @Override
-    public double getBound(){
-        double ntn = N*N*N/3.0 + N*N/2.0 + N/6.0;
-        double nt1 = N*N/2.0 + N/2.0;
-        double ep = 1/p;
-        double vp = (1-p)/(p*p);
-        double ests = ntn*ep*ep + nt1*vp;
-        double est1 = nt1*ep;
-        
-        double extx = ests - est1*est1/N;
-        
-        return var/extx;
-    }
-
-}

### 2014-06-13 code/java/src/snpe/bounds/Hamersley.java

@@ -1,65 +1,0 @@
-package snpe.bounds;
-
-import Jama.Matrix;
-import pubsim.VectorFunctions;
-
-/**
- * Hamersley-Chapman-Robbins bound that applies
- * when the index differences are geometrically
- * distributed.
- * @author Robby McKilliam
- */
-public class Hamersley extends ClairvoyantCRB implements BoundCalculator{
-    
-    int N;
-    double p;
-    
-    /** Set the vector of indicies */
-    @Override
-    public void setS(double[] s){
-        setLength(s.length);
-    }
-    
-    /** Set the data length */
-    @Override
-    public void setLength(int N){
-        this.N = N-1;
-    }
-    
-    /** 
-     * Set the parameter for the geometrically distributed
-     * index differences.
-     * @param p the geometric parameter in range [0,1]
-     */
-    @Override
-    public void setGeometicParameter(double p){
-        this.p = p;
-    }
-    
-   /** Return the clairvoyant CRB for the set parameters */
-    @Override
-    public double getBound(){
-        double ntn = N*N*N/3.0 + N*N/2.0 + N/6.0;
-        double nt1 = N*N/2.0 + N/2.0;
-        double ep = 1/p;
-        double vp = (1-p)/(p*p);
-        double ests = ntn*ep*ep + nt1*vp;
-        double est1 = nt1*ep;
-        double estd = N*ep;
-        
-                //set the numerator matrix
-        Matrix num = new Matrix(2,2);
-        num.set(0,0, N/var); num.set(0,1, T/var);
-        num.set(1,0, T/var); num.set(1,1, Math.expm1(T*T/var));
-        
-        //set the numerator matrix
-        Matrix den = new Matrix(3,3);
-        den.set(0,0, ests/var); den.set(0,1, est1/var); den.set(0,2, T*estd/var);
-        den.set(1,0, est1/var); den.set(1,1, N/var); den.set(1,2, T/var);
-        den.set(2,1, T*estd/var); den.set(2,1, T/var); den.set(2,2, Math.expm1(T*T/var));
-        
-        return VectorFunctions.stableDet(num)/VectorFunctions.stableDet(den);
-        
-    }
-
-}

### 2014-06-13 code/java/src/snpe/bounds/HamersleyFixedS.java

@@ -1,75 +1,0 @@
-package snpe.bounds;
-
-import Jama.Matrix;
-import java.util.Iterator;
-import pubsim.lattices.util.ZnWithinSphere;
-import pubsim.VectorFunctions;
-
-/**
- * This is the Hamerley-Chapmin-Robbins bound that
- * includes the discrete parameter.  It is only tight when
- * sigma^2/T^2 is less than log(3).  This is generally in
- * the range that the estimators work so it's not a
- * terrible restriction.
- * @author Robby McKilliam
- */
-public class HamersleyFixedS extends ClairvoyantCRB{
-    
-    int N;
-    double sts, st1;
-    
-    /** Set the vector of indicies */
-    @Override
-    public void setS(double[] s){
-        this.s = s;
-        sts = VectorFunctions.sum2(s);
-        st1 = VectorFunctions.sum(s);
-        N = s.length;
-        
-    }
-    
-    /** Return the Hamerley-Chapmin-Robbins bound for the set parameters */
-    @Override
-    public double getBound(){
-        
-        Matrix num = new Matrix(2,2);
-        Matrix den = new Matrix(3,3);
-        
-        System.out.println(N);
-        System.out.println(2*Math.sqrt(sts));
-        
-        ZnWithinSphere deltas = new ZnWithinSphere(N, 2*Math.sqrt(sts));
-        
-        double B = Double.NEGATIVE_INFINITY;
-        Iterator itr = deltas.iterator();
-        while(itr.hasNext()){
-            
-            double[] delta = (double[]) itr.next();
-            
-            //System.out.println(VectorFunctions.print(delta));
-            
-            
-            double dt1 = VectorFunctions.sum(delta);
-            double dts = VectorFunctions.dot(delta, s);
-            double magd = VectorFunctions.magnitude(delta);
-            double dtd = VectorFunctions.sum2(delta);
-            
-            //set the numerator matrix
-            num.set(0,0, N/var); num.set(0,1, T*dt1/(var*magd));
-            num.set(1,0, T*dt1/(var*magd)); num.set(1,1, Math.expm1(T*T*dtd/var)/dtd);
-
-            //set the numerator matrix
-            den.set(0,0, sts/var); den.set(0,1, st1/var); den.set(0,2, T*dts/(var*magd));
-            den.set(1,0, st1/var); den.set(1,1, N/var); den.set(1,2, T*dt1/(var*magd));
[…]

### 2014-06-13 code/java/src/snpe/generators/DifferencesIID.java

@@ -26,8 +26,8 @@ public class DifferencesIID implements SpareIntegerSequence {
 
     @Override
     public int[] generate() {
-        s[0] = discreterv.getNoise();
-        for(int i = 1; i < N; i++) s[i] = s[i-1] + discreterv.getNoise();
+        s[0] = discreterv.noise();
+        for(int i = 1; i < N; i++) s[i] = s[i-1] + discreterv.noise();
         return s;
     }
     

### 2014-06-13 code/java/src/snpe/generators/SparseNoisyPeriodicSignal.java

@@ -45,7 +45,7 @@ public class SparseNoisyPeriodicSignal {
     public double[] generate() {
           int[] s = sparsenoise.generate();
           for(int i = 0; i< N; i++)
-              y[i] = T * s[i] + noise.getNoise() + phase;
+              y[i] = T * s[i] + noise.noise() + phase;
           return y;
     }
     

### 2014-06-13 code/java/src/snpe/theory/CLT.java

@@ -1,4 +1,4 @@
-package snpe.bounds;
+package snpe.theory;
 
 /**
  *

### 2014-06-13 code/java/src/snpe/theory/ClairvoyantCRB.java

@@ -1,0 +1,32 @@
+package snpe.theory;
+
+import pubsim.lattices.Anstar.Anstar;
+import pubsim.VectorFunctions;
+
+/**
+ * This is the clairvoyant Cramer-Rao bound.  It is
+ * just the bound given by linear regression under the assumption
+ * that the vector of indicies, s, is known.
+ * @author Robby McKilliam
+ */
+public class ClairvoyantCRB {
+    
+    public final double bound;
+    
+    public ClairvoyantCRB(double[] s, double var){
+        double[] x = new double[s.length];
+        Anstar.project(s, x);
+        double mag2x = VectorFunctions.sum2(x);
+        bound = var/mag2x;
+    }
+
+    /** @return the clairvoyant CRB for the set parameters */
+    public double bound(){
+        return bound;
+    }
+    
+    public static double bound(double[] s, double var){
+        return new ClairvoyantCRB(s, var).bound;
+    }
+        
+}

### 2014-06-13 code/java/src/snpe/theory/NormalisedLLSCLT.java

@@ -1,4 +1,4 @@
-package snpe.bounds;
+package snpe.theory;
 
 import pubsim.distributions.RealRandomVariable;
 import pubsim.distributions.circular.CircularRandomVariable;
@@ -24,7 +24,7 @@ public class NormalisedLLSCLT implements CLT{
     public NormalisedLLSCLT(RealRandomVariable noise, double discretemean, double T0){
 
         dmean = discretemean;
-        CircularRandomVariable wrped = noise.getWrapped();
+        CircularRandomVariable wrped = noise.wrapped();
         double h = wrped.pdf(-0.5);
 
         //Is is the wrapped variance or not?  Need to check Barry's

### 2014-06-13 code/java/src/snpe/theory/PeriodogramCLT.java

@@ -1,4 +1,4 @@
-package snpe.bounds;
+package snpe.theory;
 
 import pubsim.distributions.RealRandomVariable;
 

### 2014-06-13 code/java/test/snpe/NormalisedSamplingLLSTest.java

@@ -3,8 +3,8 @@ package snpe;
 import snpe.generators.SparseNoisyPeriodicSignal;
 import static org.junit.Assert.assertEquals;
 import org.junit.*;
-import pubsim.distributions.GaussianNoise;
-import pubsim.distributions.discrete.PoissonRandomVariable;
+import pubsim.distributions.Gaussian;
+import pubsim.distributions.discrete.Poisson;
 import snpe.generators.DifferencesIID;
 
 /**
@@ -47,8 +47,8 @@ public class NormalisedSamplingLLSTest {
         NormalisedSamplingLLS instance = new NormalisedSamplingLLS(n, Tmin, Tmax, 2*n);
 
         double noisestd = 0.001;
-        GaussianNoise noise = new pubsim.distributions.GaussianNoise(0.0,noisestd*noisestd);
-        PoissonRandomVariable drv = new PoissonRandomVariable(2);
+        Gaussian noise = new Gaussian(0.0,noisestd*noisestd);
+        Poisson drv = new Poisson(2);
         
         SparseNoisyPeriodicSignal sig = new SparseNoisyPeriodicSignal(n,T, phase, new DifferencesIID(n,drv),noise);
         double[] y = sig.generate();

### 2014-06-13 code/java/test/snpe/PeriodogramEstimatorTest.java

@@ -7,8 +7,8 @@ import org.junit.Before;
 import org.junit.BeforeClass;
 import org.junit.Test;
 import static org.junit.Assert.*;
-import pubsim.distributions.GaussianNoise;
-import pubsim.distributions.discrete.PoissonRandomVariable;
+import pubsim.distributions.Gaussian;
+import pubsim.distributions.discrete.Poisson;
 import snpe.generators.DifferencesIID;
 
 /**
@@ -51,8 +51,8 @@ public class PeriodogramEstimatorTest {
         PeriodogramEstimator instance = new PeriodogramEstimator(n, Tmin, Tmax, 2*n);
 
         double noisestd = 0.001;
-        GaussianNoise noise = new pubsim.distributions.GaussianNoise(0.0,noisestd*noisestd);
-        PoissonRandomVariable drv = new PoissonRandomVariable(2);
+        Gaussian noise = new Gaussian(0.0,noisestd*noisestd);
+        Poisson drv = new Poisson(2);
         
         SparseNoisyPeriodicSignal sig = new SparseNoisyPeriodicSignal(n,T, phase, new DifferencesIID(n,drv),noise);
         double[] y = sig.generate();

### 2014-06-13 code/java/test/snpe/QuantisedPeriodogramChirpZTest.java

@@ -7,8 +7,8 @@ import org.junit.Before;
 import org.junit.BeforeClass;
 import org.junit.Test;
 import static org.junit.Assert.*;
-import pubsim.distributions.GaussianNoise;
-import pubsim.distributions.discrete.PoissonRandomVariable;
+import pubsim.distributions.Gaussian;
+import pubsim.distributions.discrete.Poisson;
 import snpe.generators.DifferencesIID;
 
 /**
@@ -51,8 +51,8 @@ public class QuantisedPeriodogramChirpZTest {
         QuantisedPeriodogramChirpZ instance = new QuantisedPeriodogramChirpZ(n, Tmin, Tmax, 4/Tmin);
 
         double noisestd = 0.0001;
-        GaussianNoise noise = new pubsim.distributions.GaussianNoise(0.0,noisestd*noisestd);
-        PoissonRandomVariable drv = new PoissonRandomVariable(2);
+        Gaussian noise = new Gaussian(0.0,noisestd*noisestd);
+        Poisson drv = new Poisson(2);
         
         SparseNoisyPeriodicSignal sig = new SparseNoisyPeriodicSignal(n,T, phase, new DifferencesIID(n,drv),noise);
         double[] y = sig.generate();

### 2014-06-13 code/java/test/snpe/QuantisedPeriodogramFFTTest.java

@@ -8,8 +8,8 @@ import org.junit.BeforeClass;
 import org.junit.Test;
 import static org.junit.Assert.*;
 import pubsim.Complex;
-import pubsim.distributions.GaussianNoise;
-import pubsim.distributions.discrete.PoissonRandomVariable;
+import pubsim.distributions.Gaussian;
+import pubsim.distributions.discrete.Poisson;
 import snpe.generators.DifferencesIID;
 
 /**
@@ -93,8 +93,8 @@ public class QuantisedPeriodogramFFTTest {
         QuantisedPeriodogramFFT instance = new QuantisedPeriodogramFFT(n, Tmin, Tmax, 10);
 
         double noisestd = 0.001;
-        GaussianNoise noise = new pubsim.distributions.GaussianNoise(0.0,noisestd*noisestd);
-        PoissonRandomVariable drv = new PoissonRandomVariable(2);
+        Gaussian noise = new Gaussian(0.0,noisestd*noisestd);
+        Poisson drv = new Poisson(2);
         
         SparseNoisyPeriodicSignal sig = new SparseNoisyPeriodicSignal(n,T, phase, new DifferencesIID(n,drv),noise);
         double[] y = sig.generate();

### 2014-06-13 code/java/test/snpe/SparseNoisyPeriodicSignalTest.java

@@ -2,7 +2,7 @@ package snpe;
 
 import snpe.generators.SparseNoisyPeriodicSignal;
 import junit.framework.TestCase;
-import pubsim.distributions.GaussianNoise;
+import pubsim.distributions.Gaussian;
 import snpe.generators.FixedIntegerSequence;
 
 /**
@@ -26,7 +26,7 @@ public class SparseNoisyPeriodicSignalTest extends TestCase {
         double T = 1.0;
         double phase = 0.1;
         FixedIntegerSequence fs = new FixedIntegerSequence(s);
-        SparseNoisyPeriodicSignal sig = new SparseNoisyPeriodicSignal(N, T, phase, fs, new GaussianNoise(0, 0));
+        SparseNoisyPeriodicSignal sig = new SparseNoisyPeriodicSignal(N, T, phase, fs, new Gaussian(0, 0));
         double[] y = sig.generate();
         
         for (int i = 0; i < N; i++) assertTrue( Math.abs(y[i] - T*s[i] - phase) < 1e-9 );

### 2014-06-13 code/java/test/snpe/YeSamplingTest.java

@@ -7,8 +7,8 @@ import org.junit.Before;
 import org.junit.BeforeClass;
 import org.junit.Test;
 import static org.junit.Assert.*;
-import pubsim.distributions.GaussianNoise;
-import pubsim.distributions.discrete.PoissonRandomVariable;
+import pubsim.distributions.Gaussian;
+import pubsim.distributions.discrete.Poisson;
 import snpe.generators.DifferencesIID;
 
 /**
@@ -51,8 +51,8 @@ public class YeSamplingTest {
         YeSampling instance = new YeSampling(new PeriodogramEstimator(n, Tmin, Tmax));
 
         double noisestd = 0.001;
-        GaussianNoise noise = new pubsim.distributions.GaussianNoise(0.0,noisestd*noisestd);
-        PoissonRandomVariable drv = new PoissonRandomVariable(2);
+        Gaussian noise = new Gaussian(0.0,noisestd*noisestd);
+        Poisson drv = new Poisson(2);
         
         SparseNoisyPeriodicSignal sig = new SparseNoisyPeriodicSignal(n,T, phase, new DifferencesIID(n,drv),noise);
         double[] y = sig.generate();
@@ -79,8 +79,8 @@ public class YeSamplingTest {
         YeSampling instance = new YeSampling(new NormalisedSamplingLLS(n, Tmin, Tmax));
 
         double noisestd = 0.001;
-        GaussianNoise noise = new pubsim.distributions.GaussianNoise(0.0,noisestd*noisestd);
-        PoissonRandomVariable drv = new PoissonRandomVariable(2);
+        Gaussian noise = new Gaussian(0.0,noisestd*noisestd);
+        Poisson drv = new Poisson(2);
         
         SparseNoisyPeriodicSignal sig = new SparseNoisyPeriodicSignal(n,T, phase, new DifferencesIID(n,drv),noise);
         double[] y = sig.generate();

### 2014-06-13 code/java/test/snpe/ZnLLSTest.java

@@ -9,8 +9,8 @@ package snpe;
 
 import snpe.generators.SparseNoisyPeriodicSignal;
 import junit.framework.TestCase;
-import pubsim.distributions.GaussianNoise;
-import pubsim.distributions.discrete.PoissonRandomVariable;
+import pubsim.distributions.Gaussian;
+import pubsim.distributions.discrete.Poisson;
 import snpe.generators.DifferencesIID;
 
 /**
@@ -38,8 +38,8 @@ public class ZnLLSTest extends TestCase {
         ZnLLS instance = new ZnLLS(N,Tmin,Tmax);
         
         double noisestd = 0.0001;
-        GaussianNoise noise = new pubsim.distributions.GaussianNoise(0.0,noisestd*noisestd);
-        PoissonRandomVariable drv = new PoissonRandomVariable(2);
+        Gaussian noise = new Gaussian(0.0,noisestd*noisestd);
+        Poisson drv = new Poisson(2);
         
         SparseNoisyPeriodicSignal sig = new SparseNoisyPeriodicSignal(N,T, phase, new DifferencesIID(N,drv),noise);
         double[] y = sig.generate();
@@ -66,9 +66,9 @@ public class ZnLLSTest extends TestCase {
         ZnLLS instance = new ZnLLS(N,1/fmax,1/fmin);
 
         double noisestd = 0.00001;
-        GaussianNoise noise = new pubsim.distributions.GaussianNoise(0.0,noisestd*noisestd);
+        Gaussian noise = new Gaussian(0.0,noisestd*noisestd);
 
-        PoissonRandomVariable drv = new PoissonRandomVariable(2);
+        Poisson drv = new Poisson(2);
         
         SparseNoisyPeriodicSignal sig = new SparseNoisyPeriodicSignal(N,T, phase, new DifferencesIID(N,drv),noise);
         double[] y = sig.generate();

### 2014-06-13 response/response.tex
added new data for responses

--- 2014-06-11 response/response.tex
@@
 This paper proposes a new fast method for estimating the period of sparse and noisy
-observations of a point process. The new idea is to used the periodogram of quantized
+observations of a point process. The new idea is to use the periodogram of quantized
 observations that can be computed faster than the standard periodogram by using
@@
 Despite being simple, the results are novel and useful.
-The quantisated periodogram estimator described in this paper is as accurate as the most accurate estimators in the literature (the least squares and periodogram estimator),  but, for moderate and large data sets our estimator is substantially more computationally simple.
+The quantisated periodogram estimator described in this paper is as accurate as the most accurate estimators in the literature (the least squares and periodogram estimator).
+At the same time our estimator is substantially faster to compute.
 For these reasons, we firmly believe this will become the estimator of choice for periodic point processes.
-It enables period estimation for extremely large data sets in reasonable time.
-This is not possible with previous estimators that require a number of operations growing quadratically with data size.
-We have provided all the simulation code as supplementary material.
+For the first time, it enables highly accurate period estimation for extremely large data sets in reasonable time.
+This was not possible with previous estimators that require a number of operations growing quadratically with data size.
+%We have provided all the simulation code as supplementary material.
 This will enable simple implementation for future research.
@@
 The main reviewer concern is about the choice of the quantization parameter
-$q$ that should be explained more carefully. It is not suﬃcient to display two sets
-of curves obtained for two diﬀerent values of $q$ and to choose the best one without
+$q$ that should be explained more carefully. It is not sufficient to display two sets
+of curves obtained for two different values of $q$ and to choose the best one without
 explanations. For instance, it is not clear whether the value of $q$ chosen in this paper
@@
 These theoretical difficulties do not negate the immediate practical applicability of our estimator.
+We believe it would be unwise to halt dissemination of these practical results simply because a difficult theoretical problem has not yet been solved.
 We agree with the reviewer that the simulations presented are limited.
@@
 In the paper we use our simulations to motivate the choice $q = 5 f_{\text{max}}$ for the quantisation parameter.
-It is natural to wonder whether this choice also works in other scenarios, for example (as the reviewer suggests), for different methods of generating the integers $s_1,\dots,s_N$.
+It is natural to question whether this choice also works in other scenarios, for example (as the reviewer suggests), for different noise distributions, and different methods of generating the integers $s_1,\dots,s_N$.
+The results of such simulations are displayed in Figures~\ref{plot:poisson} and~\ref{plot:geomuniform} in this document.
+The simulations are setup as in the paper, but with the following modifications:
+
+\begin{enumerate}
+\item In Figure~\ref{plot:poisson} the noise is Gaussian distributed, $N=200$, and the integers are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Poisson distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right.
+\item In Figure~\ref{plot:geomuniform} the noise if uniformly distributed, $N=200$, and the integers are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right.
+\end{enumerate}
+
+Observe that, in both simulations, the choice $q=5 f_{\text{max}}$ is a good one.
+With this choice the accuracy of the quantised periodogram estimator is similar to the original periodgram estimator (computed without quantisation and FFT).
+%That the choice $q=5 f_{\text{max}}$ works in a wide range of scenarios should not come as a
+
+\begin{figure*}[tp] 
+  \centering 
+  \begin{tikzpicture} 
+    \selectcolormodel{gray} 
+    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.24)}},xlabel style={at={(0.5,0.07)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=7, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
+      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200poisson1}; 
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200poisson1};
+            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5poisson1};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0poisson1};
+      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN200poisson1};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN200poisson1}; 
+ %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200poisson1};
+      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200poisson1};
+     \legend{Periodogram theory, Least squares theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
+     %\node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=200$};
+   \end{axis} 
+ \begin{axis}[name=plot2,at={(10.5,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.23)}},xlabel style={at={(0.5,0.07)}}]
+      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200poisson10}; 
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200poisson10};
+            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5poisson10};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0poisson10};
+      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN200poisson10};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN200poisson10}; 
+  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200poisson10};
+       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200poisson10};
+     %\node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=200$}; 
+   \end{axis} 
+  \end{tikzpicture}  
+  \caption{Mean square period error versus noise variance $\sigma^2$ with Gaussain noise and $N=200$ observations.
+The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Poisson distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$.}\label{plot:poisson}
+\end{figure*}
+
+\begin{figure*}[tp]  
+  \centering
+  \begin{tikzpicture}  
+    \selectcolormodel{gray} 
+    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.24)}},xlabel style={at={(0.5,0.07)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=7, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
+\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200uniformgeom1}; 
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200uniformgeom1};
+            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5uniformgeom1};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0uniformgeom1};
+      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN200uniformgeom1};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN200uniformgeom1}; 
+ %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200uniformgeom1};
+      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200uniformgeom1};
+     \legend{Periodogram theory, Least squares theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
+     %\node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=200$};
+   \end{axis}   \begin{axis}[name=plot2,at={(4.7,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.23)}},xlabel style={at={(0.5,0.07)}}]
+\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200uniformgeom10}; 
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200uniformgeom10};
+            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5uniformgeom10};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0uniformgeom10};
+      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN200uniformgeom10};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN200uniformgeom10}; 
+  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200uniformgeom10};
+       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200uniformgeom10};
+     %\node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=200$}; 
+   \end{axis} 
+  \end{tikzpicture}  
+  \caption{Mean square period error versus noise variance $\sigma^2$ with uniform noise and $N=200$ observations.
+The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$.
+The least squares estimator is the most accurate in this simulation.
+The is predicted by the asymptotic theory~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}.}\label{plot:geomuniform}
+\end{figure*}
 
@@
 compare their results in term of accuracy and of computational complexity to
-the results of [24]. Including the Cramér-Rao bounds for the estimation of T0
+the results of [24]. Including the Cramer-Rao bounds for the estimation of T0
 (as in [24]) would also be appreciated.
@@
 \textbf{Response:}
-Section VI in~\cite{726812} is not the conclusion.
+Section VI in the paper~\cite{726812} entitled ``On periodic pulse interval analysis with outliers and missing observations'' is not the conclusion.
 It is the section entitled ``Numerical Results''.
-It is clear from Figure 1 to 4 in~\cite{726812} that the periogram estimator implemented by Casey and Sadler is inferior to their MEA estimator.
+In Figures 1 to 4 in~\cite{726812} the periogram estimator implemented by Casey and Sadler is seen to be inferior to their MEA estimator.
 Casey and Sadler do use an FFT to compute the periodgram, but fail to implement an optimisation proceedure.
 We used the Newton-Raphson proceedure in the old manuscript, but have used Brent's method~\cite{Brent_opt_no_derivs_1973} in the new manuscript as dicussed in our response to comment~\ref{com:rev1:nr} below.
-This leads Casey and Sadler to the conclusion that very large FFTs are required and that the periodgram estimator is not particularly accurate.
+Casey and Sadler arrive at the conclusion that very large FFTs are required and that the periodgram estimator is not particularly accurate.
 In the conclusion of~\cite{726812} it is stated that:
@@
 The periodgram estimator, computed by quanitisation and FFT and combined with an appropriate optimisation proceedure, is both compuationally efficient and highly accurate.
-Our paper shows that (for moderately large $N$) this is the most computationally efficient proceedure currently available for this problem.
-Simultaneously, it is as accurate at the most accuraate estimators available, namely the least squares estimator, and the periodgram estimator computed directly without FFT.
+For moderately large $N$ this is the most computationally efficient proceedure currently available for this problem.
+Simultaneously, it is as accurate at the most accurate estimators available, namely the least squares estimator, and the periodgram estimator computed directly without FFT.
 
-BLERG: Cramer-Rao bound not available for randomly chosen $s_1,\dots,s_N$.
+In the simulations we regenerate the integers $s_1,\dots,s_N$ on each Monte-Carlo trial.
+To the authors knowledge, a Cramer-Rao bound is not known in this case.
+The \emph{clairvoyant} Cramer-Rao bound from~\cite{Sidiropoulos2005} (also in \cite{726812,Clarkson2007}) supposes a single fixed realisation for $s_1,\dots,s_N$.
+This bound is therefore not valid for the simulations we present in Figure~\ref{plot:multipleN} in Section~\ref{sec:simulations} of the paper.
+
+In Figure~\ref{plot:crb} in this document we perform a simulation with integers $s_1,\dots,s_N$.
+The integers are generated once such $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed.
+This single fixed sequence is used for all of the $1000$ Monte-Carlo trials.
+Observe that for small noise variance the \emph{clairvoyant} Cramer-Rao bound is similar to the asymptotic variance of the least square and periodogram estimator predicted in~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}.
+As the noise variance increases the \emph{clairvoyant} Cramer-Rao bound and the asymptotic variances diverge.
+As expected, the behaviour of the estimators themselves follows the asymptotic theory, rather than the \emph{clairvoyant} Cramer-Rao bound.
+
+%We prefer to show to plot the MSE of the estimatorsThe least squares estimator is the maximum likelihood estimator, and so, the dashed curve serves a similar type of purpose as the CRB often does in these types of simulations.
+Namely, to indicate whether an estimator is performing similarly to what might be considered `optimal', i.e.
+close the what might theoretically be expected from the maximum likelihood estimator.
+
+\begin{figure*}[tp]  
+  \centering
+  \begin{tikzpicture}  
+    \selectcolormodel{gray} 
+    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.24)}},xlabel style={at={(0.5,0.07)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=7, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
+\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200fixed1}; 
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200fixed1};
+      \addplot[mark=none,color=black,dotted] table {code/data/CRBN200fixed1};
+            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5fixed1};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0fixed1};
+      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN200fixed1};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN200fixed1}; 
+ %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200fixed1};
+      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200fixed1};
+     \legend{Periodogram theory, Least squares theory, CRB, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
+     %\node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=200$};
+   \end{axis}   \begin{axis}[name=plot2,at={(4.7,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.23)}},xlabel style={at={(0.5,0.07)}}]
+\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200fixed10}; 
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200fixed10};
+      \addplot[mark=none,color=black,dotted] table {code/data/CRBN200fixed10};
+            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5fixed10};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0fixed10};
+      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN200fixed10};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN200fixed10}; 
+  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200fixed10};
+       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200fixed10};
+     %\node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=200$}; 
+   \end{axis} 
+  \end{tikzpicture}   
+  \caption{Mean square period error versus noise variance $\sigma^2$ with Gaussian noise and $N=200$ observations.
+A singled fixed set of integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$.
+This single fixed sequence is used for all of the $1000$ Monte-Carlo trials.
+In this case the  \emph{clairvoyant} Cramer-Rao bound from~\cite{Sidiropoulos2005} (also in \cite{726812,Clarkson2007}) exists and is displayed by the dotted line.
+In this simulation the \emph{clairvoyant} Cramer-Rao bound is equal to the variance of the maximum likelihood (i.e.
+least squares) estimator in the case that the integers $s_1,\dots,s_N$ are known.}\label{plot:crb}
+\end{figure*}
 

### 2014-06-13 code/java/src/snpe/theory/ClairvoyantCRB.java

@@ -1,7 +1,8 @@
 package snpe.theory;
 
 import pubsim.lattices.Anstar.Anstar;
-import pubsim.VectorFunctions;
+import static pubsim.VectorFunctions.intArrayToDoubleArray;
+import static pubsim.VectorFunctions.sum2;
 
 /**
  * This is the clairvoyant Cramer-Rao bound.  It is
@@ -13,10 +14,14 @@ public class ClairvoyantCRB {
     
     public final double bound;
     
+    public ClairvoyantCRB(int[] s, double var){
+        this(intArrayToDoubleArray(s), var);
+    }
+    
     public ClairvoyantCRB(double[] s, double var){
         double[] x = new double[s.length];
-        Anstar.project(s, x);
-        double mag2x = VectorFunctions.sum2(x);
+        Anstar.project(s, x); //projects into the zero mean plane (i.e. the plane orthogonal to (1,...1)
+        double mag2x = sum2(x);
         bound = var/mag2x;
     }
 
@@ -26,6 +31,10 @@ public class ClairvoyantCRB {
     }
     
     public static double bound(double[] s, double var){
+        return new ClairvoyantCRB(s, var).bound;
+    }
+    
+    public static double bound(int[] s, double var){
         return new ClairvoyantCRB(s, var).bound;
     }
         

### 2014-06-13 code/montecarlo.scala

@@ -1,12 +1,12 @@
 /**
  * Run simulations of various sparse noise period estimators.
  */
-import snpe.generators.DifferencesIID
 import pubsim.distributions.discrete.DiscreteRandomVariable
-import pubsim.distributions.discrete.GeometricRandomVariable
-import pubsim.distributions.GaussianNoise
+import pubsim.distributions.discrete.Geometric
+import pubsim.distributions.Gaussian
 import pubsim.distributions.RealRandomVariable
 import snpe.generators.SparseNoisyPeriodicSignal
+import snpe.generators.DifferencesIID
 import snpe.PRIEstimator
 import snpe.YeSampling
 import snpe.PeriodogramEstimator
@@ -14,8 +14,8 @@ import snpe.NormalisedSamplingLLS
 import snpe.QuantisedPeriodogramFFT
 import snpe.SLS2novlp
 import snpe.SLS2new
-import snpe.bounds.PeriodogramCLT
-import snpe.bounds.NormalisedLLSCLT
+import snpe.theory.PeriodogramCLT
+import snpe.theory.NormalisedLLSCLT
 
 val iters = 100
 val T_0 = scala.math.Pi/3 //true period
@@ -24,14 +24,14 @@ val T_max = 1.5 //maximum period
 val f_min = 1.0/T_max //minimum frequency
 val f_max = 1.0/T_min //maximum frequency
 val theta_0 = 0.2 //true phase
-def noisedist(v : Double) = new GaussianNoise(0,v) //the noise distribution we use with variance v
+def noisedist(v : Double) = new Gaussian(0,v) //the noise distribution we use with variance v
 
 val starttime = (new java.util.Date).getTime
 
 for( N <- List(30,1200) ) {  
   //for discrete distributions with mean 1 and 10
   for( m <- List( 1, 10 ) ) {
-    def discretedist() =  new GeometricRandomVariable.StartingAtZero(1.0/(m+1.0))
+    def discretedist() =  new Geometric.StartingAtZero(1.0/(m+1.0))
     def sparsenoisygenerator(v : Double) = new SparseNoisyPeriodicSignal(N,T_0,theta_0,new DifferencesIID(N,discretedist),noisedist(v))
     val varbs = -38.0 to -5.0 by 1.5 //variances we use in dB
[…]

### 2014-06-13 response/code/montecarlo.scala

@@ -1,13 +1,15 @@
 /**
  * Run simulations of various sparse noise period estimators.
  */
-import snpe.generators.DifferencesIID
 import pubsim.distributions.discrete.DiscreteRandomVariable
-import pubsim.distributions.discrete.GeometricRandomVariable
-import pubsim.distributions.discrete.PoissonRandomVariable
-import pubsim.distributions.GaussianNoise
+import pubsim.distributions.discrete.Geometric
+import pubsim.distributions.discrete.Poisson
+import pubsim.distributions.Gaussian
+import pubsim.distributions.Uniform
 import pubsim.distributions.RealRandomVariable
 import snpe.generators.SparseNoisyPeriodicSignal
+import snpe.generators.FixedIntegerSequence
+import snpe.generators.DifferencesIID
 import snpe.PRIEstimator
 import snpe.YeSampling
 import snpe.PeriodogramEstimator
@@ -15,8 +17,9 @@ import snpe.NormalisedSamplingLLS
 import snpe.QuantisedPeriodogramFFT
 import snpe.SLS2novlp
 import snpe.SLS2new
-import snpe.bounds.PeriodogramCLT
-import snpe.bounds.NormalisedLLSCLT
+import snpe.theory.PeriodogramCLT
+import snpe.theory.NormalisedLLSCLT
+import snpe.theory.ClairvoyantCRB
 
 val iters = 100
 val T_0 = scala.math.Pi/3 //true period
@@ -25,17 +28,19 @@ val T_max = 1.5 //maximum period
 val f_min = 1.0/T_max //minimum frequency
 val f_max = 1.0/T_min //maximum frequency
 val theta_0 = 0.2 //true phase
-def noisedist(v : Double) = new GaussianNoise(0,v) //the noise distribution we use with variance v
 
 val starttime = (new java.util.Date).getTime
 
-//trials with Poisson discrete random variables
-for( N <- List(30,500) ) {  
+/////// Simulations with Poisson discrete random variables and Gaussian noise
+/*{
+ val N = 200
   //for discrete distributions with mean 1 and 10
   for( m <- List( 1, 10 ) ) {
-    def discretedist() = new PoissonRandomVariable(m);
+    def discretedist() = new Poisson(m);
[…]

### 2014-06-13 response/response.tex
added new suff

@@
       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200poisson1};
-     \legend{Periodogram theory, Least squares theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
+     \legend{Periodogram theory, LS theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, LS, SLS2-new}
      %\node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=200$};
@@
       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200uniformgeom1};
-     \legend{Periodogram theory, Least squares theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
+     \legend{Periodogram theory, LS theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, LS, SLS2-new}
      %\node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=200$};
@@
 
+Observe that in the right hand plot the quantised periodogram estimator with coarse quantisation $q = 1.5 f_{\text{max}}$ actually perform well for sufficently small noise variance.
+It appears that the fixed integers $s_1,\dots,s_N$ generated in this simulation favour the quantised periodogram estimator.
+This is one motivation for performing our simulations with integer $s_1,\dots,s_N$ generated randomly for each Monte-Carlo trial, rather than choosing a single fixed set of integers $s_1,\dots,s_N$.
+It avoid the scenario where a particular favourable (or unfavourable) set $s_1,\dots,s_N$ is chosen by chance.
+
+We have not implemented and do not present simulations results with any of the MEA estimators from the papers by Casey and Sadler~\cite{536682,726812}.
+The MEA estimators were supersceeded by the Separable Line Search (SLS) estimators of Sidiropoulos, Swami and Sadler~\cite{Sidiropoulos2005} in 2005.
+Comparison between the MEA estimator and the SLS are presented in~\cite{Sidiropoulos2005} where the superiority of the SLS estimators is shown.
+An enhanced SLS estimator was developed in~\cite{Clarkson2007} and this is the estimator we show simulation results for in our paper and this document.
+Even this enhanced SLS estimator performs poorly compared to the least squares and periodogram estimators.
+The MEA estimators are no longer state-of-the-art for this problem.
+
 %We prefer to show to plot the MSE of the estimatorsThe least squares estimator is the maximum likelihood estimator, and so, the dashed curve serves a similar type of purpose as the CRB often does in these types of simulations.
@@
     \selectcolormodel{gray}
-    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.24)}},xlabel style={at={(0.5,0.07)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=7, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
-\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200fixed1}; 
-      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200fixed1};
-      \addplot[mark=none,color=black,dotted] table {code/data/CRBN200fixed1};
-            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5fixed1};
-      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0fixed1};
-      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN200fixed1};
-       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN200fixed1}; 
- %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200fixed1};
-      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200fixed1};
-     \legend{Periodogram theory, Least squares theory, CRB, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
+    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.24)}},xlabel style={at={(0.5,0.07)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=8, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
+\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN500fixed1}; 
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN500fixed1};
+      \addplot[mark=none,color=black,dotted] table {code/data/CRBN500fixed1};
+            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN500q1.5fixed1};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN500q5.0fixed1};
+      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN500fixed1};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN500fixed1}; 
+ %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN500fixed1};
+      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN500fixed1};
+     \legend{Periodogram theory, LS theory, CRB, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, LS, SLS2-new}
      %\node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=200$};
-   \end{axis}   \begin{axis}[name=plot2,at={(4.7,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.23)}},xlabel style={at={(0.5,0.07)}}]
-\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200fixed10}; 
-      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200fixed10};
-      \addplot[mark=none,color=black,dotted] table {code/data/CRBN200fixed10};
-            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5fixed10};
-      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0fixed10};
-      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN200fixed10};
-       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN200fixed10}; 
-  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200fixed10};
-       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200fixed10};
+   \end{axis}   \begin{axis}[name=plot2,at={(9.5,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.23)}},xlabel style={at={(0.5,0.07)}}]
+\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN500fixed10}; 
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN500fixed10};
+      \addplot[mark=none,color=black,dotted] table {code/data/CRBN500fixed10};
+            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN500q1.5fixed10};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN500q5.0fixed10};
+      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN500fixed10};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN500fixed10}; 
+  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN500fixed10};
+       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN500fixed10};
      %\node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=200$}; 
@@
   \end{tikzpicture}
-  \caption{Mean square period error versus noise variance $\sigma^2$ with Gaussian noise and $N=200$ observations.
+  \caption{Mean square period error versus noise variance $\sigma^2$ with Gaussian noise and $N=500$ observations.
 A singled fixed set of integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$.
-This single fixed sequence is used for all of the $1000$ Monte-Carlo trials.
+This single fixed sequence is used for all of the $500$ Monte-Carlo trials.
 In this case the \emph{clairvoyant} Cramer-Rao bound from~\cite{Sidiropoulos2005} (also in \cite{726812,Clarkson2007}) exists and is displayed by the dotted line.

### 2014-06-13 response/code/montecarlo.scala

@@ -32,7 +32,7 @@ val theta_0 = 0.2 //true phase
 val starttime = (new java.util.Date).getTime
 
 /////// Simulations with Poisson discrete random variables and Gaussian noise
-/*{
+{
  val N = 200
   //for discrete distributions with mean 1 and 10
   for( m <- List( 1, 10 ) ) {
@@ -52,10 +52,10 @@ val starttime = (new java.util.Date).getTime
     runCLT(N, vardbsfiner, () => discretedist, (v : Double, dm : Double) => new PeriodogramCLT(noisedist(v/T_0/T_0), dm, T_0), "PeriodogramCLTN" + N + "poisson" + m)
     runCLT(N, vardbsfiner, () => discretedist, (v : Double, dm : Double) => new NormalisedLLSCLT(noisedist(v/T_0/T_0), dm, T_0), "NormalisedLLSCLTN" + N + "poisson" + m)
   }
-}*/
+}
 
 /////// Simulations with Geometric discrete random variables and uniform noise
-/*{
+{
 val N = 200
   //for discrete distributions with mean 1 and 10
   for( m <- List( 1, 10 ) ) {
@@ -76,11 +76,11 @@ val N = 200
     runCLT(N, vardbsfiner, () => discretedist, (v : Double, dm : Double) => new PeriodogramCLT(noisedist(v/T_0/T_0), dm, T_0), "PeriodogramCLTN" + N + "uniformgeom" + m)
     runCLT(N, vardbsfiner, () => discretedist, (v : Double, dm : Double) => new NormalisedLLSCLT(noisedist(v/T_0/T_0), dm, T_0), "NormalisedLLSCLTN" + N + "uniformgeom" + m)
   }
-}*/
+}
 
 /////// Simulations with fixed Geometric discrete random variables and Gaussian noise
 {
-val N = 500
+  val N = 500
   //for discrete distributions with mean 1 and 10
   for( m <- List( 1, 10 ) ) {
     val discretedist = new Geometric.StartingAtZero(1.0/(m+1.0))
@@ -88,7 +88,7 @@ val N = 500
     def noisedist(v : Double) = new Gaussian(0,v) //the noise distribution we use with variance v
     def sparsenoisygenerator(v : Double) = new SparseNoisyPeriodicSignal(N,T_0,theta_0,fixedseq,noisedist(v))
     val twelthdB = scala.math.log10(1.0/12)*10
[…]

### 2014-06-13 response/response.tex
more edits

@@
 Comparison between the MEA estimator and the SLS are presented in~\cite{Sidiropoulos2005} where the superiority of the SLS estimators is shown.
-An enhanced SLS estimator was developed in~\cite{Clarkson2007} and this is the estimator we show simulation results for in our paper and this document.
+An enhanced SLS estimator (SLS2-new) was developed in~\cite{Clarkson2007} and this is the estimator we show simulations results for.
 Even this enhanced SLS estimator performs poorly compared to the least squares and periodogram estimators.
@@
 \textbf{Response:}
-We are only aware of a single Netwon-Raphson method that proceeds by the iteration
+We reponse to each of the questions of the reviewer in turn:
+
+\begin{enumerate}
+\item \textbf{For instance, what is the exact NR scheme used in this paper (there are several options)?}  The Netwon-Raphson method used proceeds by the iteration
 \[
@@
 to find a stationary point of the function $f$ where $f^\prime$ and $f^{\prime\prime}$ are the first and second derivatives of $f$.
-As was stated Section~\zref{sec:simulations}: ``Both the periodogram and quanitised periodogram estimators use the Newton-Raphson method.''  This was the case both for the simulations presented in Figure~\zref{plot:multipleN} and for the computational benchmarks presented in Figure~\zref{plot:benchmark}.
-
 To avoid the potential ambguity regarding the definition of the Newton-Raphson method we have, in the new manuscipt, replaced the Newton-Raphson method with the optimisation proceedure due to Brent~\cite[Ch.~5]{Brent_opt_no_derivs_1973}.
 This optimiation method is likely to be more familiar with readers because it is used internally by the \texttt{fminbnd} function in the popular MATLAB environment.
+
+BLERG add some text to the paper.
+
+\item \text{How have the authors selected the total number of iterations?} In our implementation iterations are performed until the absolute difference between consequentive iterations is less than $10^{-7}$.
+
+BLERG: Something like the following should be added:
+
+``Brent's method is iterative.
+During the method a number of evaluations of the periodogram $I_y$ are made.
+The number of calls depends on the grid width $D$, but not on $N$.
+Each call requires $O(N)$ operations, and so, the total number of operations required by Brent's method is $O(N)$.
+In practice we find that that computational cost of Brent's method is far smaller than the cost of either that FFT or chirp-z transform.''
+
+\end{enumerate}
+
+We are only aware of a single   As was stated Section~\zref{sec:simulations}: ``Both the periodogram and quanitised periodogram estimators use the Newton-Raphson method.''  This was the case both for the simulations presented in Figure~\zref{plot:multipleN} and for the computational benchmarks presented in Figure~\zref{plot:benchmark}.
 
@@
 \textbf{Response:}
-We have not done this.
+Respectfully, we have not done this.
 The parameter $M$ is reserved for the length of the FFT used for the single Fourier transform method.
-We feel it will lead to confusion if $M$ is also used for the length of the transforms required for the chirp-z transform.
+We feel it is confusing if $M$ is also used for the length of the transforms required by the chirp-z transform.
 
@@
 \textbf{Response:}
-We have connected the data points for the $N=1200$ case by gray lines.  We believe this make the plot clear.  We are happy to make further adjustments to this plot should the reviewer require them.
+We have connected the data points for the $N=1200$ case by gray lines.  We believe this makes the plot clear.  We are happy to make further adjustments to this plot should the reviewer require them.
 
 \item\textbf{Comment:}
-Typos: p. 2, “anomolies" should be “anomalies", p. 7, “quanitised" should be
-“quantized".
+Typos: p. 2, ``anomolies'' should be ``anomalies'', p. 7, ``quanitised'' should be
+``quantized''.
 \\

### 2014-06-13 code/benchmark.scala

@@ -2,10 +2,11 @@
 * Run benchmarks.
 */
 
-import pubsim.distributions.discrete.GeometricRandomVariable
-import pubsim.distributions.GaussianNoise
+import pubsim.distributions.discrete.Geometric
+import pubsim.distributions.Gaussian
 import pubsim.distributions.RealRandomVariable
-import snpe.SparseNoisyPeriodicSignal
+import snpe.generators.SparseNoisyPeriodicSignal
+import snpe.generators.DifferencesIID
 import snpe.PRIEstimator
 import snpe.YeSampling
 import snpe.PeriodogramEstimator
@@ -22,8 +23,8 @@ val f_min = 1.0/T_max //minimum frequency
 val f_max = 1.0/T_min //maximum frequency
 val theta_0 = 0.2 //true phase
 val p = 1.0/3.0 //parameter of geometric random variable for generating sparse noise
-def discretedist = new GeometricRandomVariable(p) //the discrete distribution for generating different sN - sN-1
-def noisedist = new GaussianNoise(0,0.01) //the noise distribution we use with variance v
+def discretedist = new Geometric(p) //the discrete distribution for generating different sN - sN-1
+def noisedist = new Gaussian(0,0.01) //the noise distribution we use with variance v
 
 //time used for benchmark warmup in nanoseconds.  Benchmark runtime is approximately twice this.
 val MIN_BENCH_DURATION : Long = 5000000000L; // (10 secs)  
@@ -33,12 +34,12 @@ val Ns = List(10,14,20,30,40,50,60,80,100,120,150,200,250,300,400,600,900,1200,1
 //val Ls = List(1000)
 //val Ns = (1 to 40).map(i => (5*scala.math.pow(1.1,i)).toInt).distinct //number of symbols
 
-//runbench(Ns, (N : Int) => new YeSampling(new PeriodogramEstimator(N,T_min,T_max)), "Periodogram")
+runbench(Ns, (N : Int) => new YeSampling(new PeriodogramEstimator(N,T_min,T_max)), "Periodogram")
 runbench(Ns, (N : Int) => new YeSampling(new NormalisedSamplingLLS(N,T_min,T_max),6.0), "LeastSquares")
 //runbench(Ns, (N : Int) => new YeSampling(new SLS2novlp(N,T_min,T_max)), "SLS2novlp")
[…]

### 2014-06-13 code/montecarlo.scala

@@ -17,7 +17,7 @@ import snpe.SLS2new
 import snpe.theory.PeriodogramCLT
 import snpe.theory.NormalisedLLSCLT
 
-val iters = 100
+val iters = 3000
 val T_0 = scala.math.Pi/3 //true period
 val T_min = 0.8 //minimum period
 val T_max = 1.5 //maximum period

### 2014-06-13 code/montecarlo.scala
minor

@@ -107,7 +107,7 @@ def runCLT(N : Int, varbs : Seq[Double], discretedist : () => DiscreteRandomVari
   //now write all the data to a file
   val file = new java.io.FileWriter("data/" + name) //list of files to write to
   vars.foreach{ v =>
-    val mse = cltf(v, discretedist().getMean).periodVar(N)
+    val mse = cltf(v, discretedist().mean).periodVar(N)
     file.write(v.toString.replace('E', 'e') + "\t" + mse.toString.replace('E', 'e')  + "\n")
   }
   file.close //close all the files we wrote to 

### 2014-06-13 response/response.tex
added minor

@@
 \begin{enumerate}
-\item \textbf{For instance, what is the exact NR scheme used in this paper (there are several options)?}  The Netwon-Raphson method used proceeds by the iteration
+\item \textbf{For instance, what is the exact NR scheme used in this paper (there are several options)?}  The Netwon-Raphson method that was used proceeds by the iteration
 \[
@@
 
-\item \text{How have the authors selected the total number of iterations?} In our implementation iterations are performed until the absolute difference between consequentive iterations is less than $10^{-7}$.
+\item \textbf{How have the authors selected the total number of iterations?} In our implementation iterations are performed until the absolute difference between consequentive iterations is less than $10^{-7}$.
 
@@
 In practice we find that that computational cost of Brent's method is far smaller than the cost of either that FFT or chirp-z transform.''
+
+\item \textbf{Is the quantized periodogram
+just used to initialize a Newton-Raphson procedure? Do the MSEs displayed
+in Fig. 3 correspond to the estimator obtained after applying the NR algorithm?}
+
+\item \textbf{Also, do the results displayed in Fig. 2 correspond to the total time
+required to compute the estimator (including the NR procedure) or just to the
+initialization of the NR procedure?}
 
@@
 \\
-\textbf{Response:}
+\textbf{Response:} The same identifiability requirements are commonly imposed in the literature, see for example, Sidiropoulos~\cite[eq.
+(2)]{Sidiropoulos2005}.
+Suppose that $T_{\text{max}} \geq 2T_{\text{min}}$ and that the true period $T_0$ happened to be contained in the nonempty interval $(2T_{\text{min}}, T_0)$.
+Put $T^\prime = T_0/2 \in (T_{\text{min}}, T_{\text{max}})$ and observe that
+\[
+y_n = T_0 s_n + \theta_0 + w_n = 2 T^\prime s_n + \theta_0 + w_n = T^\prime r_n + \theta_0 + w_n 
+\]
+where $r_n = 2 s_n \in \ints$.
+Since the integers are not known both $T_0$ and $T^\prime$ would describe the observations equally well.
+This problem is averted only if we suppose that $T_0 \in (T_{\text{min}}, T_{\text{max}})$ and that  $T_{\text{max}} < 2T_{\text{min}}$.
+
+%Perhaps it is possible to increase the range of identifiability if further assumptions are made regarding the integers $s_1,\dots,s_N$.  We don't consider such propsitions in this paper.
+
+% To see this another way, define least squares objective function $LS : \reals \times \reals \times \ints^N \mapsto \reals$ as
+% \[
+% LS(T, \theta, z_1,\dots,z_N) = \sum_{n=1}^N (y_n - T z_n - \theta)^2.
+% \]
+% In the case that the noise variables $w_1,\dots,w_n$ are i.i.d. and normally distributed the minimiser of this objective function with respect to $T$ is the maximum likelihood estimator.  That is,
+% \[
+% \hat{T} = \arg\min_{T \in (T_{\text{min}}, T_{\text{max}})} \min_{\theta \in \reals} \min_{z_1,\dots,z_N \in \ints} LS(T, \theta, z_1,\dots,z_N)
+% \]
+% is the maximum likelihood period estimator.
+
+In the paper we have modifed the s BLERG
 
@@
 \textbf{Response:}
-BLERG
+Please see our response to comment~\ref{com:rev1:nr} of reviewer 1.
 

### 2014-06-14 response/response.tex
started added more edits.  Getting there

@@
 \textbf{Response:}
-We reponse to each of the questions of the reviewer in turn:
+We repond to each of the questions of the reviewer in turn:
 
@@
 We have removed reference [29] from the original manuscript because the material in that paper is also covered in [27].
-We feel that the all 37 references in the new version of the paper are highly relevant.
+We have also removed references [33] and [34] from the original manuscript.
+We feel that the all 35 references in the new version of the paper are highly relevant.
 The paper would be weakened without them.
@@
 
-In the paper we have modifed the s BLERG
+We have included reference~\cite{Sidiropoulos2005} and simplified the text at the start of Section~\zref{sec:peri-estim} to read:
+
+``To ensure identifiability it is necessary to assume, as in~\cite{Sidiropoulos2005}, that $T_0$ lies in some interval $(T_{\text{min}}, T_{\text{max}})$ where $T_{\text{max}} < 2T_{\text{min}}$.''
 

### 2014-06-14 paper.tex
more edits
added minor
started added more edits.  Getting there

@@
 Solutions have applications to synchronisation for telecommunications devices~\cite{Fogel1988,Fogel1989_bit_synch_zero_crossings,Sidiropoulos2005,5621928}, estimation of the pulse repetition interval in electronic support~\cite{EltonGray_puilse_train_rader_1994,Gray_more_pri_1994,Clarkson_thesis,clarkson_estimate_period_pulse_train_1996,Hauochan_pri_2012}, and beat and tempo estimation in music~\cite{dixon_beat_extraction_2001}.
-The problem is closely related to spike interval estimation in neurology~\cite{Arnett_neuro_pri_1976,Brillinger_spike_trains_1988,Rossoni200630} and a similar model has been proposed for the detection of anomalies, such as denial-of-service attacks, in internet traffic~\cite{He_detecting_periodic_patterns_in_internet_2006,5585849,5947313}.
+The problem is related to spike interval estimation in neurology~\cite{Arnett_neuro_pri_1976,Brillinger_spike_trains_1988,Rossoni200630} and a similar model has been proposed for the detection of anomalies, such as denial-of-service attacks, in internet traffic~\cite{He_detecting_periodic_patterns_in_internet_2006,5585849,5947313}.
 The problem is also connected with classical problems in number theory such as Diophantine approximation, lattice reduction, and the study of prime numbers~\cite{Cassels_geom_numbers_1997,490557,Clarkson_thesis,Lenstra1982,Wubben_2011,536682,726812,CaseySadler_primes_2013}.
@@
 %Before describing the periodogram estimator we state some required assumptions.  
-Observe that
-\[
-y_n = T_0 s_n + \theta_0 + w_n = \frac{T_0}{2} r_n + \theta_0 + w_n
-\]
-where $r_n = 2s_n \in \ints$.
-Thus, to ensure identifiability, it is necessary to assume that $T_0$ lies in some interval $(T_{\text{min}}, T_{\text{max}})$ where $T_{\text{max}} < 2T_{\text{min}}$.%  We also assume that $y_1,\dots,y_N$ are in ascending order.
+%Observe that
+%\[
+%y_n = T_0 s_n + \theta_0 + w_n = \frac{T_0}{2} r_n + \theta_0 + w_n
+%\]
+%where $r_n = 2s_n \in \ints$.  
+To ensure identifiability it is necessary to assume, as in~\cite{Sidiropoulos2005}, that $T_0$ lies in some interval $(T_{\text{min}}, T_{\text{max}})$ where $T_{\text{max}} < 2T_{\text{min}}$.%  We also assume that $y_1,\dots,y_N$ are in ascending order.
 The periodogram estimator is unchanged by a permutation observations, and so, this assumption can be made without loss of generality because it amounts only to sorting the observations if required.
-
-%BLERG. $T_{\text{min}}, T_{\text{max}}$ identifiability and $y_1,\dots,y_N$ in accending order.
-
 Fogel and Gavish~\cite{Fogel1988,Fogel1989_bit_synch_zero_crossings} proposed the estimator $\hat{T} = 1/\hat{f}$ where $\hat{f}$ is the maximiser of the~\emph{periodogram}
@@
 This is called the \emph{periodogram estimator}.
-This definition of the periodogram differs from the common definition in the spectral and frequency estimation literature~\cite{Hannan_time_series_1967,Quinn2001}.
+%This definition of the periodogram differs from the common definition in the spectral and frequency estimation literature~\cite{Hannan_time_series_1967,Quinn2001}.
 The statistical properties of the periodogram estimator were examined in~\cite{Quinn20013asilomar_period_est}.
@@
 However, the choice~\eqref{eq:gridwidthrecommended} does appear effective in practice.
-A statistically rigorous justification for this choice of $D$ is beyond the scope of the present paper, but could potentially be developed using the techniques in~\cite{Quinn2008maximizing_the_periodogram}.
+A statistically rigorous justification for this choice of $D$ is beyond the scope of the present paper, but could potentially be developed using techniques like those in~\cite{Quinn2008maximizing_the_periodogram}.
 
@@
 
-Denote by $\round{x}$ the nearest integer to $x$.
-The direction of rounding for half integers is not important.
-Half integers are rounded up in our implementation.
-Let $q$ be a real number and define the quantization function
+Denote by $\round{x} = \sfloor{x + \tfrac{1}{2}}$ the nearest integer to $x$.  %The direction of rounding for half integers is not important.  Half integers are rounded up in our implementation.  
+Let $q$ be a positive real number and define the quantization function
 \[

### 2014-06-14 paper.tex
minor

@@
 \end{align*}
+Observe that $I_z$ is periodic with period $q$.  This indicates that quantization should be fine enough that 
+\begin{equation}\label{eq:boundaliasingP}
+q > f_{\text{max}} - f_{\text{min}}.
+\end{equation}
+Otherwise there would exist frequencies in the interval $(f_{\text{min}}, f_{\text{max}})$ that are aliases with respect to $I_z$.
+%That is, there would exists two frequences $f_1,f_2 \in (f_{\text{min}}, f_{\text{max}})$ such that $f_1 = f_2 + 1/P$ and $I_z(f_1) = I_z(f_2)$.
+
 Let $\widetilde{f}^\prime$ be the maximiser of $I_z$ over the grid $G$, that is,
@@
 An iterative procedure %or Brent's method~\cite{Brent_opt_no_derivs_1973}
-can then be used to compute $\hat{f}$ starting from $\widetilde{f}^\prime$.  
-Observe that $I_z$ is periodic with period $q$.  This indicates that quantization should be fine enough that 
-\begin{equation}\label{eq:boundaliasingP}
-q > f_{\text{max}} - f_{\text{min}}.
-\end{equation}
-Otherwise there would exist frequencies in the interval $(f_{\text{min}}, f_{\text{max}})$ that are aliases with respect to $I_z$.
-%That is, there would exists two frequences $f_1,f_2 \in (f_{\text{min}}, f_{\text{max}})$ such that $f_1 = f_2 + 1/P$ and $I_z(f_1) = I_z(f_2)$.
+can then be used to compute $\hat{f}$ by maximising $I_y$ starting from $\widetilde{f}^\prime$.
 

### 2014-06-14 paper.tex
minor

@@
 \pgfplotsset{compat=1.8}
+\usetikzlibrary{pgfplots.groupplots}
 
@@
     \selectcolormodel{gray}
-    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.24)}},xlabel style={at={(0.5,0.07)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=7, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
+\begin{groupplot}[
+    group style={
+        group name=my plots,
+        group size=3 by 3,
+        ylabels at=edge left
+    },
+    legend style={
+      draw=none,
+      fill=none,
+      % legend pos=north west,
+      % cells={anchor=west},
+      % font=\footnotesize
+    },
+    footnotesize,
+    width=9cm,
+    height=11cm,
+    tickpos=left,
+    ytick align=outside,
+    xtick align=outside,
+    xmode=log,
+    ymode=log,
+    enlarge x limits=false 
+]
+
+\nextgroupplot[
+xlabel={$\sigma^2$},
+ylabel={MSE},
+]
+
 \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30geom1}; 
@@
 \addplot[color=gray,mark=triangle,mark options={solid,color=black,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q1.5geom1};
-      %\addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q3};
+% \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q3};
 \addplot[color=gray,mark=*,mark options={solid,color=black,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q5.0geom1};
@@
 \node[rotate=17,font=\footnotesize] at (axis cs:7e-4,1.5e-12) {$N=1200$};
-   \end{axis}  
- \begin{axis}[name=plot2,at={(11,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.23)}},xlabel style={at={(0.5,0.07)}}]
+
+\nextgroupplot[
+xlabel={$\sigma^2$}
+]
       \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN30geom10}; 
@@
      \node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=1200$};
-   \end{axis}
+
+\end{groupplot}
+
   \end{tikzpicture}  

### 2014-06-14 code/java/src/snpe/QuantisedPeriodogramChirpZ.java
switched back to the Newton-Raphson method

@@ -46,8 +46,8 @@ public class QuantisedPeriodogramChirpZ extends PeriodogramEstimator {
             }
         }        
         double ftilde = fmin + D*khat;
-        //double fhat = refine(fhat, y); //refine coarse frequencey estimate using Newton's method
-        double fhat = refine(ftilde, D, y); //refine coarse frequencey estimate using Brent's method
+        double fhat = refine(ftilde, y); //refine coarse frequencey estimate using Newton's method
+        //double fhat = refine(ftilde, D, y); //refine coarse frequencey estimate using Brent's method
         That = 1.0 / fhat;
         phat = phasestor.phase(y, That); //now compute the phase estimate        
     }

### 2014-06-14 code/java/src/snpe/QuantisedPeriodogramFFT.java

@@ -48,8 +48,8 @@ public class QuantisedPeriodogramFFT extends PeriodogramEstimator {
             }
         }        
         double ftilde = fmin + D*khat;
-        //double fhat = refine(fhat, y); //refine coarse frequencey estimate using Newton's method
-        double fhat = refine(ftilde, D, y); //refine coarse frequencey estimate using Brent's method
+        double fhat = refine(ftilde, y); //refine coarse frequencey estimate using Newton's method
+        //double fhat = refine(ftilde, D, y); //refine coarse frequencey estimate using Brent's method
         That = 1.0 / fhat;
         phat = phasestor.phase(y, That); //now compute the phase estimate        
     }

### 2014-06-14 response/response.tex
minor

@@
 
-In the simulations we regenerate the integers $s_1,\dots,s_N$ on each Monte-Carlo trial.
+In the simulations we regenerate the integers $s_1,\dots,s_N$ on each Monte-Carlo trial.  To make this clear in the paper the following sentence included
+
+``
+For each Monte-Carlo trial the integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed starting from zero with means $\mu=1$ and $10$.
+''
+
 To the authors knowledge, a Cramer-Rao bound is not known in this case.
@@
 
-In Figure~\ref{plot:crb} in this document we perform a simulation with integers $s_1,\dots,s_N$.
+In Figure~\ref{plot:crb} in this document we perform a simulation with fixed integers $s_1,\dots,s_N$.
 The integers are generated once such $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed.

### 2014-06-14 paper.tex
minor

@@
 Sadler and Casey~\cite{726812} have previously suggested use of the fast Fourier transform for period estimation.
-However, their estimator requires far larger transforms and produces less accurate results than the estimator described here (see Section~VI and Figures 1-4 in~\cite{726812}).
-In contrast, our estimator has accuracy similar to the least squares estimator, i.e., similar to the maximum likelihood estimator if the noise variables $w_1,\dots,w_N$ are assumed independent and identically normally distributed.
+However, their estimator requires larger transforms and produces less accurate results than the estimator described here (see Sec.~VI and Fig.
+1-4 in~\cite{726812}).
+Our estimator has accuracy similar to the least squares estimator, i.e., similar to the maximum likelihood estimator if the noise $w_1,\dots,w_N$ is assumed independent and identically normally distributed.
 
@@
 
-The previous section showed how $\widetilde{f}^\prime$ could be computed using either the chirp z-transform or a fast Fourier transform of length $M$.
+The previous section shows how $\widetilde{f}^\prime$ can be computed using either the chirp z-transform or a fast Fourier transform of length $M$.
 We now describe recommended values for the parameters $q, M$, and $D$.
-The simulations presented in Section~\ref{sec:simulations} suggest that a reasonable value for the quantization parameter is $q = 5 f_{\text{max}}$.
+The simulations presented in Section~\ref{sec:simulations} suggest that an effective value for the quantization parameter is $q = 5 f_{\text{max}}$.
 If the chirp z-transform is used then $D$ is chosen according to~\eqref{eq:gridwidthrecommended}.
-Computation of the chirp z-transform requires the computation of three fast Fourier transforms, each of length $L + K - 1$, where
+The chirp z-transform requires computation of three fast Fourier transforms, each of length $L + K - 1$, where
 \[
@@
 
-Observe that $L$ grows proportionally with $q$ and that $L = O(qN)$ under reasonable assumptions about the integers $s_1,\dots,s_N$ and the noise variables $w_1,\dots,w_N$.
-The most computationally expensive part of the chirp z-transform is the computation of three fast Fourier transforms of length $L+K-1$.
+Observe that $L$ grows proportionally with $q$ and that $L = O(qN)$ under reasonable assumptions about the integers $s_1,\dots,s_N$ and the noise $w_1,\dots,w_N$.
+%The most computationally expensive part of the chirp z-transform is the computation of three fast Fourier transforms of length $L+K-1$.
 Alternatively, a single fast Fourier transform of length $M = 2L$ can be used.
-Both approaches require $O(qN\log(qN)) = O(N\log N)$ operations because $q$ does not depend on $N$.
+Thus, both chirp-z and fast Fourier transform methods require $O(qN\log(qN)) = O(N\log N)$ operations because $q$ does not depend on $N$.
 However, it is apparent that the length of the transforms, and hence the complexity, increases as the quantization becomes finer, i.e., as $q$ increases.
 This presents a trade off between computational complexity and quantization accuracy.
-This trade off is investigated in the next section.
+This trade off is investigated in Section~\ref{sec:simulations}.
+
+Given the approximate maximiser $\widetilde{f}^\prime$ we apply the Newton-Raphson method to find a nearby stationary point of the periodogram $I_y$.  This proceeds by the iteration
+\[
+f_{n+1} = f_{n} - \frac{I_y^\prime(f_n)}{I_y^{\prime\prime}(f_n)}, \qquad f_{1} = \widetilde{f}^\prime
+\]
+where $I_y^\prime$ and $I_y^{\prime\prime}$ are the first and second derivative of $I_y$.
+This procedure converges to the maximiser $\hat{f}$ of $I_y$ provided that $\widetilde{f}^\prime$ is sufficiently close to $\hat{f}$.
+In our implementation iterations are performed until $\sabs{f_{n+1}-f_n} < 10^{-10}$.
+The number of iterations depends upon the distance $D$, but not on $N$.
+Each iteration requires $O(N)$ operations and so the total number of operations required by the Newton-Raphson procedure has order $O(N)$.
+In practice we find that the computational cost of the Newton-Raphson procedure is dominated by the cost of the chirp-z or fast Fourier transforms.
 
@@
 
-\begin{figure*}[t] 
+\begin{figure}[p] 
   \centering 
@@
         group name=my plots,
-        group size=3 by 3,
-        ylabels at=edge left
+        group size=1 by 4,
+        %ylabels at=edge left,
+        xlabels at=edge bottom,
+        xticklabels at=edge bottom,
+        vertical sep=0pt
     },
-    legend style={
-      draw=none,
-      fill=none,
-      % legend pos=north west,
-      % cells={anchor=west},
-      % font=\footnotesize
-    },
+    ylabel style={at={(0.07,0.4)}},
     footnotesize,
-    width=9cm,
-    height=11cm,
+    width=9.2cm,
+    height=6.9cm,
     tickpos=left,
@@
 \nextgroupplot[
-xlabel={$\sigma^2$},
 ylabel={MSE},
+legend style={
+  legend columns=2,
+  draw=none,
+  fill=none,
+  %legend pos=north west,
+  at={(0.75,1)},
+  cells={anchor=west},
+  font=\footnotesize,
+},
+ymin=2e-8
 ]
@@
 \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN30geom1};
-% \addplot[mark=none,dashed,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q2};
-% \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200};
-% \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200}; 
-% \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/PeriodogramN200};
-% \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q3};
-% \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q10};
-% \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/NormalisedSamplingLLSN200};
-\addplot[color=gray,mark=none,color=black] table {code/data/PeriodogramCLTN1200geom1}; 
-\addplot[color=gray,mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1200geom1};
-\addplot[color=gray,mark=o,mark options={color=black,solid,fill=black,scale=1}] table {code/data/PeriodogramN1200geom1};
-\addplot[color=gray,mark=x,mark options={color=black,solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN1200geom1};
-% \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1200geom1};
-\addplot[color=gray,mark=square,mark options={solid,color=black,fill=black,scale=1}] table {code/data/SLS2newN1200geom1};
-\addplot[color=gray,mark=triangle,mark options={solid,color=black,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q1.5geom1};
-% \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q3};
-\addplot[color=gray,mark=*,mark options={solid,color=black,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q5.0geom1};
-% \legend{Periodogram Theory, Least squares theory, Quantised periodogram, Periodogram, Least squares}
-\legend{Periodogram theory, Least squares theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
-\node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=30$};
-\node[rotate=17,font=\footnotesize] at (axis cs:7e-4,1.5e-12) {$N=1200$};
+
+\legend{Perdg. theory, Least squares theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
+\node[font=\footnotesize] at (axis cs:0.3,2e-8) [anchor=south east] {$\mu=1,N=30$};
 
 \nextgroupplot[
-xlabel={$\sigma^2$}
+ylabel={MSE},
+ymin=2e-10
 ]
@@
 %      \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/NormalisedSamplingLLSN200}; 
+       \node[font=\footnotesize] at (axis cs:0.3,2e-10) [anchor=south east] {$\mu=10,N=30$};`
+
+\nextgroupplot[
+ylabel={MSE},
+ymin=2e-13
+]
+
+\addplot[color=gray,mark=none,color=black] table {code/data/PeriodogramCLTN1200geom1}; 
+\addplot[color=gray,mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1200geom1};
+\addplot[color=gray,mark=o,only marks,mark options={color=black,solid,fill=black,scale=1}] table {code/data/PeriodogramN1200geom1};
+\addplot[color=gray,mark=x,only marks,mark options={color=black,solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN1200geom1};
+% \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1200geom1};
+\addplot[color=gray,mark=square,only marks,mark options={solid,color=black,fill=black,scale=1}] table {code/data/SLS2newN1200geom1};
+\addplot[color=gray,mark=triangle,only marks,mark options={solid,color=black,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q1.5geom1};
+% \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q3};
+\addplot[color=gray,mark=*,only marks,mark options={solid,color=black,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q5.0geom1};
+% \legend{Periodogram Theory, Least squares theory, Quantised periodogram, Periodogram, Least squares}
+\node[font=\footnotesize] at (axis cs:0.3,2e-13) [anchor=south east] {$\mu=1,N=1200$};
+
+\nextgroupplot[
+xlabel={$\sigma^2$},
+xlabel style={at={(0.5,0.1)}},
+ylabel={MSE},
+ymin=2e-15,
+ymax=8e-1
+%ytickten={-1,-3,-5,-7,-9,-11,-13}
+]
+
      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN1200geom10};
       \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN1200geom10};
-       \addplot[color=gray,mark=o,mark options={solid,color=black,fill=black,scale=1.1}] table {code/data/PeriodogramN1200geom10};
-       \addplot[color=gray,mark=x,mark options={solid,color=black,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN1200geom10};
+       \addplot[color=gray,mark=o,only marks,mark options={solid,color=black,fill=black,scale=1.1}] table {code/data/PeriodogramN1200geom10};
+       \addplot[color=gray,mark=x,only marks,mark options={solid,color=black,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN1200geom10};
  %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1100geom10};
-       \addplot[color=gray,mark=square,mark options={solid,color=black,fill=black,scale=1}] table {code/data/SLS2newN1200geom10};
-       \addplot[color=gray,mark=triangle,mark options={solid,color=black,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q1.5geom10};
+       \addplot[color=gray,mark=square,only marks,mark options={solid,color=black,fill=black,scale=1}] table {code/data/SLS2newN1200geom10};
+       \addplot[color=gray,mark=triangle,only marks,mark options={solid,color=black,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1200q1.5geom10};
       %\addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1100q3};
-      \addplot[color=gray,mark=*,mark options={solid,color=black,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q5.0geom10};
+      \addplot[color=gray,mark=*,only marks,mark options={solid,color=black,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN1200q5.0geom10};
  %     \legend{Periodogram Theory, Least squares theory, Quantised periodogram, Periodogram, Least squares}
      %\legend{Periodogram theory, $q=\tfrac{3}{2}T_{\text{max}}$, $q=4T_{\text{max}}$, Periodogram, Least squares, SLS2-adj}
-     \node[rotate=16,font=\footnotesize] at (axis cs:6e-4,9e-10) {$N=30$};
-     \node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=1200$};
+     \node[font=\footnotesize] at (axis cs:0.3,2e-15) [anchor=south east] {$\mu=10,N=1200$};
+   %  \node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=1200$};
 
@@
   \caption{Mean square period error versus noise variance $\sigma^2$.
-The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right.
-Data points for $N=1200$ are connected by gray lines.}\label{plot:multipleN}
-\end{figure*}
+For each Monte-Carlo trial the integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed with mean $\mu$.}\label{plot:multipleN}
+\end{figure}
 
@@
 
-This section presents the results of Monte-Carlo simulations with the periodogram estimator~\cite{Haohuan2013435,Fogel1989_bit_synch_zero_crossings}, the modified least squares estimator~\cite{Clarkson2007,McKilliam2007,Quinn_sparse_noisy_SSP_2012}, Clarkson's modification of the separable least squares line search (SLS2-new) estimator~\cite{Clarkson2007,Sidiropoulos2005}, and the quantized periodogram estimator from Section~\ref{sec:quant-peri}.
+This section presents the results of Monte-Carlo simulations with the periodogram estimator~\cite{Haohuan2013435,Fogel1989_bit_synch_zero_crossings}, the modified least squares estimator~\cite{Clarkson2007,McKilliam2007,Quinn_sparse_noisy_SSP_2012}, Clarkson's enhanced separable line search (SLS2-new) estimator~\cite{Clarkson2007,Sidiropoulos2005}, and the quantized periodogram estimator from Section~\ref{sec:quant-peri}.
 The grid distance $D$ used for the periodogram and SLS2-new estimators is given by~\eqref{eq:gridwidthrecommended}.
@@
 Both the periodogram and quanitised periodogram estimators use the Newton-Raphson method.
-In all simulations the noise variables $w_1,\dots,w_N$ are independent and identically normally distributed with variance $\sigma^2$.
-The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed starting from zero with means $\mu=1$ and $10$.
+In all simulations the noise $w_1,\dots,w_N$ are independent and identically normally distributed with variance $\sigma^2$.
+For each Monte-Carlo trial the integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed starting from zero with means $\mu=1$ and $10$.
 The true period is $T_0 = \pi/3$, the true phase is $\theta_0 = 0.2$, and $T_{\text{min}} = 0.8$ and $T_{\text{max}} = 1.5$.

### 2014-06-14 paper.tex
added more

@@
 
-Figure~\ref{plot:multipleN} shows the mean square error of the estimators versus $\sigma^2$ for $N = 30$ and $N = 1200$.
+Figure~\ref{plot:multipleN} shows the mean square error of the estimators versus noise variance $\sigma^2$.
+%for $N = 30$ and $N = 1200$.
 In the left hand  plot $\mu=1$ and in the right hand plot $\mu=10$.
@@
 Three thousand Monte-Carlo trials are run for each value of $\sigma^2$.
-The plots show the mean square error of the quantized periodogram estimator for two quantization parameters $q = \tfrac{3}{2}f_{\text{max}} = \tfrac{15}{8}$ and $q = 5 f_{\text{max}} = \tfrac{25}{4}$.
+The plots show the mean square error of the quantized periodogram estimator for two quantization parameters $q = \tfrac{3}{2}f_{\text{max}}$ and $q = 5 f_{\text{max}}$.
 The estimator performs poorly when $q = \tfrac{3}{2}f_{\text{max}}$ (triangles) but performs almost identically to the original periodogram estimator (circles) when $q = 5 f_{\text{max}}$ (dots).
 The plot also shows the mean square error of the modified least squares estimator~\cite{Clarkson2007,McKilliam2007} and the SLS2-new estimator~\cite{Clarkson2007,Sidiropoulos2005}.
-%The statistical performance of the SLS2-new estimator is poor when compared with the other estimators.
+The accuracy of the SLS2-new estimator is poor when compared with the other estimators.
 %For small $\sigma^2$ the mean square error of the SLS2 estimator is almost three orders of magnitude less accurate when $N = 50$ and almost five orders of magnitude less accurate when $N=1000$.
@@
 The dashed line shows the mean square error of the modified least squares estimator predicted by Theorem~3 from~\cite{Quinn_sparse_noisy_SSP_2012}.
+These theorems accurately predict the behaviour of the periodogram and least squares estimators when $\sigma^2$ is sufficiently small.
 
@@
 
-The new estimator has potential application to the synchronisation of communications devices~\cite{Fogel1988,Fogel1989_bit_synch_zero_crossings}, pulse repetition interval estimation in electronic support~\cite{EltonGray_puilse_train_rader_1994,Gray_more_pri_1994,Clarkson_thesis,clarkson_estimate_period_pulse_train_1996,Hauochan_pri_2012}, beat and tempo estimation in music~\cite{dixon_beat_extraction_2001}, spike interval estimation in neurology~\cite{Arnett_neuro_pri_1976,Brillinger_spike_trains_1988,Rossoni200630}, and in the analysis of internet traffic~\cite{He_detecting_periodic_patterns_in_internet_2006,5585849,5947313}.
+The new estimator has potential application to the synchronisation of communications devices~\cite{Fogel1988,Fogel1989_bit_synch_zero_crossings}, pulse interval estimation in electronic support~\cite{EltonGray_puilse_train_rader_1994,Gray_more_pri_1994,Clarkson_thesis,clarkson_estimate_period_pulse_train_1996,Hauochan_pri_2012}, beat and tempo estimation in music~\cite{dixon_beat_extraction_2001}, spike interval estimation in neurology~\cite{Arnett_neuro_pri_1976,Brillinger_spike_trains_1988,Rossoni200630}, and in the analysis of internet traffic~\cite{He_detecting_periodic_patterns_in_internet_2006,5585849,5947313}.
 

### 2014-06-14 response/response.tex
added more updates to response.  Reverted to older simulations using Newton-Raphson optimisation.

@@
 Despite being simple, the results are novel and useful.
-The quantisated periodogram estimator described in this paper is as accurate as the most accurate estimators in the literature (the least squares and periodogram estimator).
+The quantisated periodogram estimator described in this paper is as accurate as the most accurate estimators in the literature (the least squares and periodogram estimators).
 At the same time our estimator is substantially faster to compute.
 For these reasons, we firmly believe this will become the estimator of choice for periodic point processes.
-For the first time, it enables highly accurate period estimation for extremely large data sets in reasonable time.
+It enables highly accurate period estimation for extremely large data sets in reasonable time.
 This was not possible with previous estimators that require a number of operations growing quadratically with data size.
@@
 We anticipate that developing a rigorous theoretical method for choosing $q$ will be extremely complicated.
-The properties of many optimisation proceedures, such as the Newton-Raphson method, and Brent's method~\cite[Ch.~5]{Brent_opt_no_derivs_1973} are reasonably well understood for the purpose of maximising fixed known functions.
+The properties of many optimisation proceedures, such as the Newton-Raphson method, are reasonably well understood for the purpose of maximising fixed known functions.
 However, very few theorectical statistical analyses for these proceedures exist when the functions to be maximised are themselves random variables, i.e., when noise is involved.
@@
 We agree with the reviewer that the simulations presented are limited.
-This is entirely due to the 4 page limit required IEEE Signal Processing Letters.
+This is due to the 4 page limit required IEEE Signal Processing Letters (a 5th page containing only references is also allowed).
 In the paper we use our simulations to motivate the choice $q = 5 f_{\text{max}}$ for the quantisation parameter.
-It is natural to question whether this choice also works in other scenarios, for example (as the reviewer suggests), for different noise distributions, and different methods of generating the integers $s_1,\dots,s_N$.
-The results of such simulations are displayed in Figures~\ref{plot:poisson} and~\ref{plot:geomuniform} in this document.
-The simulations are setup as in the paper, but with the following modifications:
+We have found this choice works well under an array of different noise distributions and methods for generating the intergers $s_1,\dots,s_N$  For example, Figures~\ref{plot:poisson} and~\ref{plot:geomuniform} in this document show simulations similar to those in the paper, but with the following modifications:
 
@@
 \item In Figure~\ref{plot:poisson} the noise is Gaussian distributed, $N=200$, and the integers are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Poisson distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right.
+One hundred Monte-Carlo trials are used for each value of $\sigma^2$.
 \item In Figure~\ref{plot:geomuniform} the noise if uniformly distributed, $N=200$, and the integers are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right.
+One hundred Monte-Carlo trials are used for each value of $\sigma^2$.
+The least squares estimator is the most accurate in this simulation.
+The is predicted by the asymptotic theory~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}.
 \end{enumerate}
 
-Observe that, in both simulations, the choice $q=5 f_{\text{max}}$ is a good one.
-With this choice the accuracy of the quantised periodogram estimator is similar to the original periodgram estimator (computed without quantisation and FFT).
+Observe that, in both simulations, the choice $q=5 f_{\text{max}}$ provide good accuracy.
+With this choice, the accuracy of the quantised periodogram estimator is similar to the original periodgram estimator computed without quantisation and FFT.
 %That the choice $q=5 f_{\text{max}}$ works in a wide range of scenarios should not come as a
 
-\begin{figure*}[tp] 
+\begin{figure*}[p] 
   \centering 
@@
 
-\begin{figure*}[tp]  
+\begin{figure*}[p]  
   \centering
@@
 close the what might theoretically be expected from the maximum likelihood estimator.
-
-\begin{figure*}[tp]  
-  \centering
-  \begin{tikzpicture}  
-    \selectcolormodel{gray}
-    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.24)}},xlabel style={at={(0.5,0.07)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=8, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
-\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN500fixed1}; 
-      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN500fixed1};
-      \addplot[mark=none,color=black,dotted] table {code/data/CRBN500fixed1};
-            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN500q1.5fixed1};
-      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN500q5.0fixed1};
-      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN500fixed1};
-       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN500fixed1}; 
- %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN500fixed1};
-      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN500fixed1};
-     \legend{Periodogram theory, LS theory, CRB, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, LS, SLS2-new}
-     %\node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=200$};
-   \end{axis}   \begin{axis}[name=plot2,at={(9.5,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.23)}},xlabel style={at={(0.5,0.07)}}]
-\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN500fixed10}; 
-      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN500fixed10};
-      \addplot[mark=none,color=black,dotted] table {code/data/CRBN500fixed10};
-            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN500q1.5fixed10};
-      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN500q5.0fixed10};
-      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN500fixed10};
-       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN500fixed10}; 
-  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN500fixed10};
-       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN500fixed10};
-     %\node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=200$}; 
-   \end{axis} 
-  \end{tikzpicture}
-  \caption{Mean square period error versus noise variance $\sigma^2$ with Gaussian noise and $N=500$ observations.
-A singled fixed set of integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$.
-This single fixed sequence is used for all of the $500$ Monte-Carlo trials.
-In this case the \emph{clairvoyant} Cramer-Rao bound from~\cite{Sidiropoulos2005} (also in \cite{726812,Clarkson2007}) exists and is displayed by the dotted line.
-In this simulation the \emph{clairvoyant} Cramer-Rao bound is equal to the variance of the maximum likelihood (i.e.
-least squares) estimator in the case that the integers $s_1,\dots,s_N$ are known.}\label{plot:crb}
-\end{figure*}
 
@@
 
+\begin{figure}[tp]  
+  \centering
+  \begin{tikzpicture}  
+    \selectcolormodel{gray}
+    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.24)}},xlabel style={at={(0.5,0.07)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=8, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
+\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN500fixed1}; 
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN500fixed1};
+      \addplot[mark=none,color=black,dotted] table {code/data/CRBN500fixed1};
+            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN500q1.5fixed1};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN500q5.0fixed1};
+      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN500fixed1};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN500fixed1}; 
+ %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN500fixed1};
+      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN500fixed1};
+     \legend{Periodogram theory, LS theory, CRB, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, LS, SLS2-new}
+     %\node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=200$};
+   \end{axis}   \begin{axis}[name=plot2,at={(9.5,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.23)}},xlabel style={at={(0.5,0.07)}}]
+\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN500fixed10}; 
+      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN500fixed10};
+      \addplot[mark=none,color=black,dotted] table {code/data/CRBN500fixed10};
+            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN500q1.5fixed10};
+      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN500q5.0fixed10};
+      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN500fixed10};
+       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN500fixed10}; 
+  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN500fixed10};
+       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN500fixed10};
+     %\node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=200$}; 
+   \end{axis} 
+  \end{tikzpicture}
+  \caption{Mean square period error versus noise variance $\sigma^2$ with Gaussian noise and $N=500$ observations.
+A singled fixed set of integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$.
+This single fixed sequence is used for all of the $500$ Monte-Carlo trials.
+In this case the \emph{clairvoyant} Cramer-Rao bound from~\cite{Sidiropoulos2005} (also in \cite{726812,Clarkson2007}) exists and is displayed by the dotted line.
+In this simulation the \emph{clairvoyant} Cramer-Rao bound is equal to the variance of the maximum likelihood (i.e.
+least squares) estimator in the case that the integers $s_1,\dots,s_N$ are known.}\label{plot:crb}
+\end{figure}
+
 \section*{Reviewer 2}

### 2014-06-14 response/response.tex
added more edits.  Getting there.

@@
 We do not believe this to be a weakness.
-Despite being simple, the results are novel and useful.
+Whilst being simple, the results are also novel and useful.
 The quantisated periodogram estimator described in this paper is as accurate as the most accurate estimators in the literature (the least squares and periodogram estimators).
@@
 \begin{enumerate}
-\item In Figure~\ref{plot:poisson} the noise is Gaussian distributed, $N=200$, and the integers are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Poisson distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right.
-One hundred Monte-Carlo trials are used for each value of $\sigma^2$.
-\item In Figure~\ref{plot:geomuniform} the noise if uniformly distributed, $N=200$, and the integers are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right.
-One hundred Monte-Carlo trials are used for each value of $\sigma^2$.
+\item In Figure~\ref{plot:poisson} the noise is Gaussian distributed, $N=200$, and the integers are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically \emph{Poisson distributed} with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right.
+Five hundred Monte-Carlo trials are used for each value of $\sigma^2$.
+\item In Figure~\ref{plot:geomuniform} the noise if \emph{uniformly distributed}, $N=200$, and the integers are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right.
+Five hundred Monte-Carlo trials are used for each value of $\sigma^2$.
 The least squares estimator is the most accurate in this simulation.
-The is predicted by the asymptotic theory~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}.
+This is predicted by the asymptotic theory~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}.
 \end{enumerate}
@@
 
-\begin{figure*}[p] 
+\begin{figure}[p] 
   \centering 
@@
 The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Poisson distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$.}\label{plot:poisson}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[p]  
+\begin{figure}[p]  
   \centering
@@
 The least squares estimator is the most accurate in this simulation.
-The is predicted by the asymptotic theory~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}.}\label{plot:geomuniform}
-\end{figure*}
+This is predicted by the asymptotic theory~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}.}\label{plot:geomuniform}
+\end{figure}
 
@@
 compare their results in term of accuracy and of computational complexity to
-the results of [24]. Including the Cramer-Rao bounds for the estimation of T0
+the results of [24]. Including the Cramer-Rao bounds for the estimation of $T_0$
 (as in [24]) would also be appreciated.
@@
 \textbf{Response:}
+
+We response to each of the reviewers comments in turn.
+
+\begin{enumerate}
+
+\item \textbf{Unfortunately, there is nothing in
+Section VI (which is the conclusion) supporting this claim.}
+
 Section VI in the paper~\cite{726812} entitled ``On periodic pulse interval analysis with outliers and missing observations'' is not the conclusion.
@@
 Casey and Sadler do use an FFT to compute the periodgram, but fail to implement an optimisation proceedure.
-We used the Newton-Raphson proceedure in the old manuscript, but have used Brent's method~\cite{Brent_opt_no_derivs_1973} in the new manuscript as dicussed in our response to comment~\ref{com:rev1:nr} below.
-Casey and Sadler arrive at the conclusion that very large FFTs are required and that the periodgram estimator is not particularly accurate.
-In the conclusion of~\cite{726812} it is stated that:
+% We used the Newton-Raphson proceedure in the old manuscript, but have used Brent's method~\cite{Brent_opt_no_derivs_1973} in the new manuscript as dicussed in our response to comment~\ref{com:rev1:nr} below.
+Casey and Sadler arrive at the conclusion that very large FFTs are required and that the periodgram estimator is not particularly accurate.  In the conclusion of~\cite{726812} it is stated that:
 
@@
 Our paper clearly shows that such limitation do not exist.
-The periodgram estimator, computed by quanitisation and FFT and combined with an appropriate optimisation proceedure, is both compuationally efficient and highly accurate.
+The periodgram estimator, computed by quantization and FFT and combined with an appropriate optimisation proceedure, is both compuationally efficient and highly accurate.
 For moderately large $N$ this is the most computationally efficient proceedure currently available for this problem.
@@
 
-In the simulations we regenerate the integers $s_1,\dots,s_N$ on each Monte-Carlo trial.  To make this clear in the paper the following sentence included
+\item \textbf{ The authors should
+compare their results in term of accuracy and of computational complexity to
+the results of [24].}
 
-``
-For each Monte-Carlo trial the integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed starting from zero with means $\mu=1$ and $10$.
-''
+We have not implemented and do not present simulations results with any of the MEA estimators from the papers by Casey and Sadler~\cite{536682,726812}.
+The MEA estimators were supersceeded by the Separable Line Search (SLS) estimators of Sidiropoulos, Swami and Sadler~\cite{Sidiropoulos2005} in 2005.
+Comparison between the MEA estimators and the SLS estimators are presented in~\cite{Sidiropoulos2005} where the superiority of the SLS estimators is shown.
+An enhanced SLS estimator (SLS2-new) was developed in~\cite{Clarkson2007} and this is the estimator we show simulations results for\footnote{One of the SLS estimators (SLS2-all) has excellent statistical accuracy.
+It is almost as accurate as the least squares and periodogram estimators as can be see in the simulations in~\cite{Clarkson2007} and~\cite{McKilliam2007}.
+However, it is extremely computationally complex, requiring atleast $O(N^3)$ operations to compute.
+We cannot present simulation results for this estimator when $N=1200$ for example.
+This estimator is discussed and compared in~\cite{Sidiropoulos2005} and also~\cite{Clarkson2007,McKilliam2007}.}.
+Even this enhanced SLS estimator performs poorly when compared to the least squares and periodogram estimators.
+The MEA estimators are no longer state-of-the-art for this problem.
+
+\item \textbf{Including the Cramer-Rao bounds for the estimation of $T_0$
+(as in [24]) would also be appreciated.}
+
+In the simulations we regenerate the integers $s_1,\dots,s_N$ on each Monte-Carlo trial.  To make this clear in the paper the following sentence is included:
+
+``For each Monte-Carlo trial the integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed starting from zero with means $\mu=1$ and $10$.''
 
 To the authors knowledge, a Cramer-Rao bound is not known in this case.
-The \emph{clairvoyant} Cramer-Rao bound from~\cite{Sidiropoulos2005} (also in \cite{726812,Clarkson2007}) supposes a single fixed realisation for $s_1,\dots,s_N$.
-This bound is therefore not valid for the simulations we present in Figure~\ref{plot:multipleN} in Section~\ref{sec:simulations} of the paper.
+The \emph{clairvoyant} Cramer-Rao bound from~\cite{Sidiropoulos2005} (also \cite{726812,Clarkson2007}) supposes a single fixed realisation for $s_1,\dots,s_N$.
+This bound is therefore not valid for the simulations we present in Figure~\zref{plot:multipleN} in Section~\zref{sec:simulations} of the paper.
 
@@
 The integers are generated once such $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed.
-This single fixed sequence is used for all of the $1000$ Monte-Carlo trials.
-Observe that for small noise variance the \emph{clairvoyant} Cramer-Rao bound is similar to the asymptotic variance of the least square and periodogram estimator predicted in~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}.
-As the noise variance increases the \emph{clairvoyant} Cramer-Rao bound and the asymptotic variances diverge.
+This single fixed sequence is used for all 500 Monte-Carlo trials.
+Observe that for small noise variance the \emph{clairvoyant} Cramer-Rao bound is similar to the asymptotic variance of the least squares and periodogram estimators predicted in~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}.
+As the noise variance increases the \emph{clairvoyant} Cramer-Rao bound and the predicted asymptotic variances diverge.
 As expected, the behaviour of the estimators themselves follows the asymptotic theory, rather than the \emph{clairvoyant} Cramer-Rao bound.
 
-Observe that in the right hand plot the quantised periodogram estimator with coarse quantisation $q = 1.5 f_{\text{max}}$ actually perform well for sufficently small noise variance.
+Observe that in the left hand plot of Figure~\ref{plot:crb} the quantised periodogram estimator with coarse quantisation $q = 1.5 f_{\text{max}}$ (triangles) actually performs well for sufficently small noise variance.
 It appears that the fixed integers $s_1,\dots,s_N$ generated in this simulation favour the quantised periodogram estimator.
-This is one motivation for performing our simulations with integer $s_1,\dots,s_N$ generated randomly for each Monte-Carlo trial, rather than choosing a single fixed set of integers $s_1,\dots,s_N$.
-It avoid the scenario where a particular favourable (or unfavourable) set $s_1,\dots,s_N$ is chosen by chance.
-
-We have not implemented and do not present simulations results with any of the MEA estimators from the papers by Casey and Sadler~\cite{536682,726812}.
-The MEA estimators were supersceeded by the Separable Line Search (SLS) estimators of Sidiropoulos, Swami and Sadler~\cite{Sidiropoulos2005} in 2005.
-Comparison between the MEA estimator and the SLS are presented in~\cite{Sidiropoulos2005} where the superiority of the SLS estimators is shown.
-An enhanced SLS estimator (SLS2-new) was developed in~\cite{Clarkson2007} and this is the estimator we show simulations results for.
-Even this enhanced SLS estimator performs poorly compared to the least squares and periodogram estimators.
-The MEA estimators are no longer state-of-the-art for this problem.
+This is one motivation for performing simulations with integers $s_1,\dots,s_N$ generated randomly for each Monte-Carlo trial, rather than choosing a single fixed set of integers.
+It avoids the scenario where a particular favourable set $s_1,\dots,s_N$ is chosen by chance.
 
@@
 close the what might theoretically be expected from the maximum likelihood estimator.
+
+\end{enumerate}
 
@@
 \begin{enumerate}
-\item \textbf{For instance, what is the exact NR scheme used in this paper (there are several options)?}  The Netwon-Raphson method that was used proceeds by the iteration
+\item \textbf{For instance, what is the exact NR scheme used in this paper (there are several options)?}  
+In the new version of the paper the following paragraph has been included in Section~\zref{sec:comp-compl}:
+
+``Given the approximate maximiser $\widetilde{f}^\prime$ we apply the Newton-Raphson method to find a nearby stationary point of the periodogram $I_y$.  This proceeds by the iteration
 \[
-x_{n+1} = x_n - \frac{f^\prime(x_{n})}{f^{\prime\prime}(x_n)}
+f_{n+1} = f_{n} - \frac{I_y^\prime(f_n)}{I_y^{\prime\prime}(f_n)}, \qquad f_{1} = \widetilde{f}^\prime
 \]
-to find a stationary point of the function $f$ where $f^\prime$ and $f^{\prime\prime}$ are the first and second derivatives of $f$.
+where $I_y^\prime$ and $I_y^{\prime\prime}$ are the first and second derivative of $I_y$.
+This procedure converges to the maximiser $\hat{f}$ of $I_y$ provided that $\widetilde{f}^\prime$ is sufficiently close to $\hat{f}$.
+In our implementation iterations are performed until $\sabs{f_{n+1}-f_n} < 10^{-10}$.
+The number of iterations depends upon the distance $D$, but not on $N$.
+Each iteration requires $O(N)$ operations and so the total number of operations required by the Newton-Raphson procedure has order $O(N)$.
+In practice we find that the computational cost of the Newton-Raphson procedure is dominated by the cost of the chirp-z or fast Fourier transforms.''
+
+% The Netwon-Raphson method that was used proceeds by the iteration
+% \[
+% x_{n+1} = x_n - \frac{f^\prime(x_{n})}{f^{\prime\prime}(x_n)}
+% \]
+% to find a stationary point of the function $f$ where $f^\prime$ and $f^{\prime\prime}$ are the first and second derivatives of $f$.
 To avoid the potential ambguity regarding the definition of the Newton-Raphson method we have, in the new manuscipt, replaced the Newton-Raphson method with the optimisation proceedure due to Brent~\cite[Ch.~5]{Brent_opt_no_derivs_1973}.
@@
 
-BLERG add some text to the paper.
+% BLERG add some text to the paper.
 
-\item \textbf{How have the authors selected the total number of iterations?} In our implementation iterations are performed until the absolute difference between consequentive iterations is less than $10^{-7}$.
+\item \textbf{How have the authors selected the total number of iterations?} 
+As described in the above paragraph ``iterations are performed until $\sabs{f_{n+1}-f_n} < 10^{-10}$.''
 
-BLERG: Something like the following should be added:
+% In our implementation iterations are performed until the absolute difference between consequentive iterations is less than $10^{-7}$.
 
-``Brent's method is iterative.
+% BLERG: Something like the following should be added:
+
+% ``Brent's method is iterative.
 During the method a number of evaluations of the periodogram $I_y$ are made.
@@
 
-\item \textbf{Is the quantized periodogram
-just used to initialize a Newton-Raphson procedure? Do the MSEs displayed
-in Fig. 3 correspond to the estimator obtained after applying the NR algorithm?}
+\item \textbf{Is the quantized periodogram just used to initialize a Newton-Raphson procedure?}
+Yes.  As described in the paper BLERG
+
+\item \textbf{Do the MSEs displayed in Fig. 3 correspond to the estimator obtained after applying the NR algorithm?}
+Yes.
 
@@
 initialization of the NR procedure?}
+Yes.
+The estimators run in the benchmarks in Figure~\zref{plot:benchmark} are exactly the same as those in the simulations in Figure~\zref{plot:multipleN}.
+For the periodogram and quantised peridogram estimators this includes the Newton-Raphson proceedure.
 
 \end{enumerate}
-
-We are only aware of a single   As was stated Section~\zref{sec:simulations}: ``Both the periodogram and quanitised periodogram estimators use the Newton-Raphson method.''  This was the case both for the simulations presented in Figure~\zref{plot:multipleN} and for the computational benchmarks presented in Figure~\zref{plot:benchmark}.
-
-BLERG: In Section~\zref{sec:peri-estim} the text has been modifed to read:
-
-``An iterative procedure, such as the Newton-Raphson method, or Brent's method~\cite{Brent_opt_no_derivs_1973} 
-can then be used to compute $\hat{f}$ starting from $\widetilde{f}$.''
 
@@
 \textbf{Response:}
-We have connected the data points for the $N=1200$ case by gray lines.  We believe this makes the plot clear.  We are happy to make further adjustments to this plot should the reviewer require them.
+The formatting of Figure~\zref{plot:multipleN} has been improved.  Each simulation for $N=30,1200$ and $\mu=1,10$ is now presented in its own plot.
 

### 2014-06-14 paper.tex
added more updates to response.  Reverted to older simulations using Newton-Raphson optimisation.
added more edits.  Getting there.

@@
 1-4 in~\cite{726812}).
-Our estimator has accuracy similar to the least squares estimator, i.e., similar to the maximum likelihood estimator if the noise $w_1,\dots,w_N$ is assumed independent and identically normally distributed.
+%Our estimator has accuracy similar to the least squares estimator, i.e., similar to the maximum likelihood estimator if the noise $w_1,\dots,w_N$ is assumed independent and identically normally distributed.
 
@@
 %where $r_n = 2s_n \in \ints$.  
-To ensure identifiability it is necessary to assume, as in~\cite{Sidiropoulos2005}, that $T_0$ lies in some interval $(T_{\text{min}}, T_{\text{max}})$ where $T_{\text{max}} < 2T_{\text{min}}$.%  We also assume that $y_1,\dots,y_N$ are in ascending order.
+To ensure identifiability it is necessary to assume, as in~\cite{Sidiropoulos2005}, that $T_0$ lies in an interval $(T_{\text{min}}, T_{\text{max}})$ where $T_{\text{max}} < 2T_{\text{min}}$.
+%We also assume that $y_1,\dots,y_N$ are in ascending order.
 The periodogram estimator is unchanged by a permutation observations, and so, this assumption can be made without loss of generality because it amounts only to sorting the observations if required.
@@
 If the distance $D$ is sufficiently small, then $\widetilde{f}$ is a good approximation to the maximiser $\hat{f}$.
-An iterative procedure, such as the Newton-Raphson method, or Brent's method~\cite[Ch.~5]{Brent_opt_no_derivs_1973}
+An iterative procedure, such as the Newton-Raphson method, %or Brent's method~\cite[Ch.~5]{Brent_opt_no_derivs_1973}
 can then be used to compute $\hat{f}$ starting from $\widetilde{f}$.
@@
 \end{align*}
-Observe that $I_z$ is periodic with period $q$.  This indicates that quantization should be fine enough that 
+Observe that $I_z$ is periodic with period $q$ and so quantization should be fine enough that 
 \begin{equation}\label{eq:boundaliasingP}
@@
 can then be used to compute $\hat{f}$ by maximising $I_y$ starting from $\widetilde{f}^\prime$.
+Section~\ref{sec:comp-compl} describes an implementation of the Newton-Raphson optimisation procedure for this purpose.
 

### 2014-06-14 response/code/montecarlo.scala

@@ -21,7 +21,7 @@ import snpe.theory.PeriodogramCLT
 import snpe.theory.NormalisedLLSCLT
 import snpe.theory.ClairvoyantCRB
 
-val iters = 100
+val iters = 500
 val T_0 = scala.math.Pi/3 //true period
 val T_min = 0.8 //minimum period
 val T_max = 1.5 //maximum period

### 2014-06-14 response/response.tex
minor

@@
 Each iteration requires $O(N)$ operations and so the total number of operations required by the Newton-Raphson procedure has order $O(N)$.
-In practice we find that the computational cost of the Newton-Raphson procedure is dominated by the cost of the chirp-z or fast Fourier transforms.''
+In practice we find that the computational cost of the Newton-Raphson procedure negligible when compared to the cost of the chirp-z or fast Fourier transforms.''
 

### 2014-06-14 paper.tex
minor

@@
 
-The previous section shows how $\widetilde{f}^\prime$ can be computed using either the chirp z-transform or a fast Fourier transform of length $M$.
+The previous section shows how the approximate maximiser $\widetilde{f}^\prime$ can be computed using either the chirp z-transform or a fast Fourier transform of length $M$.
 We now describe recommended values for the parameters $q, M$, and $D$.
@@
 Thus, both chirp-z and fast Fourier transform methods require $O(qN\log(qN)) = O(N\log N)$ operations because $q$ does not depend on $N$.
-However, it is apparent that the length of the transforms, and hence the complexity, increases as the quantization becomes finer, i.e., as $q$ increases.
+However, the length of the transforms, and hence the complexity, increases as the quantization becomes finer, i.e., as $q$ increases.
 This presents a trade off between computational complexity and quantization accuracy.
@@
 Each iteration requires $O(N)$ operations and so the total number of operations required by the Newton-Raphson procedure has order $O(N)$.
-In practice we find that the computational cost of the Newton-Raphson procedure is dominated by the cost of the chirp-z or fast Fourier transforms.
+In practice we find that the computational cost of the Newton-Raphson procedure negligible when compared to the cost of the chirp-z or fast Fourier transforms.
 

### 2014-06-14 paper.tex
minor

@@
       \addplot[mark=square,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2newBenchmark};
-      \legend{Periodogram, Least squares, Fast Fourier transform, Chirp z-transform, SLS2-new}
+      \legend{Periodogram, Least squares, Fast Fourier transform, Chirp z-transform, SLS-new}
    \end{axis}
@@
 
-\legend{Perdg. theory, Least squares theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
+\legend{Perdg. theory, Least squares theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, Least squares, SLS-new}
 \node[font=\footnotesize] at (axis cs:0.3,2e-8) [anchor=south east] {$\mu=1,N=30$};
@@
 
-This section presents the results of Monte-Carlo simulations with the periodogram estimator~\cite{Haohuan2013435,Fogel1989_bit_synch_zero_crossings}, the modified least squares estimator~\cite{Clarkson2007,McKilliam2007,Quinn_sparse_noisy_SSP_2012}, Clarkson's enhanced separable line search (SLS2-new) estimator~\cite{Clarkson2007,Sidiropoulos2005}, and the quantized periodogram estimator from Section~\ref{sec:quant-peri}.
-The grid distance $D$ used for the periodogram and SLS2-new estimators is given by~\eqref{eq:gridwidthrecommended}.
+This section presents the results of Monte-Carlo simulations with the periodogram estimator~\cite{Haohuan2013435,Fogel1989_bit_synch_zero_crossings}, the modified least squares estimator~\cite{Clarkson2007,McKilliam2007,Quinn_sparse_noisy_SSP_2012}, Clarkson's enhanced separable line search (SLS-new) estimator~\cite{Clarkson2007,Sidiropoulos2005}, and the quantized periodogram estimator from Section~\ref{sec:quant-peri}.
+The grid distance $D$ used for the periodogram and SLS-new estimators is given by~\eqref{eq:gridwidthrecommended}.
 The modified least squares estimator requires a smaller grid distance to work accurately and we choose~\eqref{eq:gridwidthrecommended} divided by three in our simulations.
@@
 The estimator performs poorly when $q = \tfrac{3}{2}f_{\text{max}}$ (triangles) but performs almost identically to the original periodogram estimator (circles) when $q = 5 f_{\text{max}}$ (dots).
-The plot also shows the mean square error of the modified least squares estimator~\cite{Clarkson2007,McKilliam2007} and the SLS2-new estimator~\cite{Clarkson2007,Sidiropoulos2005}.
-The accuracy of the SLS2-new estimator is poor when compared with the other estimators.
-%For small $\sigma^2$ the mean square error of the SLS2 estimator is almost three orders of magnitude less accurate when $N = 50$ and almost five orders of magnitude less accurate when $N=1000$.
+The plot also shows the mean square error of the modified least squares estimator~\cite{Clarkson2007,McKilliam2007} and the SLS-new estimator~\cite{Clarkson2007,Sidiropoulos2005}.
+The accuracy of the SLS-new estimator is poor when compared with the other estimators.
+%For small $\sigma^2$ the mean square error of the SLS estimator is almost three orders of magnitude less accurate when $N = 50$ and almost five orders of magnitude less accurate when $N=1000$.
 The solid line shows the mean square error of the periodogram estimator predicted by Theorem~2 from~\cite{Quinn20013asilomar_period_est}.
@@
 %The quantized periodogram estimator computed using a single fast Fourier transform is faster than the periodogram estimator when $N \geq 20$.
-It is faster than the least squares estimator when $N \geq 80$ and it is faster than the SLS2-new estimator when $N \geq 400$.
+It is faster than the least squares estimator when $N \geq 80$ and it is faster than the SLS-new estimator when $N \geq 400$.
 The quantized periodogram estimator is significantly faster than the other estimators for large $N$.

### 2014-06-15 response/response.tex
added final versions.

--- 2014-06-14 response/response.tex
@@
 Whilst being simple, the results are also novel and useful.
-The quantisated periodogram estimator described in this paper is as accurate as the most accurate estimators in the literature (the least squares and periodogram estimators).
+The quantized periodogram estimator described in this paper is as accurate as the most accurate estimators in the literature (the least squares and periodogram estimators).
 At the same time our estimator is substantially faster to compute.
@@
 We anticipate that developing a rigorous theoretical method for choosing $q$ will be extremely complicated.
-The properties of many optimisation proceedures, such as the Newton-Raphson method, are reasonably well understood for the purpose of maximising fixed known functions.
-However, very few theorectical statistical analyses for these proceedures exist when the functions to be maximised are themselves random variables, i.e., when noise is involved.
-Add to this the existence of quantisation (itself typically very tricky to analyse) and it appears that a theorectical analysis of this estimator will be extremely difficult.
+The properties of many optimisation procedures, such as the Newton-Raphson method, are reasonably well understood for the purpose of maximising fixed known functions.
+However, very few theoretical statistical analyses for these procedures exist when the functions to be maximised are themselves random variables, i.e., when noise is involved.
+Add to this the existence of quantization (itself typically very tricky to analyse) and it appears that a theoretical analysis of this estimator will be extremely difficult.
 
@@
 This is due to the 4 page limit required IEEE Signal Processing Letters (a 5th page containing only references is also allowed).
-In the paper we use our simulations to motivate the choice $q = 5 f_{\text{max}}$ for the quantisation parameter.
-We have found this choice works well under an array of different noise distributions and methods for generating the intergers $s_1,\dots,s_N$  For example, Figures~\ref{plot:poisson} and~\ref{plot:geomuniform} in this document show simulations similar to those in the paper, but with the following modifications:
+In the paper we use our simulations to motivate the choice $q = 5 f_{\text{max}}$ for the quantization parameter.
+We have found this choice works well under an array of different noise distributions and methods for generating the integers $s_1,\dots,s_N$.
+For example, Figures~\ref{plot:poisson} and~\ref{plot:geomuniform} in this document show simulations similar to those in the paper, but with the following modifications:
 
@@
 Five hundred Monte-Carlo trials are used for each value of $\sigma^2$.
-\item In Figure~\ref{plot:geomuniform} the noise if \emph{uniformly distributed}, $N=200$, and the integers are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right.
+\item In Figure~\ref{plot:geomuniform} the noise if \emph{uniformly distributed}, $N=200$, and the integers are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right.
 Five hundred Monte-Carlo trials are used for each value of $\sigma^2$.
@@
 
-Observe that, in both simulations, the choice $q=5 f_{\text{max}}$ provide good accuracy.
-With this choice, the accuracy of the quantised periodogram estimator is similar to the original periodgram estimator computed without quantisation and FFT.
+Observe that, in both simulations, the choice $q=5 f_{\text{max}}$ provides good accuracy.
+With this choice, the accuracy of the quantized periodogram estimator is similar to the original periodogram estimator computed without quantization and FFT.
 %That the choice $q=5 f_{\text{max}}$ works in a wide range of scenarios should not come as a
@@
   \end{tikzpicture}  
-  \caption{Mean square period error versus noise variance $\sigma^2$ with Gaussain noise and $N=200$ observations.
+  \caption{Mean square period error versus noise variance $\sigma^2$ with Gaussian noise and $N=200$ observations.
 The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Poisson distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$.}\label{plot:poisson}
@@
 
-We response to each of the reviewers comments in turn.
+We respond to each of the reviewers comments in turn.
 
@@
 It is the section entitled ``Numerical Results''.
-In Figures 1 to 4 in~\cite{726812} the periogram estimator implemented by Casey and Sadler is seen to be inferior to their MEA estimator.
-Casey and Sadler do use an FFT to compute the periodgram, but fail to implement an optimisation proceedure.
-% We used the Newton-Raphson proceedure in the old manuscript, but have used Brent's method~\cite{Brent_opt_no_derivs_1973} in the new manuscript as dicussed in our response to comment~\ref{com:rev1:nr} below.
-Casey and Sadler arrive at the conclusion that very large FFTs are required and that the periodgram estimator is not particularly accurate.  In the conclusion of~\cite{726812} it is stated that:
+In Figures 1 to 4 in~\cite{726812} the periodogram estimator implemented by Casey and Sadler is seen to be inferior to their MEA estimator.
+Casey and Sadler do use an FFT to compute the periodogram, but fail to implement an optimisation procedure.
+We used the Newton-Raphson proceedure (see our response to comment~\ref{com:rev1:nr} below).
+Casey and Sadler arrive at the conclusion that very large FFTs are required and that the periodogram estimator is not particularly accurate.
+In the conclusion of~\cite{726812} it is stated that:
 
@@
 Our paper clearly shows that such limitation do not exist.
-The periodgram estimator, computed by quantization and FFT and combined with an appropriate optimisation proceedure, is both compuationally efficient and highly accurate.
-For moderately large $N$ this is the most computationally efficient proceedure currently available for this problem.
-Simultaneously, it is as accurate at the most accurate estimators available, namely the least squares estimator, and the periodgram estimator computed directly without FFT.
+The periodogram estimator, computed by quantization and FFT and combined with an appropriate optimisation procedure, is both computationally efficient and highly accurate.
+This is the most computationally efficient procedure currently available for this problem.
+Simultaneously, it is as accurate at the most accurate estimators available, namely the least squares estimator, and the periodogram estimator computed directly without FFT.
 
@@
 
-We have not implemented and do not present simulations results with any of the MEA estimators from the papers by Casey and Sadler~\cite{536682,726812}.
-The MEA estimators were supersceeded by the Separable Line Search (SLS) estimators of Sidiropoulos, Swami and Sadler~\cite{Sidiropoulos2005} in 2005.
+We have presented simulation results with the MEA estimators of Casey and Sadler~\cite{536682,726812}.
+The MEA estimators were superseded by the Separable Line Search (SLS) estimators of Sidiropoulos, Swami and Sadler~\cite{Sidiropoulos2005} in 2005.
 Comparison between the MEA estimators and the SLS estimators are presented in~\cite{Sidiropoulos2005} where the superiority of the SLS estimators is shown.
-An enhanced SLS estimator (SLS2-new) was developed in~\cite{Clarkson2007} and this is the estimator we show simulations results for\footnote{One of the SLS estimators (SLS2-all) has excellent statistical accuracy.
+An enhanced SLS estimator (SLS-new) was developed in~\cite{Clarkson2007} and this is the estimator we show simulations results for\footnote{One of the SLS estimators (SLS2-all) has excellent statistical accuracy~\cite{Sidiropoulos2005}.
 It is almost as accurate as the least squares and periodogram estimators as can be see in the simulations in~\cite{Clarkson2007} and~\cite{McKilliam2007}.
-However, it is extremely computationally complex, requiring atleast $O(N^3)$ operations to compute.
+However, it is extremely computationally complex, requiring at least $O(N^3)$ operations.
 We cannot present simulation results for this estimator when $N=1200$ for example.
-This estimator is discussed and compared in~\cite{Sidiropoulos2005} and also~\cite{Clarkson2007,McKilliam2007}.}.
+The SLS2-all estimator is discussed and compared in~\cite{Sidiropoulos2005} and also~\cite{Clarkson2007,McKilliam2007}.}.
 Even this enhanced SLS estimator performs poorly when compared to the least squares and periodogram estimators.
@@
 As the noise variance increases the \emph{clairvoyant} Cramer-Rao bound and the predicted asymptotic variances diverge.
-As expected, the behaviour of the estimators themselves follows the asymptotic theory, rather than the \emph{clairvoyant} Cramer-Rao bound.
+As expected, the behaviour of the estimators measured by simulation follows the asymptotic theory, rather than the \emph{clairvoyant} Cramer-Rao bound.
 
-Observe that in the left hand plot of Figure~\ref{plot:crb} the quantised periodogram estimator with coarse quantisation $q = 1.5 f_{\text{max}}$ (triangles) actually performs well for sufficently small noise variance.
-It appears that the fixed integers $s_1,\dots,s_N$ generated in this simulation favour the quantised periodogram estimator.
-This is one motivation for performing simulations with integers $s_1,\dots,s_N$ generated randomly for each Monte-Carlo trial, rather than choosing a single fixed set of integers.
+Observe that in the left hand plot of Figure~\ref{plot:crb} the quantized periodogram estimator with coarse quantization $q = 1.5 f_{\text{max}}$ (triangles) actually performs well for sufficiently small noise variance.
+It appears that the fixed integers $s_1,\dots,s_N$ generated in this simulation favour the quantized periodogram estimator.
+This is one motivation for performing simulations with integers $s_1,\dots,s_N$ generated randomly for each Monte-Carlo trial.
 It avoids the scenario where a particular favourable set $s_1,\dots,s_N$ is chosen by chance.
@@
 \textbf{Response:}
-We repond to each of the questions of the reviewer in turn:
+We respond to each of the questions of the reviewer in turn.
 
@@
 
-``Given the approximate maximiser $\widetilde{f}^\prime$ we apply the Newton-Raphson method to find a nearby stationary point of the periodogram $I_y$.  This proceeds by the iteration
+``Given the approximate maximiser $\widetilde{f}^\prime$ we apply the Newton-Raphson method~\cite{Ypma_historical_newton_raphson_1995} to find a nearby stationary point of the periodogram $I_y$.
+This proceeds by the iteration
 \[
@@
 Each iteration requires $O(N)$ operations and so the total number of operations required by the Newton-Raphson procedure has order $O(N)$.
-In practice we find that the computational cost of the Newton-Raphson procedure is dominated by the cost of the chirp-z or fast Fourier transforms.''
+In practice we find that the computational cost of the Newton-Raphson procedure is negligible when compared to the cost of the chirp-z or fast Fourier transforms.''
 
@@
 \item \textbf{Is the quantized periodogram just used to initialize a Newton-Raphson procedure?}
-Yes.  As described in the paper BLERG
+Yes.  In the paper the following text is included in Section~\zref{sec:quant-peri}:
+
+``Let $\widetilde{f}^\prime$ be the maximiser of $I_z$ over the grid $G$, that is,
+\[
+\widetilde{f}^\prime = \arg\max_{f \in G} I_z(f).
+\]
+If the quantization parameter $q$ is sufficiently large we expect that $I_z$ is close to $I_y$ and so $\widetilde{f}^\prime$ will be close to $\widetilde{f}$.
+Thus, if $q$ is sufficiently large and the distance $D$ is sufficiently small we expect $\widetilde{f}^\prime$ to be a good approximation to $\hat{f}$.
+An iterative procedure %or Brent's method~\cite{Brent_opt_no_derivs_1973}
+can then be used to compute $\hat{f}$ by maximising $I_y$ starting from $\widetilde{f}^\prime$.
+Section~\zref{sec:comp-compl} describes an implementation of the Newton-Raphson optimisation procedure for this purpose.''
 
@@
 initialization of the NR procedure?}
-Yes.
 The estimators run in the benchmarks in Figure~\zref{plot:benchmark} are exactly the same as those in the simulations in Figure~\zref{plot:multipleN}.
-For the periodogram and quantised peridogram estimators this includes the Newton-Raphson proceedure.
+For the periodogram and quantized periodogram estimators this includes the Newton-Raphson procedure.
 
@@
 \textbf{Response:}
-Thankyou for spotting this error.  A similar error was fixed in equation for $I_z$ directly above this.
+Thank you for spotting this error.  A similar error was fixed in equation for $I_z$ directly above this.
 
@@
 We have also removed references [33] and [34] from the original manuscript.
-We feel that the all 35 references in the new version of the paper are highly relevant.
+We feel that all 35 references in the new version of the paper are highly relevant.
 The paper would be weakened without them.
@@
 (2)]{Sidiropoulos2005}.
-Suppose that $T_{\text{max}} \geq 2T_{\text{min}}$ and that the true period $T_0$ happened to be contained in the nonempty interval $(2T_{\text{min}}, T_0)$.
+To  see why this is necessary, suppose that $T_{\text{max}} \geq 2T_{\text{min}}$ and that the true period $T_0$ happened to be contained in the nonempty interval $(2T_{\text{min}}, T_{\text{max}})$.
 Put $T^\prime = T_0/2 \in (T_{\text{min}}, T_{\text{max}})$ and observe that
@@
 where $r_n = 2 s_n \in \ints$.
-Since the integers are not known both $T_0$ and $T^\prime$ would describe the observations equally well.
+Because the integers $s_1,\dots,s_N$ (respectively $r_1,\dots,r_N$) are not known both $T_0$ and $T^\prime$ would describe the observations equally well.
 This problem is averted only if we suppose that $T_0 \in (T_{\text{min}}, T_{\text{max}})$ and that  $T_{\text{max}} < 2T_{\text{min}}$.
@@
 \textbf{Response:}
-Yes.  The terms inside the sum on the left and right are conjugates of one another.  That is,
+Yes.  The terms inside the sum on the left and right are conjugates.  That is,
 \[
@@
 \item\textbf{Comment:}
-Newton-Rapson iteration to further improve the frequency estimation is advocated in Sec. II. Can a reference supporting this statement be added?
+Newton-Raphson iteration to further improve the frequency estimation is advocated in Sec. II. Can a reference supporting this statement be added?
 \\
@@
 Please see our response to comment~\ref{com:rev1:nr} of reviewer 1.
+The reference~\cite{Ypma_historical_newton_raphson_1995} that gives an overview of the Newton-Raphson method has also been included.
 
@@
 \textbf{Response:}
-The quantised periodgram estimator does not suffer a significant performance loss with respect to the classical LS and periodogram estimators in any scenario we have found.
-Observe in Figure 3 that the solid dots ($q=5 f_{\text{max}}$) the crosses (LS), the circles (periodogram) are all close for $N=30,1200$ and $\mu=1,10$.
-Be aware that the solid lines indicates the asymptotic theory ($N\to\infty$)~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est} for the periodogram and LS estimators and not the measured accuracy for finite $N$.
-As $N\to\infty$, the MSE of the periogram and LS estimators will converge to the solid and dashed curves.
+The quantized periodogram estimator does not suffer a significant performance loss with respect to the classical LS and periodogram estimators in any scenario we have found.
+Observe in Figure 3 that the dots ($q=5 f_{\text{max}}$), the crosses (LS), and the circles (periodogram) are all close for $N=30,1200$ and $\mu=1,10$.
+The solid and dashed lines indicate the asymptotic theory ($N\to\infty$)~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est} for the periodogram and LS estimators and not the measured MSE for finite $N$.
+As $N\to\infty$, the MSE of the periodogram and LS estimators will converge to the solid and dashed curves.
+
+In general we don't expect there to exist an estimator accurate for small $N$ and large noise variance.
 

### 2014-06-15 paper.tex
added final versions.

@@
 %\documentclass[a4paper,10pt]{article}
-%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+%\documentclass[draftcls, onecolumn, 10pt]{IEEEtran}
 \documentclass[10pt,twocolumn,twoside]{IEEEtran}
@@
 If the distance $D$ is sufficiently small, then $\widetilde{f}$ is a good approximation to the maximiser $\hat{f}$.
-An iterative procedure, such as the Newton-Raphson method, %or Brent's method~\cite[Ch.~5]{Brent_opt_no_derivs_1973}
+An iterative procedure, such as the Newton-Raphson method~\cite{Ypma_historical_newton_raphson_1995}, %or Brent's method~\cite[Ch.~5]{Brent_opt_no_derivs_1973}
 can then be used to compute $\hat{f}$ starting from $\widetilde{f}$.
@@
 
-Given the approximate maximiser $\widetilde{f}^\prime$ we apply the Newton-Raphson method to find a nearby stationary point of the periodogram $I_y$.  This proceeds by the iteration
+Given the approximate maximiser $\widetilde{f}^\prime$ we apply the Newton-Raphson method~\cite{Ypma_historical_newton_raphson_1995} to find a nearby stationary point of the periodogram $I_y$.
+This proceeds by the iteration
 \[
@@
 Each iteration requires $O(N)$ operations and so the total number of operations required by the Newton-Raphson procedure has order $O(N)$.
-In practice we find that the computational cost of the Newton-Raphson procedure negligible when compared to the cost of the chirp-z or fast Fourier transforms.
+In practice we find that the computational cost of the Newton-Raphson procedure is negligible when compared to the cost of the chirp-z or fast Fourier transforms.
 
@@
   cells={anchor=west},
-  font=\footnotesize,
+  font=\footnotesize
 },

### 2014-07-29 paper.tex
minor

@@
 By sparse it is meant that only a potentially small unknown subset of the process is observed.
-By noisy it is meant that the subset that is observed, is observed with some error, or noise.
+By noisy it is meant that the subset that is observed, is observed with error, or noise.
 Existing accurate algorithms for estimating the period require $O(N^2)$ operations where $N$ is the number of observations.
@@
 Sadler and Casey~\cite{726812} have previously suggested use of the fast Fourier transform for period estimation.
-However, their estimator requires larger transforms and produces less accurate results than the estimator described here (see Sec.~VI and Fig.
+However, their estimator requires larger transforms and produces less accurate results than the estimator described here (see Sec.~VI in~~\cite{726812} and Fig.
 1-4 in~\cite{726812}).
@@
 \end{align*}
-Observe that $I_z$ is periodic with period $q$ and so quantization should be fine enough that 
+Observe that $I_z$ is periodic with period $q$ and so quantization must be fine enough that 
 \begin{equation}\label{eq:boundaliasingP}
@@
 The previous section shows how the approximate maximiser $\widetilde{f}^\prime$ can be computed using either the chirp z-transform or a fast Fourier transform of length $M$.
-We now describe recommended values for the parameters $q, M$, and $D$.
+We now describe recommended values for the parameters $q$, $M$, and $D$.
 The simulations presented in Section~\ref{sec:simulations} suggest that an effective value for the quantization parameter is $q = 5 f_{\text{max}}$.

### 2014-08-20 response/proof_corrections.tex
added proof corrections

\documentclass[a4paper,10pt]{article}
%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
%\documentclass[journal]{IEEEtran}

\usepackage{../mathbf-abbrevs}
\input{../defs.tex}

\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{pgfplots}
\pgfplotsset{compat=1.8}

\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\usepackage{amsmath,amsfonts,amssymb, amsthm, bm}

%\usepackage[square,comma,numbers,sort&compress]{natbib}

\usepackage{color}
\newcommand{\comment}[1]{\textcolor{red}{#1}}

\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\sinc}{\operatorname{sinc}}
\newcommand{\rect}[1]{\operatorname{rect}\left(#1\right)}

%opening
\title{Fast sparse period estimation: Reviewer responses}
\author{R.~G.~McKilliam, I.~V.~L.~Clarkson and B.~G.~Quinn 
%\thanks{}
}

\renewcommand{\theenumii}{(\alph{enumii})}
\renewcommand{\labelenumii}{\theenumii}

\begin{document}

%\maketitle

\begin{enumerate}

\item Page 1, column 2, \emph{periodogramestimator} appears to be written as all one word.  \\ It should be \emph{periodogram~estimator}.

\item Page 2, column 1, again \emph{periodogramestimator} appears to be written as all one word.  \\ It should be \emph{periodogram~estimator}.

\item Page 2, column 1.  The equation
\[
G = \{ f_{\text{min}} + Dk \mid k = 0, \dots K-1 \}
\]
In the proof version there no space between $Dk$ and the vertical bar $\vert$ and $k$.  It should have some space as above.

\item Page 2, bottom of column 2.  The equation
\[
b_i = \#\{ n \mid \ell_n = i, n = 1,\dots, N\}
\]
In the proof version there no space between $n$ and the vertical bar $\vert$ and $\ell_n$.  It should have some space as above.

\newcommand{\ellmin}{\ell_{\text{min}}}
\newcommand{\ellmax}{\ell_{\text{max}}}

\item Page 3, top of column 1.  The equation
\[
I_z(f) = \babs{ \sum_{i = \ellmin}^{\ellmax} e^{ -2\pi j f i/q} b_i }^2 = \babs{ \sum_{m = 0}^{\ellmax-\ellmin} e^{ -2\pi j f m/q} b_{m+\ellmin} }^2.
\]
does not quite fit on one line in your format.  I don't personally mind this, but if it is preferable you can split this equation over two lines, such as,
\begin{align*}
I_z(f) &= \babs{ \sum_{i = \ellmin}^{\ellmax} e^{ -2\pi j f i/q} b_i }^2 \\
&= \babs{ \sum_{m = 0}^{\ellmax-\ellmin} e^{ -2\pi j f m/q} b_{m+\ellmin} }^2.
\end{align*}

\end{enumerate}

\end{document}

### 2015-04-10 poster/ITR_Poster.tex
added start of ICASSP poster

\documentclass[final,hyperref={pdfpagelabels=false}]{beamer}

\usepackage[orientation=portrait,size=a1,scale=1.1]{beamerposter}

\mode<presentation>{
    \usetheme{ITR}    
}

%%% fonts & encoding %%%
\usepackage{newcent}
\renewcommand{\familydefault}{\sfdefault}
\usepackage[english]{babel}
%\usepackage[latin1]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}

\input{presdefs}

\definecolor{darkgreen}{rgb}{0,0.3,0}
\newcommand{\term}[1]{{\color{darkgreen}\textbf{#1}}}

%%% title, author & contact %%%%%%%%%%%%%%%%%%%%%%%%%
\title[]{Noncoherent least squares estimators \\ of carrier phase and amplitude}
\author{Robby McKilliam, Andr\'{e} Pollok, Bill Cowley, \\ I. Vaughan L. Clarkson and Barry G. Quinn}
\newcommand{\email}{robby.mckilliam@unisa.edu.au}
\newcommand{\worksupportedby}{Supported under the Australian Government's Australian Space Research Program}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
%\frame{

%\vspace{-1.4cm}

\begin{columns}[t] % <-- columns with top alignment
\begin{column}{.49\textwidth} % <-- start the first column

\newcommand{\calC}{{\mathcal C}}
\begin{block}{Received signal model}
Receive $L$ noisy $M$-ary phase-shifted-keyed ($M$-PSK) symbols of the form
\[
y_i = a_0 s_i + w_i, \qquad i = 1, \dots, L.
\]
\begin{itemize}
\item $a_0 = \rho_0 e^{j\theta_0}$ is the unknown carrier phase $\theta_0$ and amplitude $\rho_0$,
\item $s_1,\dots,s_L$ are transmitted $M$-PSK symbols,
\item $w_1,\dots,w_L \in \complex$ are i.i.d. circularly symmetric complex random variables representing noise.
\end{itemize}

Interested in estimating $\rho_0$ and $\theta_0$.
\end{block}

\begin{block}{Least squares estimator}
If all symbols $s_1,\dots,s_L$ are known
\[
\hat{a}_{\text{uc}} = \arg \min_{a \in \complex} \sum_{i = 1}^L \abs{ y_i - a s_i }^2  = \frac{1}{L} \sum_{i = 1}^L y_i s_i^*.
\]
More interested in the practical situation where symbols are not known,
\[
\hat{a} = \arg\min_{a \in \complex} \min_{s_1, \dots, s_L} \sum_{i = 1}^L \abs{ y_i - a s_i }^2.
\]
This estimator can be computed in $O(L\log L)$ operations.
\end{block}

\begin{block}{Theorem (Almost sure convergence)}

Let $R_i \geq 0$ and $\Phi_i \in [-\pi,\pi)$ be real random variables satisfying
%\begin{equation}\label{eq:RiandPhii}
\vspace{-0.15cm}
\[
R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i},
\]
%\end{equation}
and define the continuous function
\[
G(x) = \expect R_1 \cos\sfracpart{ x + \Phi_1}.
\] 
If $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$, then:
\begin{enumerate}
\item \vspace{-0.15cm} $\sfracpart{\hat{\theta} - \theta_0} \rightarrow 0$ almost surely as $L \rightarrow \infty$,
\item \vspace{-0.2cm} $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$,
\end{enumerate}
where $\sfracpart{\cdot}$ takes its argument `modulo $\tfrac{2\pi}{M}$' into $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$.
%\[
%\fracpart{x} = x - \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right).
%\]  
\end{block}

\begin{block}{Theorem (Asymptotic normality)}
Let $f(r,\phi)$ be the joint pdf of $R_1$ and $\Phi_1$, and let
\[
g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
\]
Put $\hat{\lambda}_L = \sfracpart{\hat{\theta} - \theta_0}$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$. %If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then 
The distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance matrix
\[
\left( \begin{array}{cc} 
H^{-2} A & 0 \\
0 & \rho_0^2 B
\end{array} \right)
\]
as $L \rightarrow \infty$, where
\vspace{-0.15cm}
\[
H = G(0) -  2 \sin\big(\tfrac{\pi}{M}\big) \sum_{k = 0}^{M-1} g\big(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}\big),
\]
\vspace{-0.15cm}
\[
A = \expect R_1^2\sin^2\fracpart{\Phi_1}, \;\;\; B = \expect R_1^2 \cos^2\fracpart{\Phi_1} - G^2(0). 
\]

\end{block}

\end{column}
\begin{column}{.49\textwidth} % <-- start the second column

\begin{block}{Simulations}

\begin{figure}
	\centering
		\includegraphics[width=\linewidth]{../code/data/posterplot-2.mps}
		\caption{Phase mean square error for $4$-PSK (QPSK)}
		\label{fig:plotphaseM4}
\end{figure}

\vspace{0.79cm}
\begin{figure}
	\centering
		\includegraphics[width=\linewidth]{../code/data/posterplot-1.mps}
		\caption{Amplitude mean square error for $4$-PSK (QPSK)}
		\label{fig:plotampM4}
\end{figure}

\end{block}

\end{column}
\end{columns}

 %}

\end{document}

### 2015-04-10 poster/build.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+pushd ../code/data
+mpost -interaction=nonstopmode posterplot.mp
+popd
+pdflatex -interaction=nonstopmode ITR_Poster.tex

### 2015-04-10 poster/ITR_Poster.tex
addition to icassp poster

@@
 
+\usepackage{tikz}
+\usetikzlibrary{calc}
+\usepackage{pgfplots}
+\pgfplotsset{compat=1.8}
+\usetikzlibrary{pgfplots.groupplots}
+
 \input{presdefs}
@@
 %%% title, author & contact %%%%%%%%%%%%%%%%%%%%%%%%%
-\title[]{Noncoherent least squares estimators \\ of carrier phase and amplitude}
-\author{Robby McKilliam, Andr\'{e} Pollok, Bill Cowley, \\ I. Vaughan L. Clarkson and Barry G. Quinn}
+\title[]{Fast sparse period estimation}
+\author{Robby.~G.~McKilliam, I.~Vaughan.~L.~Clarkson and Barry.~G.~Quinn}
 \newcommand{\email}{robby.mckilliam@unisa.edu.au}
@@
   \centering 
-		\includegraphics[width=\linewidth]{../code/data/posterplot-2.mps}
-		\caption{Phase mean square error for $4$-PSK (QPSK)}
-		\label{fig:plotphaseM4}
+  \begin{tikzpicture}
+    \selectcolormodel{gray} 
+    \begin{axis}[font=\footnotesize,xmode=log,ymode=log,height=15cm,width=25cm,xlabel={$N$},ylabel={time (ms)},ylabel style={at={(0.055,0.42)}},xlabel style={at={(0.5,0.1)}}, legend style={draw=none,fill=none,legend pos=north west,cells={anchor=west},font=\footnotesize}]
+      \addplot[mark=o,mark options={scale=3}] table {../code/data/PeriodogramBenchmark};
+      \addplot[mark=x,mark options={solid,fill=black,scale=3.6}] table {../code/data/LeastSquaresBenchmark};
+      \addplot[mark=*,mark options={solid,fill=black,scale=1.5}] table {../code/data/QuantisedPeriodogramq4.0Benchmark};
+      \addplot[mark=triangle,mark options={solid,fill=black,scale=2.7}] table {../code/data/QuantisedPeriodogramChirpZq4.0Benchmark};
+      %\addplot[mark=square,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpBenchmark};
+      \addplot[mark=square,color=black,mark options={solid,fill=black,scale=2.4}] table {../code/data/SLS2newBenchmark};
+      \legend{Periodogram, Least squares, Fast Fourier transform, Chirp z-transform, SLS-new}
+   \end{axis}
+  \end{tikzpicture}  
+  \caption{Computation time in milliseconds versus $N$ for the periodogram, least squares, and quantized periodogram estimators computed using the chirp z-transform and a single fast Fourier transform.}\label{plot:benchmark}
 \end{figure}
 
-\vspace{0.79cm}
-\begin{figure}
-	\centering
-		\includegraphics[width=\linewidth]{../code/data/posterplot-1.mps}
-		\caption{Amplitude mean square error for $4$-PSK (QPSK)}
-		\label{fig:plotampM4}
-\end{figure}
+% \begin{figure}
+% 	\centering
+% 		\includegraphics[width=\linewidth]{../code/data/posterplot-2.mps}
+% 		\caption{Phase mean square error for $4$-PSK (QPSK)}
+% 		\label{fig:plotphaseM4}
+% \end{figure}
+
+% \vspace{0.79cm}
+% \begin{figure}
+% 	\centering
+% 		\includegraphics[width=\linewidth]{../code/data/posterplot-1.mps}
+% 		\caption{Amplitude mean square error for $4$-PSK (QPSK)}
+% 		\label{fig:plotampM4}
+% \end{figure}
 

### 2015-04-12 poster/ITR_Poster.tex
minor

--- 2015-04-10 poster/ITR_Poster.tex
@@
 %%% title, author & contact %%%%%%%%%%%%%%%%%%%%%%%%%
-\title[]{Noncoherent least squares estimators \\ of carrier phase and amplitude}
-\author{Robby McKilliam, Andr\'{e} Pollok, Bill Cowley, \\ I. Vaughan L. Clarkson and Barry G. Quinn}
+\title[]{Fast sparse period estimation}
+\author{Robby.~G.~McKilliam, I.~Vaughan.~L.~Clarkson \\ and Barry.~G.~Quinn}
 \newcommand{\email}{robby.mckilliam@unisa.edu.au}
@@
 \begin{block}{Received signal model}
-Receive $L$ noisy $M$-ary phase-shifted-keyed ($M$-PSK) symbols of the form
-\[
-y_i = a_0 s_i + w_i, \qquad i = 1, \dots, L.
-\]
-\begin{itemize}
-\item $a_0 = \rho_0 e^{j\theta_0}$ is the unknown carrier phase $\theta_0$ and amplitude $\rho_0$,
-\item $s_1,\dots,s_L$ are transmitted $M$-PSK symbols,
-\item $w_1,\dots,w_L \in \complex$ are i.i.d. circularly symmetric complex random variables representing noise.
-\end{itemize}
 
-Interested in estimating $\rho_0$ and $\theta_0$.
-\end{block}
-
-\begin{block}{Least squares estimator}
-If all symbols $s_1,\dots,s_L$ are known
-\[
-\hat{a}_{\text{uc}} = \arg \min_{a \in \complex} \sum_{i = 1}^L \abs{ y_i - a s_i }^2  = \frac{1}{L} \sum_{i = 1}^L y_i s_i^*.
-\]
-More interested in the practical situation where symbols are not known,
-\[
-\hat{a} = \arg\min_{a \in \complex} \min_{s_1, \dots, s_L} \sum_{i = 1}^L \abs{ y_i - a s_i }^2.
-\]
-This estimator can be computed in $O(L\log L)$ operations.
-\end{block}
-
-\begin{block}{Theorem (Almost sure convergence)}
-
-Let $R_i \geq 0$ and $\Phi_i \in [-\pi,\pi)$ be real random variables satisfying
-%\begin{equation}\label{eq:RiandPhii}
-\vspace{-0.15cm}
-\[
-R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i},
-\]
-%\end{equation}
-and define the continuous function
-\[
-G(x) = \expect R_1 \cos\sfracpart{ x + \Phi_1}.
-\] 
-If $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$, then:
-\begin{enumerate}
-\item \vspace{-0.15cm} $\sfracpart{\hat{\theta} - \theta_0} \rightarrow 0$ almost surely as $L \rightarrow \infty$,
-\item \vspace{-0.2cm} $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$,
-\end{enumerate}
-where $\sfracpart{\cdot}$ takes its argument `modulo $\tfrac{2\pi}{M}$' into $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$.
-%\[
-%\fracpart{x} = x - \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right).
-%\]  
-\end{block}
-
-\begin{block}{Theorem (Asymptotic normality)}
-Let $f(r,\phi)$ be the joint pdf of $R_1$ and $\Phi_1$, and let
-\[
-g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
-\]
-Put $\hat{\lambda}_L = \sfracpart{\hat{\theta} - \theta_0}$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
-%If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then
-The distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance matrix
-\[
-\left( \begin{array}{cc} 
-H^{-2} A & 0 \\
-0 & \rho_0^2 B
-\end{array} \right)
-\]
-as $L \rightarrow \infty$, where
-\vspace{-0.15cm}
-\[
-H = G(0) -  2 \sin\big(\tfrac{\pi}{M}\big) \sum_{k = 0}^{M-1} g\big(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}\big),
-\]
-\vspace{-0.15cm}
-\[
-A = \expect R_1^2\sin^2\fracpart{\Phi_1}, \;\;\; B = \expect R_1^2 \cos^2\fracpart{\Phi_1} - G^2(0). 
-\]
+{
+\def\vertgap{2}
+\def\ph{0.4}
+\def\T{1.1}
+\newcommand{\raxis}{\draw[->] (-0.25,0) -- (8,0) node[above] {$\reals$}; \draw (0,-0.06)-- node[below] {$0$} (0,0.06) }
+\newcommand{\pulse}[1]{ \draw[->,>=latex] (#1,0) -- (#1,1) }
+\newcommand{\pulsewithnode}[2]{ \draw[->,>=latex] (#1,0) -- node[right] {#2} (#1,1) }
+\begin{figure}[t]
+	\centering
+\begin{tikzpicture}[font=\small,scale=3]
+    %\draw[very thin,color=gray] (-0.1,-1.1) grid (3.9,3.9);
+\begin{scope}
+    \raxis;
+    \foreach \k in {0,1,...,6} { \k, \pulse{\T*\k+\ph}; }
+    \draw[<->] (\T*2+\ph,1.2) -- node[above] {$T_0$} (\T*3+\ph,1.2) ;
+    \draw[<->] (0.0,1.2) -- node[above] {$\theta_0$} (\ph,1.2) ; 
+%    \draw[->] (-0.25,-2*\vertgap) -- (8,-2*\vertgap) node[above] {$\reals$};
+\end{scope}
+\begin{scope}[yshift=-\vertgap cm]
+  \raxis;
+  \node[below] at (\T*1+\ph-0.3,0) {$y_1$}; \pulsewithnode{\T*1+\ph-0.3}{$s_1=1$}; \draw[->] (\T*1+\ph,1.2) -- node[above] {$w_1$} (\T*1+\ph-0.3,1.2) ; 
+  \node[below] at (\T*3+\ph+0.5,0) {$y_2$}; \pulsewithnode{\T*3+\ph+0.5}{$s_2=3$}; \draw[->] (\T*3+\ph,1.2) -- node[above] {$w_2$} (\T*3+\ph+0.5,1.2) ;
+  \node[below] at (\T*6+\ph-0.4,0) {$y_3$}; \pulsewithnode{\T*6+\ph-0.4}{$s_3=6$};\draw[->] (\T*6+\ph,1.2) -- node[above] {$w_3$} (\T*6+\ph-0.4,1.2) ;
+\end{scope}
+\end{tikzpicture} 
+%		\includegraphics{figs/figures-1.mps}
+		\caption{Sparse noisy observations of a periodic sequence.
+The original sequence~\eqref{eq:periodicprocess} is shown on the top and sparse noisy observations according to~\eqref{eq:sigmodel} are shown on the bottom.
+In this example the observed subset is represented by integers $s_1=1$, $s_2=3$ and $s_3=6$.}
+		\label{fig_stat_model}
+\end{figure}
+}
 
@@
 
+\begin{figure}[p] 
+  \centering 
+  \begin{tikzpicture} 
+\begin{groupplot}[
+    group style={ 
+        group name=my plots,
+        group size=1 by 2,
+        %ylabels at=edge left,
+        xlabels at=edge bottom,
+        xticklabels at=edge bottom,
+        vertical sep=0pt
+    },
+    ylabel style={at={(0.06,0.5)}},
+    footnotesize,
+    width=25cm,
+    height=15cm,
+    tickpos=left,
+    %ytick align=outside,
+    %xtick align=outside,
+    xmode=log,
+    ymode=log,
+    xmin=11e-5,
+    xmax=0.4
+%    enlarge x limits=false 
+]
+
+\nextgroupplot[
+ylabel={MSE},
+ymin=2e-10
+]
+
+      %\addplot[mark=none,color=black,dashed] table {../code/data/NormalisedLLSCLTN30geom10};
+
+            \addplot[mark=*,mark options={scale=1.5},color=cyan] table {../code/data/QuantisedPeriodogramFFTN30q1.5geom10};
+      \addplot[mark=square*,mark options={scale=2.5},color=green] table {../code/data/QuantisedPeriodogramFFTN30q5.0geom10};
+\addplot[mark=triangle*,mark options={scale=3},color=blue] table {../code/data/SLS2newN30geom10};
+\addplot[mark=none,color=black] table {../code/data/PeriodogramCLTN30geom10}; 
+      \addplot[mark=*,mark options={scale=1.75},color=red] table {../code/data/PeriodogramN30geom10};
+%       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {../code/data/NormalisedSamplingLLSN30geom10}; 
+  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN30geom10};
+
+%      \addplot[mark=none,dashed,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q2};
+ %     \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200};
+ %     \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200}; 
+ %     \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/PeriodogramN200};
+ %     \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q3};
+ %     \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q10};
+%      \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/NormalisedSamplingLLSN200}; 
+       \node[font=\footnotesize] at (axis cs:0.2,4e-10) {$N=30$};
+
+\nextgroupplot[
+xlabel={$\sigma^2$},
+xlabel style={at={(0.5,0.1)}},
+ylabel={MSE},
+ymin=2e-15,
+ymax=8e-1,
+legend style={
+  legend columns=1,
+  draw=none,
+  %fill=none,
+  legend pos=north west,
+  %at={(0.75,1)},
+  cells={anchor=west},
+  font=\footnotesize
+}
+%ytickten={-1,-3,-5,-7,-9,-11,-13}
+]
+
+\addplot[mark=*,mark options={scale=1.5},color=cyan] table {../code/data/QuantisedPeriodogramFFTN1200q1.5geom10};
+ \addplot[mark=square*,mark options={scale=2.5},color=green] table {../code/data/QuantisedPeriodogramFFTN1200q5.0geom10};
+       \addplot[mark=triangle*,mark options={scale=3},color=blue] table {../code/data/SLS2newN1200geom10};
+\addplot[mark=none,color=black] table {../code/data/PeriodogramCLTN1200geom10};
+      %\addplot[mark=none,color=black,dashed] table {../code/data/NormalisedLLSCLTN1200geom10};
+       \addplot[mark=*,mark options={scale=1.75},color=red] table {../code/data/PeriodogramN1200geom10};
+%       \addplot[color=gray,mark=x,only marks,mark options={solid,color=black,fill=black,scale=1.3}] table {../code/data/NormalisedSamplingLLSN1200geom10};
+ %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1100geom10};
+      %\addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1100q3};
+ %     \legend{Periodogram Theory, Least squares theory, Quantised periodogram, Periodogram, Least squares}
+     %\legend{Periodogram theory, $q=\tfrac{3}{2}T_{\text{max}}$, $q=4T_{\text{max}}$, Periodogram, Least squares, SLS2-adj}
+     \node[font=\footnotesize] at (axis cs:0.32,2e-15) [anchor=south east] {$N=1200$};
+   %  \node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=1200$};
+
+\legend{Quantised periodogram $q=1.5$, Quantised periodogram $q=5$,Seperable line search,Periodogram Theory,Periodogram}
+
+\end{groupplot}
+
+  \end{tikzpicture}  
+  \caption{Mean square error versus noise variance $\sigma^2$.  Trials are generated so that only $10\%$ of periods are observed on average. }\label{plot:multipleN}
+\end{figure}
+
 \begin{figure}
   \centering 
-		\includegraphics[width=\linewidth]{../code/data/posterplot-2.mps}
-		\caption{Phase mean square error for $4$-PSK (QPSK)}
-		\label{fig:plotphaseM4}
+  \begin{tikzpicture}
+%    \selectcolormodel{gray} 
+    \begin{axis}[font=\footnotesize,xmode=log,ymode=log,height=15cm,width=25cm,xlabel={$N$},ylabel={time (ms)},ylabel style={at={(0.055,0.42)}},xlabel style={at={(0.5,0.1)}}, legend style={draw=none,fill=none,legend pos=north west,cells={anchor=west},font=\footnotesize},xmin=7,xmax=36000]
+      \addplot[mark=*,mark options={scale=1.75},color=red] table {../code/data/PeriodogramBenchmark};
+      %\addplot[mark=triangle*,mark options={scale=3},color=cyan] table {../code/data/LeastSquaresBenchmark};
+      \addplot[mark=square*,mark options={scale=2.5},color=green] table {../code/data/QuantisedPeriodogramq4.0Benchmark};
+      %\addplot[mark=triangle*,mark options={scale=3},color=cyan] table {../code/data/QuantisedPeriodogramChirpZq4.0Benchmark};
+      %\addplot[mark=square,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpBenchmark};
+      \addplot[mark=triangle*,mark options={scale=3},color=blue] table {../code/data/SLS2newBenchmark};
+      %\legend{Periodogram, Least squares, Fast Fourier transform, Chirp z-transform, SLS-new}
+      \legend{Periodogram, Quantised periodogram $q=5$, Seperable line search}
+   \end{axis}
+  \end{tikzpicture}  
+  \caption{Computation time in milliseconds versus data length $N$.}\label{plot:benchmark}
 \end{figure}
 
-\vspace{0.79cm}
-\begin{figure}
-	\centering
-		\includegraphics[width=\linewidth]{../code/data/posterplot-1.mps}
-		\caption{Amplitude mean square error for $4$-PSK (QPSK)}
-		\label{fig:plotampM4}
-\end{figure}
+% \begin{figure}
+% 	\centering
+% 		\includegraphics[width=\linewidth]{../code/data/posterplot-2.mps}
+% 		\caption{Phase mean square error for $4$-PSK (QPSK)}
+% 		\label{fig:plotphaseM4}
+% \end{figure}
+
+% \vspace{0.79cm}
+% \begin{figure}
+% 	\centering
+% 		\includegraphics[width=\linewidth]{../code/data/posterplot-1.mps}
+% 		\caption{Amplitude mean square error for $4$-PSK (QPSK)}
+% 		\label{fig:plotampM4}
+% \end{figure}
 

### 2015-04-12 code/benchmark.scala

@@ -26,8 +26,7 @@ val p = 1.0/3.0 //parameter of geometric random variable for generating sparse n
 def discretedist = new Geometric(p) //the discrete distribution for generating different sN - sN-1
 def noisedist = new Gaussian(0,0.01) //the noise distribution we use with variance v
 
-//time used for benchmark warmup in nanoseconds.  Benchmark runtime is approximately twice this.
-val MIN_BENCH_DURATION : Long = 5000000000L; // (10 secs)  
+
 
 val Ns = List(10,14,20,30,40,50,60,80,100,120,150,200,250,300,400,600,900,1200,1500,2000,2600,3500,4700,6000,7800,10000,13000,18000,25000)
 //val Ns = List(10,14,20,30,40,50,60,80,100,120,150,200,250,300,400,600,900,1200,1500,2000,2600,3500,4700,6000,7800,10000,13000,18000,25000,35000,50000,70000)

### 2015-04-14 poster/ITR_Poster.tex
added poster

@@
 \usepackage{pgfplots}
-\pgfplotsset{compat=1.8}
+%\pgfplotsset{compat=1.8}
 \usetikzlibrary{pgfplots.groupplots}
+\usetikzlibrary{arrows,arrows.new}
 
@@
 \newcommand{\email}{robby.mckilliam@unisa.edu.au}
-\newcommand{\worksupportedby}{Supported under the Australian Government's Australian Space Research Program}
+\newcommand{\worksupportedby}{Supported by Australian Research Council Discovery Project DP130102228}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\newcommand{\calC}{{\mathcal C}}
-\begin{block}{Received signal model}
+\newcommand{\Tmin}{T_{\text{min}}}
+\newcommand{\Tmax}{2T_{\text{min}}}
+
+\begin{block}{Sparse noisy period estimation}
+
+Observations of the form
+\[
+y_n = T_0 s_n + \theta_0 + w_n, \qquad n = 1,\dots,N
+\]
+where
+\begin{itemize}
+\item $T_0 \in (\Tmin, \Tmax) \subset \reals$ is the unknown period,
+\item $\theta_0 \in \reals$ is an unknown phase,
+\item $w_1,\dots,w_N \in \reals$ represent noise,
+\item $s_1,\dots,s_N \in \ints$ represent the subset of periods observed.
+\end{itemize}
+Aim to estimate $T_0$.
 
@@
 \def\T{1.1}
-\newcommand{\raxis}{\draw[->] (-0.25,0) -- (8,0) node[above] {$\reals$}; \draw (0,-0.06)-- node[below] {$0$} (0,0.06) }
-\newcommand{\pulse}[1]{ \draw[->,>=latex] (#1,0) -- (#1,1) }
-\newcommand{\pulsewithnode}[2]{ \draw[->,>=latex] (#1,0) -- node[right] {#2} (#1,1) }
+
+\newcommand{\raxis}{\draw[-> new,arrow head=2mm] (-0.25,0) -- (8,0) node[above] {$\reals$}; \draw (0,-0.06)-- node[below] {$0$} (0,0.06) }
+\newcommand{\pulse}[1]{ \draw[-latex new,arrow head=4mm] (#1,0) -- (#1,1) }
+\newcommand{\pulsewithnode}[2]{ \draw[-latex new,arrow head=4mm] (#1,0) -- node[right] {#2} (#1,1) }
 \begin{figure}[t]
@@
     \foreach \k in {0,1,...,6} { \k, \pulse{\T*\k+\ph}; }
-    \draw[<->] (\T*2+\ph,1.2) -- node[above] {$T_0$} (\T*3+\ph,1.2) ;
-    \draw[<->] (0.0,1.2) -- node[above] {$\theta_0$} (\ph,1.2) ; 
+    \draw[< new-> new, arrow head=2mm] (\T*2+\ph,1.2) -- node[above] {$T_0$} (\T*3+\ph,1.2);
+    \draw[< new-> new, arrow head=2mm] (0.0,1.2) -- node[above] {$\theta_0$} (\ph,1.2); 
 %    \draw[->] (-0.25,-2*\vertgap) -- (8,-2*\vertgap) node[above] {$\reals$};
@@
   \raxis;
-  \node[below] at (\T*1+\ph-0.3,0) {$y_1$}; \pulsewithnode{\T*1+\ph-0.3}{$s_1=1$}; \draw[->] (\T*1+\ph,1.2) -- node[above] {$w_1$} (\T*1+\ph-0.3,1.2) ; 
-  \node[below] at (\T*3+\ph+0.5,0) {$y_2$}; \pulsewithnode{\T*3+\ph+0.5}{$s_2=3$}; \draw[->] (\T*3+\ph,1.2) -- node[above] {$w_2$} (\T*3+\ph+0.5,1.2) ;
-  \node[below] at (\T*6+\ph-0.4,0) {$y_3$}; \pulsewithnode{\T*6+\ph-0.4}{$s_3=6$};\draw[->] (\T*6+\ph,1.2) -- node[above] {$w_3$} (\T*6+\ph-0.4,1.2) ;
+  \node[below] at (\T*1+\ph-0.3,0) {$y_1$}; \pulsewithnode{\T*1+\ph-0.3}{$s_1=1$}; \draw[-> new arrow head=2mm] (\T*1+\ph,1.2) -- node[above] {$w_1$} (\T*1+\ph-0.3,1.2); 
+  \node[below] at (\T*3+\ph+0.5,0) {$y_2$}; \pulsewithnode{\T*3+\ph+0.5}{$s_2=3$}; \draw[-> new,arrow head=2mm] (\T*3+\ph,1.2) -- node[above] {$w_2$} (\T*3+\ph+0.5,1.2);
+  \node[below] at (\T*6+\ph-0.4,0) {$y_3$}; \pulsewithnode{\T*6+\ph-0.4}{$s_3=6$};\draw[-> new,arrow head=2mm] (\T*6+\ph,1.2) -- node[above] {$w_3$} (\T*6+\ph-0.4,1.2);
 \end{scope}
@@
 %		\includegraphics{figs/figures-1.mps}
-		\caption{Sparse noisy observations of a periodic sequence.
-The original sequence~\eqref{eq:periodicprocess} is shown on the top and sparse noisy observations according to~\eqref{eq:sigmodel} are shown on the bottom.
-In this example the observed subset is represented by integers $s_1=1$, $s_2=3$ and $s_3=6$.}
+	%	\caption{Sparse noisy observations of a periodic sequence.}
 		\label{fig_stat_model}
@@
 }
+
+\end{block}
+
+\begin{block}{Applications}
+\begin{itemize}
+\item Pulse interval estimation in electronic support%~\cite{Clarkson_thesis,clarkson_estimate_period_pulse_train_1996,Hauochan_pri_2012}.
+\item Synchronisation for telecommunications%~\cite{Fogel1989_bit_synch_zero_crossings,Sidiropoulos2005,5621928}.
+\item Spike interval estimation in neurology%~\cite{Brillinger_spike_trains_1988}.
+\item Detection of anomalies in internet traffic%~\cite{He_detecting_periodic_patterns_in_internet_2006,5585849,5947313}.
+\end{itemize}
+Connected with Diophantine approximation, lattice reduction, and the study of prime numbers.%~\cite{Clarkson_thesis,CaseySadler_primes_2013}.
+\end{block}
+
+\newcommand{\abs}[1]{\left\vert #1 \right\vert}
+\newcommand{\sabs}[1]{\vert #1 \vert}
+\newcommand{\babs}[1]{\big\vert #1 \big\vert}
+
+\begin{block}{The periodogram estimator}
+An accurate estimator is $\hat{T} = 1/\hat{f}$ where $\hat{f}$ maximises the \emph{periodogram}
+\[
+I_y(f) = \babs{ \sum_{n=1}^N e^{ 2\pi j f y_n} }^2.
+\]
+over the interval $(1/\Tmax, 1/\Tmin)$.  
+\begin{itemize}
+\item Requires $O(N^2)$ operations to compute. 
+\item Observations $y_1,\dots,y_N$ occur in the exponent preventing use of fast transforms.
+\end{itemize}
+\end{block}
+
+\newcommand{\ellmin}{\ell_{\text{min}}}
+\newcommand{\ellmax}{\ell_{\text{max}}}
+
+\begin{block}{A fast approximation}
+
+Compute quantized observations
+\[
+z_n =  \ell_n/q, \qquad n = 1,\dots,N
+\] 
+where $\ell_n = \round{qy_n} \in \ints$ and maximise
+\[
+I_z(f) = \babs{ \sum_{n=1}^N e^{ 2\pi j f z_n} }^2 = \babs{ \sum_{i = \ellmin}^{\ellmax} e^{2\pi j f i/q} b_i }^2
+\]
+where $\ellmin$ and $\ellmax$ are the minimum and maximum value of $\ell_n$ and
+\[
+b_i= \#\{ n \mid \ell_n = i, n = 1,\dots, N\}
+\]
+where $\#$ indicates the number of elements in a set.
+
+\begin{itemize}
+\item Can be computed using $O(N\log N)$ operations by fast Fourier transform or chirp-z transform.
+\end{itemize}
 
@@
     },
-    ylabel style={at={(0.06,0.5)}},
+    ylabel style={at={(0.085,0.4)}},
     footnotesize,
     width=25cm,
-    height=15cm,
+    height=15.83cm,
     tickpos=left,
@@
 
-      %\addplot[mark=none,color=black,dashed] table {../code/data/NormalisedLLSCLTN30geom10};
+      \addplot[mark=none,color=black,dashed] table {../code/data/NormalisedLLSCLTN30geom10};
 
@@
 \addplot[mark=none,color=black] table {../code/data/PeriodogramCLTN30geom10}; 
+ \addplot[mark=o,mark options={scale=2.5},color=black, thick] table {../code/data/NormalisedSamplingLLSN30geom10}; 
       \addplot[mark=*,mark options={scale=1.75},color=red] table {../code/data/PeriodogramN30geom10};
-%       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {../code/data/NormalisedSamplingLLSN30geom10}; 
+%      
   %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN30geom10};
@@
 xlabel={$\sigma^2$},
-xlabel style={at={(0.5,0.1)}},
+xlabel style={at={(0.5,0.13)}},
 ylabel={MSE},
@@
        \addplot[mark=triangle*,mark options={scale=3},color=blue] table {../code/data/SLS2newN1200geom10};
+\addplot[mark=none,color=black,dashed] table {../code/data/NormalisedLLSCLTN1200geom10};
 \addplot[mark=none,color=black] table {../code/data/PeriodogramCLTN1200geom10};
-      %\addplot[mark=none,color=black,dashed] table {../code/data/NormalisedLLSCLTN1200geom10};
+       \addplot[mark=o,mark options={scale=2.5},color=black, thick] table {../code/data/NormalisedSamplingLLSN1200geom10};
        \addplot[mark=*,mark options={scale=1.75},color=red] table {../code/data/PeriodogramN1200geom10};
-%       \addplot[color=gray,mark=x,only marks,mark options={solid,color=black,fill=black,scale=1.3}] table {../code/data/NormalisedSamplingLLSN1200geom10};
  %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1100geom10};
@@
 
-\legend{Quantised periodogram $q=1.5$, Quantised periodogram $q=5$,Seperable line search,Periodogram Theory,Periodogram}
+\legend{Quantised periodogram $q=1.5$, Quantised periodogram $q=5$,Seperable line search,Least squares Theory,Periodogram Theory, Least squares,Periodogram}
 
@@
 %    \selectcolormodel{gray} 
-    \begin{axis}[font=\footnotesize,xmode=log,ymode=log,height=15cm,width=25cm,xlabel={$N$},ylabel={time (ms)},ylabel style={at={(0.055,0.42)}},xlabel style={at={(0.5,0.1)}}, legend style={draw=none,fill=none,legend pos=north west,cells={anchor=west},font=\footnotesize},xmin=7,xmax=36000]
+    \begin{axis}[font=\footnotesize,xmode=log,ymode=log,height=15cm,width=25cm,xlabel={$N$},ylabel={time (ms)},ylabel style={at={(0.085,0.42)}},xlabel style={at={(0.5,0.13)}}, legend style={draw=none,fill=none,legend pos=north west,cells={anchor=west},font=\footnotesize},xmin=7,xmax=36000]
       \addplot[mark=*,mark options={scale=1.75},color=red] table {../code/data/PeriodogramBenchmark};
-      %\addplot[mark=triangle*,mark options={scale=3},color=cyan] table {../code/data/LeastSquaresBenchmark};
-      \addplot[mark=square*,mark options={scale=2.5},color=green] table {../code/data/QuantisedPeriodogramq4.0Benchmark};
+      \addplot[mark=o,mark options={scale=2.5},color=black, thick] table {../code/data/LeastSquaresBenchmark};
       %\addplot[mark=triangle*,mark options={scale=3},color=cyan] table {../code/data/QuantisedPeriodogramChirpZq4.0Benchmark};
@@
       \addplot[mark=triangle*,mark options={scale=3},color=blue] table {../code/data/SLS2newBenchmark};
+      \addplot[mark=square*,mark options={scale=2.5},color=green] table {../code/data/QuantisedPeriodogramq4.0Benchmark};
       %\legend{Periodogram, Least squares, Fast Fourier transform, Chirp z-transform, SLS-new}
-      \legend{Periodogram, Quantised periodogram $q=5$, Seperable line search}
+      \legend{Periodogram, Least squares, Seperable line search, Quantised periodogram $q=5$}
    \end{axis}
@@
 
+\begin{block}{References}
+
+\nocite{clarkson_estimate_period_pulse_train_1996,Sidiropoulos2005,5585849,Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}
+
+\bibliographystyle{../IEEEbib}
+\footnotesize
+\bibliography{../bib}
+
+\end{block}
+
 \end{column}

### 2015-04-14 poster/build.sh

@@ -1,5 +1,3 @@
 #!/bin/bash
-pushd ../code/data
-mpost -interaction=nonstopmode posterplot.mp
-popd
 pdflatex -interaction=nonstopmode ITR_Poster.tex
+bibtex ITR_Poster.aux

### 2015-04-14 poster/ITR_Poster.tex
minor edits to poster

@@
 \title[]{Fast sparse period estimation}
-\author{Robby.~G.~McKilliam, I.~Vaughan.~L.~Clarkson \\ and Barry.~G.~Quinn}
+\author{Robby~G.~McKilliam, I.~Vaughan~L.~Clarkson \\ and Barry~G.~Quinn}
 \newcommand{\email}{robby.mckilliam@unisa.edu.au}
@@
 \begin{itemize}
-\item Pulse interval estimation in electronic support%~\cite{Clarkson_thesis,clarkson_estimate_period_pulse_train_1996,Hauochan_pri_2012}.
-\item Synchronisation for telecommunications%~\cite{Fogel1989_bit_synch_zero_crossings,Sidiropoulos2005,5621928}.
-\item Spike interval estimation in neurology%~\cite{Brillinger_spike_trains_1988}.
-\item Detection of anomalies in internet traffic%~\cite{He_detecting_periodic_patterns_in_internet_2006,5585849,5947313}.
+\item Pulse interval estimation in electronic support.%~\cite{Clarkson_thesis,clarkson_estimate_period_pulse_train_1996,Hauochan_pri_2012}.
+\item Synchronisation for telecommunications.%~\cite{Fogel1989_bit_synch_zero_crossings,Sidiropoulos2005,5621928}.
+\item Spike interval estimation in neurology.%~\cite{Brillinger_spike_trains_1988}.
+\item Detection of anomalies in internet traffic.%~\cite{He_detecting_periodic_patterns_in_internet_2006,5585849,5947313}.
 \end{itemize}
@@
 \begin{block}{The periodogram estimator}
-An accurate estimator is $\hat{T} = 1/\hat{f}$ where $\hat{f}$ maximises the \emph{periodogram}
+An accurate estimator is $\widehat{T} = 1/\widehat{f}$ where $\widehat{f}$ maximises the \emph{periodogram}
 \[
@@
 
-Compute quantized observations
+Compute quantised observations
 \[
@@
 \]
-where $\ellmin$ and $\ellmax$ are the minimum and maximum value of $\ell_n$ and
+where $\ellmin$ and $\ellmax$ are the minimum and maximum of $\ell_1,\dots,\ell_n$ and
 \[
@@
     width=25cm,
-    height=15.83cm,
+    height=16cm,
     tickpos=left,
@@
 
-      \addplot[mark=none,color=black,dashed] table {../code/data/NormalisedLLSCLTN30geom10};
+      \addplot[mark=none,color=black,dashed,thick] table {../code/data/NormalisedLLSCLTN30geom10};
 
@@
 \addplot[mark=triangle*,mark options={scale=3},color=blue] table {../code/data/SLS2newN30geom10};
-\addplot[mark=none,color=black] table {../code/data/PeriodogramCLTN30geom10}; 
+\addplot[mark=none,color=black,thick] table {../code/data/PeriodogramCLTN30geom10}; 
  \addplot[mark=o,mark options={scale=2.5},color=black, thick] table {../code/data/NormalisedSamplingLLSN30geom10}; 
@@
 
-\legend{Quantised periodogram $q=1.5$, Quantised periodogram $q=5$,Seperable line search,Least squares Theory,Periodogram Theory, Least squares,Periodogram}
+\legend{Quantised periodogram $q=1.5$, Quantised periodogram $q=5$,Separable line search,Least squares Theory,Periodogram Theory, Least squares,Periodogram}
 
@@
 %    \selectcolormodel{gray} 
-    \begin{axis}[font=\footnotesize,xmode=log,ymode=log,height=15cm,width=25cm,xlabel={$N$},ylabel={time (ms)},ylabel style={at={(0.085,0.42)}},xlabel style={at={(0.5,0.13)}}, legend style={draw=none,fill=none,legend pos=north west,cells={anchor=west},font=\footnotesize},xmin=7,xmax=36000]
+    \begin{axis}[font=\footnotesize,xmode=log,ymode=log,height=15.25cm,width=25cm,xlabel={$N$},ylabel={time (ms)},ylabel style={at={(0.085,0.42)}},xlabel style={at={(0.5,0.13)}}, legend style={draw=none,fill=none,legend pos=north west,cells={anchor=west},font=\footnotesize},xmin=7,xmax=36000]
       \addplot[mark=*,mark options={scale=1.75},color=red] table {../code/data/PeriodogramBenchmark};
@@
       %\legend{Periodogram, Least squares, Fast Fourier transform, Chirp z-transform, SLS-new}
-      \legend{Periodogram, Least squares, Seperable line search, Quantised periodogram $q=5$}
+      \legend{Periodogram, Least squares, Separable line search, Quantised periodogram $q=5$}
    \end{axis}
@@
 
-\nocite{clarkson_estimate_period_pulse_train_1996,Sidiropoulos2005,5585849,Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}
+\nocite{Fogel1989_bit_synch_zero_crossings,Sidiropoulos2005,Clarkson2007,Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}
 
